---
title: "Basic Probability: Pt.3"
author: "Homer White"
date: "September, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# Continuous Random Variables

## Continuous Random Variable


> A **continuous random variable** is a random variable whose possible values lie in a range of real numbers.

## Examples

$X=\mbox{amount of sugar in a randomly-chosen  orange.}$

* Range of $X$ = $(0,\infty)$.

$Y=\mbox{height of a random student at GC}$

* Range of $Y=(0,\infty)$.

$Z=\mbox{grade of a random student in MAT 111} (\%)$

* Range of $Z=[0,100]$

## Probabilities 

For a *discrete* random variable you get probabilities from a table or a histogram:

Example:  Let $X$ = number of heads in 3 flips of a fair coin.

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

## Probability Histogram

```{r,echo=FALSE}
plotDist('binom',params=list(3,0.5),kind="histogram",type="percent", xlab="Number of Heads x")
```

## Probabilities

For a *continuous* random variable, you get probabilities from a smooth *density curve*.

```{r, echo=FALSE}
pchisqGC(3,region="above",df=5,graph=TRUE)
```

## Normal Random Variable

PDF for a **normal** random variable $X$:

* is called a "normal curve"
* it is bell-shaped
* Symmetric about $EV(X)=\mu$
* $SD(X)=\sigma$
* the 68-95 Rule applies

##

```{r, echo=FALSE}
plotDist("norm",params=list(mean=0,sd=1),kind='density',labels=FALSE)
```


## 68-95 Rule for Probability

When a random variable is roughly normal:

* Probability that $X$ falls within 1 standard deviation of the mean:

$$P(\mu-\sigma< X < \mu+\sigma)\approx 0.68$$

* Probability that $X$ falls within 2 standard deviations of the mean:

$$P(\mu-2\sigma< X < \mu+2\sigma)\approx 0.95$$

* Probability that $X$ falls within 3 standard deviations of the mean:

$$P(\mu-3\sigma< X < \mu+3\sigma)\approx 0.997$$

## Normal Random Variable:  Example


The heights of college males follow a normal distribution with $\mu=72$ inches and $\sigma=3.1$ inches.  Let 

$$X= \mbox{ height of a college male.}$$  

About 95% of college males are between what two heights? 

$P($ ________ < $X$ < ________$)$ $\approx$ 0.95

```{r,eval=FALSE}
require(manipulate)
EmpRuleGC(mean=72,sd=3.1)
```

## Normal Probabilities (Greater Than)

When you can't use the 68-95 Rule, use `pnormGC()`:.

To find $P(X>70.9)$:

```{r,eval=FALSE}
pnormGC(70.9,region="above",mean=72,
         sd=3.1,graph=TRUE)
```

##

```{r,echo=FALSE}
pnormGC(70.9,region="above",mean=72,
         sd=3.1,graph=TRUE)
```

## Normal Probabilities (Less Than)

To find $P(X<69)$:

```{r,eval=FALSE}
pnormGC(69,region="below",mean=72,
         sd=3.1,graph=TRUE)
```

##

```{r,echo=FALSE}
pnormGC(69,region="below",mean=72,
         sd=3.1,graph=TRUE)
```

## Normal Probabilities (Between)

To find $P(69 < X < 72)$:

```{r,eval=FALSE}
pnormGC(c(69,72),region="between",mean=72,
         sd=3.1,graph=TRUE)
```

##

```{r,echo=FALSE}
pnormGC(c(69,72),region="between",mean=72,
         sd=3.1,graph=TRUE)
```



## Normal Probabilities (Outside)


To find $P(X<69.9 \mbox{ or } X>75.1)$:

```{r,eval=FALSE}
pnormGC(c(68.9,75.1),region="outside",
        mean=72,sd=3.1,graph=TRUE)
```


##

```{r,echo=FALSE}
pnormGC(c(68.9,75.1),region="outside",mean=72,
         sd=3.1,graph=TRUE)
```

## Note on <= and >=

When a random variable is continuous,

$$P(X <= x)  = P(X < x).$$

for any value $x$.

Why?  Well,

$$P(X=x)=0.$$

## Really?

Yes!  Consider our normal random variable.  To get $P(X=73)$, we compute:

```{r,eval=FALSE}
pnormGC(c(73,73),region="between",
        mean=72,sd=3.1,graph=TRUE)
```


##

```{r,echo=FALSE}
pnormGC(c(73,73),region="between",
        mean=72,sd=3.1,graph=TRUE)
```

## Practice { .practice }

A population of bats has wingspans that are normally distributed, with a mean $\mu$ of 25 centimeters and a standard deviation of 5 centimeters.

Let $X=$ the wingspan of a randomly-captured bat.  Find:

* $P(X<27)$
* $P(X>29)$
* $P(20 < X < 30)$



# Percentiles

## Percentiale Ranks

Suppose we want to know the height of a male who is taller than 80% of college men.  We know:

$$P(X\leq x)=0.80$$

We need to find $x$.  It is called a the "80th percentile" of the heights.

## Computing Percentiles

To find the 80th percentile, compute:

```{r}
qnorm(0.80,mean=72,sd=3.1)
```


A 74.61 inch tall male is taller than 80% of college men.

We also say:  the 80th percentile of the heights of college males is 74.61 inches.

## Another Way

You can also use `qnormGC()`.  This will give you a graph, if you like:

```{r eval=FALSE}
qnormGC(0.80,region="below",mean=72,sd=3.1,graph=TRUE)
```

## The Result

```{r echo=FALSE}
qnormGC(0.80,region="below",mean=72,sd=3.1,graph=TRUE)
```



## Another Example

Find a height so that a person with that height is shorter than 10% of college males.

Such a person is taller than 90% of college males, so compute:

```{r}
qnorm(0.90,mean=72,sd=3.1)
```

Using `qnormGC()`:

```{r eval=FALSE}
qnormGC(0.10,region="above",mean=72,sd=3.1,graph=TRUE)
```

## The Result

```{r echo=FALSE}
qnormGC(0.10,region="above",mean=72,sd=3.1,graph=TRUE)
```



## Practice { .practice }

Back to the population of bats whose wingspans average 25 centimeters, with a standard deviation of 5 cm.

* Harry the Bat has a wingspan that is right at the 35th percentile for all the bats.  What is Harry's wingspan?

* Dolores the Bat has a wingspan that is longer than all but 25% of the bats.  What is her wingspan?
