---
title: "Basic Probability: Pt.1"
author: "Will Harris"
date: "September, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# Random Variables

## Random Variable

A **random variable** is a number whose value depends on the outcome of a chance process.

## Example

You flip a fair coin three times.

Let $X =$ the number of heads you get.

$X$ is a random variable.

## Example

You roll a six-sided die:

* If the six-spots side lands up, you win 10 bucks.
* If not, you lose a buck.

Let $Y=$ how much you win in one play.

$Y$ is a random variable.

## Example

You pick a person at random from a population.

Let $Z =$ his/her height.

$Z$ is a random variable.

## Example

* You take a simple random sample of 25 people from a population.
* You calculate the mean $\bar{x}$ of the sample.

$\bar{x}$ is a random variable.

# Probability Distributions

## Finding the Chances

For a random variable $X$, we would like to know the chances for $X$ to assume various values.

If we know how to do this, then we know the *distribution* of $X$.

## Example

We flip a fair coin three times.  The possible results are:

$$\{HHH,HHT,HTH,THH,\\
      HTT,THT,TTH,TTT\}$$
      
That's 8 possible outcomes.

Suppose that 

$$X = \textbf{number of heads}$$
      
## Notice:


When this is result|$X$ is
---|---
TTT|0
TTH|1
THT|1
HTT|1
THH|2
HTH|2
HHT|2
HHH|3

## Distribution of X

So we can find the "chances" for $X$ to be various values:

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

## Graph of the Distribution

```{r,echo=FALSE}
plotDist('binom',params=list(3,0.5),kind="histogram",type="percent", xlab="Number of Heads x")
```

## Using the Distribution

Now we can find many different probabilities.

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

<br>

$P(X \leq 2) = 1/8 + 3/8 + 3/8 = 7/8.$

## Graph


```{r,echo=FALSE}
plotDist('binom',params=list(3,0.5),kind="histogram",type="percent", 
         groups=x<=2,xlab="Number of Heads x")
```

The pink "shaded" area is $P(X \leq 2)$.

## Using the Distribution

Another example:

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

<br>

$P(X < 2) = 1/8 + 3/8 = 4/8 = 0.5.$

## Graph


```{r,echo=FALSE}
plotDist('binom',params=list(3,0.5),kind="histogram",type="percent", 
         groups=x<2,xlab="Number of Heads x")
```

The shaded area is $P(X < 2)$.


## Using the Distribution

Another example:

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

<br>

$P(X > 1) = 3/8 + 1/8 = 4/8 = 0.5.$

## Graph


```{r,echo=FALSE}
plotDist('binom',params=list(3,0.5),kind="histogram",type="percent", 
         groups=x>1,xlab="Number of Heads x")
```

The shaded area is $P(X > 1)$.

## The Random Variable Y

Recall:  you roll a six-sided die:

* If the six-spots side lands up, you win 10 bucks.
* If not, you lose a buck.

$Y=$ how much you win in one play.

## Distribution of Y

$y$|$P(Y=y)$
---|---
-1|5/6
10|1/6

# Expected Value

## What to Expect "On Average"

Suppose you play the dice game many, many times.

On each play, you get either 10 or -1.

About how much would you expect to win per game, *on average*?

This is called the *expected value* of $Y$, or $EV(Y)$ for short.

## Try It Out

Set up the game:

```{r}
y <- c(-1,10)
p.y <- c(5/6,1/6)
```

Play the game:


```{r eval=FALSE}
sample(y,size=1,p=p.y,replace=TRUE)
```

Run the code-chunk a few times.  Get a feel for what happens.

## Play it a LOT!

Let's play the game 10,000 times!

```{r eval=FALSE}
winnings <- sample(y,size=10000,p=p.y,replace=TRUE)
```

```{r echo=FALSE}
set.seed(3030)
winnings <- sample(y,size=10000,p=p.y,replace=TRUE)
```

Here's my first 20:

```{r}
winnings[1:20]
```

What did *you* get?

## On Average ...

I got

```{r}
mean(winnings)
```

What did *you* get?

## Expected Value

A formula:

$$EV(Y)= \sum_y y \cdot P(Y=y) \\
= (-1) \times \frac{5}{6} + 10 \times \frac{1}{6} \\
= \frac{5}{6} \approx 0.83.$$

## It Makes Sense!

When you play the dice game many, many times:

* you expect -1 about $5/6^{th}$ of the time
* you expect 10 about $1/6^{th}$ of the time.

$$EV(Y)=(-1) \times \frac{5}{6} + 10 \times \frac{1}{6}.$$

## What Does R Say?

```{r}
expected <- sum(y*p.y)
expected
```

## Coin Flip

Recall flipping the fair coin three times, $X$ = number of heads.

$x$|$P(X=x)$
---|---
0|1/8
1|3/8
2|3/8
3|1/8

## R Says

Set it up:

```{r}
x <- c(0,1,2,3)
p.x <- c(1/8,3/8,3/8,1/8)
```

Then $EV(X)$ is:

```{r}
sum(x*p.x)
```

## Try it Out:

```{r eval=FALSE}
heads <- sample(x,size=10000,p=p.x,replace=TRUE)
mean(heads)
```
```{r echo=FALSE}
set.seed(3030)
heads <- sample(x,size=10000,p=p.x,replace=TRUE)
mean(heads)
```

Did you also get near 1.5?

# Standard Deviation

## Give-or-Take

So, when you "try" a random variable $X$, you expect to get around $EV(X)$ ...

... give or take some for chance variation!

What's the "give-or-take" figure?

## Standard Deviation

$$SD(X) = \sqrt{\sum_x (x-EV(X))^2 P(X=x)}.$$

## Example

$x$|$x-EV(X)$|$(x-EV(X))^2$|$P(X=x)$
---|---|---|---
0|-1.5|2.25|1/8
1|-0.5|0.25|3/8
2|+0.5|0.25|3/8
3|1.5|2.25|1/8

<br>

$SD(X)=\sqrt{2.25/8+0.75/8+0.75/8+2.25/8} \approx 0.867.$

## So ...

When you flip a fair coin 3 times, the number of heads will differ from the expected 1.5 by about 0.867 or so.

* 1.5 - 0.867 = 0.633
* 1.5 + 0.867 = 2.367

The possible values in this range are:  1 and 2.

So typically you are supposed get 1 or 2 heads.

$P(X=1) + P(X=2) = 3/8 +3/8 = 6/8 =\textbf{ }$ a 75% chance.

Yes, that seems "typical"!

## Two Games

The rules for Game A are:

* flip a fair coin once
* if it's Heads, you win 4 dollars
* if it's Tails, you win 2 dollars

The rules for Game B are:

* flip a fair coin twice
* if it lands Heads both times, you win 40 dollars
* otherwise, you lose 8 dollars

## Many Plays

You can play Game A 10,000 times, or you can play Game B 10,000 times.

## Practice { .practice }

Which would you choose?  Why?

## Suggestion:  Compare  EV's

Compute EV(A) using:

$a$|$P(X=a)$
---|---
2|1/2
4|1/2

Compute EV(B) using:

$b$|$P(X=b)$
---|---
-8|3/4
40|1/4

## Another Suggestion: Simulation

For Game A:

```{r eval=FALSE}
a <- c(4,2)
p.a <- c(0.5,0.5)
winningsA <- sample(a,size=10000,p=p.a,replace=TRUE)
sum(winningsA)
```


For Game B:

```{r eval=FALSE}
b <- c(40,-8)
p.b <- c(0.25,0.75)
winningsB <- sample(b,size=10000,p=p.b,replace=TRUE)
sum(winningsB)
```




