---
title: "Basic Probability: Pt.2"
author: "Homer White"
date: "September, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# Binomial Random Variables

## The Binomial "Situation"

Suppose that:

* You perform a specific number $n$ of trials.

* Each trial has two outcomes: *success* or *failure*.

* The chance of success $p$ is the same on each trial.

* The outcome of one trial does not affect the outcome of any other trial.  (The trials are "independent" of each other.)

## Binomial Random Variable

Let $X =$ the number of successes in the $n$ trials.

$X$ is called a **binomial** random variable.

Sometimes for short we write:

$$X \sim binom(n,p).$$

## Example

$X=$ number of heads in three flips of a fair coin.

This is a binomial random variable.  Why?

* Each flip is a "trial".  ($n=3$)

* Two outcomes for each trials:
    * *success* = Heads, 
    * *failure* = Tails

* Probability of a *success* = $p=0.5$ is the same for each flip. 

* The outcomes of the flips don't affect each other (independence).

## Another Example

* You can make 70% of the free-throw shots you attempt.
* Your shots don't affect each other.
* You go to the line and take 20 shots.

Let $Y =$ the number of shots you make.

$$Y \sim binom(20,0.7).$$

## Another Example

Buzz the dolphin chooses a button to push, possibly with help from Doris.

* He does it 16 times.
* Each time he either chooses the correct button, or the wrong one.
* Each time his chance of getting it right is $p$.
* His tries don't affect each other.

Let $Z =$ the number of times he pushes the correct button.

$$Z \sim binom(16,p).$$

## A Mystery

What is $p$?

* If Buzz is just guessing randomly, then $p=0.5$.
* If he can communicate with Doris to some extent, then $p > 0.5$.

## A Non-Example

* Kindergarten class has 20 children.
* Each child either will or will not get the flu this winter.
* 10% of all kindergarten children get the flu.

Let $Z$= number of children in the class who get the flu this winter.

## Analysis

* Each child is a "trial":  $n=20$
* *success*=Flu, *failure* = No Flu
* Probability of a *success* is 0.10.

But ...

... the children are not *independent*.  If one child in the class gets the flu, others become more likely to get it.

## Practice { .practice }

* You choose twenty people at random.
* They are not related to each other.
* You ask each person how many brothers he/she has.

Let $X=$ the total number of brothers reported.

Is $X$ binomial?  If so, what is $n$ and what is $p$?

## Practice { .practice }

* You sample 71 people at random from a large population.
* You record the sex of each person (male/female).
* 55% of the population is female

Let $X=$ the total number of females in the sample.

Is $X$ binomial?  If so, what is $n$ and what is $p$?

# Binomial Probabilities

## Taking a Test

* You have a multiple choice test with 20 questions.
* Each questions has four choices, but only one is correct.
* You guess on each question.
* To pass, you must get at least 8 questions correct.
* What is the chance that you fail?

## Analysis

If you guess, you have a 25% chance of getting the question right.

Let $X=$ number of questions correct.

Then $X \sim binom(20,0.25)$.

We are looking for:

$$P(X \leq 7)$$.

## Solution

Use `pbinomGC()`:

```{r eval=FALSE}
pbinomGC(7,size=20,prob=0.25,graph=TRUE)
```

## Result

```{r echo=FALSE}
pbinomGC(7,size=20,prob=0.25,
         region="below",
         graph=TRUE)
```

## Another Example

* Suppose Buzz is just a guesser (no communication with Doris).

* What's the chance that he pushes the correct button 15 or more times out of 16 tries?

## Analysis

Let $Z =$ number he gets correct.

We see that $Z \sim binom(16,0.5).$

We want:

$$P(X \geq 15).$$

## Solution

```{r eval=FALSE}
pbinomGC(14,size=16,prob=0.5,
         region="above",
         graph=TRUE)
```

We do this because $X > 14$ means the same thing as $X \geq 15$.

## Result

```{r echo=FALSE}
pbinomGC(14,size=16,prob=0.5,
         region="above",
         graph=TRUE)
```

## Example

* You toss a fair coin 100 times.
* What is the chance of getting between 45 and 55 heads (including the 45 and the 55)?

Let $X=$ the number of heads.  We want $P(45 \leq X \leq 55)$.

```{r eval=FALSE}
pbinomGC(c(45,55),size=100,prob=0.5,
         region="between",
         graph=TRUE)
```

## Result

```{r echo=FALSE}
pbinomGC(c(45,55),size=100,prob=0.5,
         region="between",
         graph=TRUE)
```

## Practice { .practice }

If Buzz can get it right only 70% of the time when Doris helps, then what is the chance that he gets it right 15 or more times out of 16 tries?

If you flip a fair coin 100 times, what is the chance of getting:

* no more than 45 heads?
* less than 47 heads?
* between 40 and 60 heads?
* exactly 50 heads?

# EV and SD for Binomials

## Expected Value and SD

If $X \sim binom(n,p)$, then

$$EV(X) = np,$$

and

$$SD(X) = \sqrt{np(1-p)}.$$

## Example

You flip a fair coin 100 times.

* $EV(X) = np = 100 \times 0.5 = 50.$

* $SD(X) = \sqrt{np(1-p)}=\sqrt{100(0.5)(1-0.5)}=5.$

So when flip a fair coin 100 times, you can expect to get around 50 heads, give or take 5 or so for chance variation!

## Practice { .practice }

You are a 70% free-throw shooter.  You shoot 20 baskets.

* How many can you expect to hit?
* Give or take about how many?

# Shapes of Binomials

## An App

Explore this app:

```{r eval=FALSE}
require(manipulate)
BinomNorm()
```

## Observations

* No matter what $p$ is, as $n$ gets larger the distribution gets more "bell-shaped".

Many people think that it looks quite bell-shaped as long as:

* $np \geq 10,$ and 
* $n(1-p) \geq 10.$


## The 68-95 Rule for Binomials

Let $X$ be a binomial random variable.

If the distribution of $X$ looks roughly bell-shaped, then:

* there is about a 68% chance that $X$ will be within one SD of its EV.
* there is about a 95% chance that $X$ will be within two SDs of its EV.
* there is about a 99.7% chance that $X$ will be within three SDs of its EV.

## Example

You flip a fair coin 400 times.  Let $X=$ the number of heads you get.

* $EV(X) = np = 400 \times 0.5 = 200.$
* $SD(X) = \sqrt{400(0.5)(1-0.5)} = 10.$

* There is about a 68% chance that you will get between 190 and 210 heads.
* There is about a 95% chance that you will get between 180 and 220 heads.

## Practice { .practice }

When you flip the fair coin 400 times:

* About what is the chance that you will get fewer than 190 heads?
* More than 220 heads?

Use the 68-95 rule to answer these questions, and then compare your answers to the exact probabilities:

```{r eval=FALSE}
pbinomGC(189,size=400,prob=0.5,
         region="below",graph=TRUE)
```

```{r eval=FALSE}
pbinomGC(220,size=400,prob=0.5,
         region="above",graph=TRUE)
```


