---
title: "Cheat Sheet"
author: "Homer White"
date: "September, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# One Factor Variable

## Graphics

$$barchartGC(\sim variable, data = MyData)$$

```{r eval=FALSE}
barchartGC(~seat,data=m111survey)
```


## The Graph


```{r echo=FALSE}
barchartGC(~seat,data=m111survey)
```



## Numerical Summaries

`xtabs()` and `rowPerc()`:

```{r}
seating <- xtabs(~seat,data=m111survey)
seating
```

```{r}
rowPerc(seating)
```



# Two Factor Variables

## Graphics

$$barchartGC(\sim exp + resp, data = MyData)$$

```{r eval=FALSE}
barchartGC(~sex+seat,data=m111survey)
```

## The Graph

```{r echo=FALSE}
barchartGC(~sex+seat,data=m111survey)
```

##Numerical Summaries

`xtabs()` and `rowPerc()`:

```{r}
sexSeat <- xtabs(~sex+seat,data=m111survey)
sexSeat
```

```{r}
rowPerc(sexSeat)
```

## Inference

Use `chisqtestGC()`:

```{r eval=FALSE}
chisqtestGC(~sex+seat,data=m111survey)
```

## Some Output

```
Pearson's Chi-squared test 

Observed Counts:
        seat
sex      1_front 2_middle 3_back
  female      19       16      5
  male         8       16      7

Counts Expected by Null:
        seat
sex      1_front 2_middle 3_back
  female   15.21    18.03   6.76
  male     11.79    13.97   5.24
```
  
## More Output

```
Contributions to the chi-square statistic:
        seat
sex      1_front 2_middle 3_back
  female    0.94     0.23   0.46
  male      1.22     0.29   0.59


Chi-Square Statistic = 3.734 
Degrees of Freedom of the table = 2 
P-Value = 0.1546 
```


# One Numeric Variable


## Graphics

`histogram()`, `densityplot()`, or `bwplot()`.

$$function(\sim variable,data=myData)$$

```{r eval=FALSE}
densityplot(~fastest,data=m111survey)
```

## The Graph

```{r echo=FALSE}
densityplot(~fastest,data=m111survey)
```

## Numerical Summaries

Use `favstats()`:

```{r}
favstats(~fastest,data=m111survey)
```


# One Factor and One Numeric

## Graphics

$$histogram(\sim numeric \vert factor, data=MyData)$$

$$densityplot(\sim numeric \vert factor, data=MyData)$$

$$bwplot(numeric \sim factor, data=MyData)$$

```{r eval=FALSE}
densitplot(~fastest|sex,data=m111survey)
```

## The Graph

```{r echo=FALSE}
densityplot(~fastest|sex,data=m111survey)
```

## Numerical Summaries

`favstats()` again:

$$favstats(numeric \sim factor, data=myData)$$

```{r}
favstats(fastest~sex,data=m111survey)
```



# Two Numeric Variables

## Graphics

Scatter plots:

$$xyplot(response \sim explanatory, data = myData)$$

```{r eval=FALSE}
xyplot(GPA~fastest,data=m111survey,type=c("p","r"))
```

## The Graph


```{r echo=FALSE}
xyplot(GPA~fastest,data=m111survey,type=c("p","r"))
```

## Numerical Summaries

$$lmGC(response \sim explanaotry, data=myData)$$

```{r eval=FALSE}
lmGC(GPA~fastest,data=m111survey)
```

## The Linear Model Output


```{r}
lmGC(GPA~fastest,data=m111survey)
```


