---
title: "Confidence Intervals: Pt.3"
author: "Homer White"
date: "April, 2016"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(tigerstats)
```



# Intervals for One Proportion

## Research Question


> Are a majority of GC students female?

We will investigate this question with the `m111survey` data.



## Recall:  Four-Step Procedure


1.  Define the parameter of interest.
2.  Run the code, and perform a "safety check".
3.  Report the confidence interval and interpret it.
4.  Use the interval to answer your Research Question.

## Step 1:  Parameter


Let:

* $p=$ the proportion of all GC students that are female.

If $p > 0.5$, then a majority of GC students are female.

We will make a 95% confidence interval for $p$.


## Step 2 (1):  Run Code


Run the code:

```{r,eval=FALSE}
binomtestGC(~sex,data=m111survey,
            success="female")
```

Additional argument, `success`, specifies which value of the variable is of interest.


## The Output

```
Descriptive Results:

40 successes in 71 trials

Inferential Results:

Estimate of p:  0.5634
SE(p.hat):  0.0589

95% Confidence Interval for p:

lower.bound       upper.bound
0.440455          0.680850
```



## Step 2 (2):  Safety Check


Similar to before:

* Need to have taken a simple random sample from the population.

* No restriction on sample size when using `binomtestGC()` because this function does not rely on any kind of an approximation.

## Are We Safe?


At least with regards to the variable **sex**, the `m111survey` data is like a simple random sample of the GC population.


## Step 3 (1):  Report Interval

Report it:

``` 
95% Confidence Interval for p:

lower.bound          upper.bound
0.440455             0.680850
```



## Step 3 (2):  Interpret the Interval


* Based on the data at hand, we are 95% confident that $p$ is somewhere between 0.44 and 0.68.





## Step 4:  Conclusion

Use the interval to answer the Research Question:

* The interval (0.44 to 0.68) contains some values below 0.5,
* and some values above 0.5!
* We don't have enough evidence to conclude that $p > 0.5$

## Note

In the output we saw:

```
Estimate of p:  0.5634
SE(p.hat):  0.0589
```

* The estimate of $p$ was $\hat{p} = 40/71$
* $SE(\hat{p})$ = 0.0589

## A Rough 95%-Confidence Interval

The 69-95 Rule for Estimation says that a rough 95%-confidence interval for $p$ is:

$$\hat{p} \pm 2 \times SE(\hat{p})$$

```{r}
0.5634 - 2*0.0589 #lower bound
0.5634 + 2*0.0589 #lower bound
```

Differs a bit from the more exact interval given by `binomtestGC()`!

## Practice { .practice }

```{r eval=FALSE}
View(m111survey)
```

Do more than one-third of all GC students prefer to the in the Middle?

To answer this question, make a 95%-confidence interval for $p$, the proportion of all GC students who prefer the middle.


  
## Additional Example

Research Question:


* A simple random sample of 1000 U.S. adults showed that 49% are in favor of the Affordable Care Act.  What is the percentage of all Americans that favor this law?

## Step 1:  Parameter


Let: 

* $p=$ the proportion of all Americans in favor of the Affordable Care Act.

We will make a 95% confidence interval for $p$.

## Step 2 (1):  Run Code

Since we are given summary data instead of access to a data set in R, we run the code:


```{r eval=FALSE}
binomtestGC(x=490,n=1000)
```

where

* $x$ is the number of successes in the sample.

* $n$ is the sample size.

## Step 2 (2):  Safety Check


* We are told this is a simple random sample.

* Since we are using `binomtestGC()`, there is no restriction on sample size.

  
## Step 3:  Report & Interpret Interval

```
95% Confidence Interval for p:

lower.bound     upper.bound
0.458585        0.521474
```

**Interpretation**:  Based on the data at hand, we are 95% confident that $p$ is somewhere between 0.46 and 0.52.

## Step 4:  Conclusion

(The Research Question was not very specific, so our conclusion cannot do much more than restate the interpretation of the confidence interval.)



## Practice { .practice }

Out of 2500 randomly-selected voters, 1200 says that they approve of the job that Barak Obama is doing..

Make a 99%-confidence interval for $p$, the proportion of all voters who approve of the job Obama is doing.

(Hint:  Add the argument `conf.level=0.99` to your code.)


# Difference of Two Proportions


## Research Question


*In the GC population, who is more likely to believe in extra-terrestrial life:  gals or guy?*

We will investigate this question with the `m111survey` data.

## Step 1:  Parameters


Let

* $p_1=$ proportion of all GC gals that believe in extra-terrestrial life.
* $p_2=$ proportion of all GC guys that believe in extra-terrestrial life.



* Explanatory variable is **sex**, a factor.
* Response variable is **belief in extra-terrestrial life**, also a factor.
* Both variables have two values.
* We are interested in estimating the parameter $p_1-p_2$ using a 95% confidence interval.

## Step 2 (1):  Run Code

We will use a new function called `proptestGC()`.

```{r eval=FALSE}
proptestGC(~sex+extra_life,data=m111survey,
           success="yes",first="female")
```

Note the input format (for two factor variables)

$$\sim  explanatory + response$$

Important:

* `success` specifies the value of the factor variable **extra_life** that we want to count
* `first="female"` guarantees that the females will be the first group, males second.

## Some Output

```
Descriptive Results:

       yes  n estimated.prop
female  10 40         0.2500
male    20 31         0.6452
```

Safety Check:

  * $10 \geq 10$ gals that believe (success)
  * $30 \geq 10$ gals that do not believe (failure)
  * $20 \geq 10$ guys that believe (success)
  * $11 \geq 10$ guys that do not believe (failure)
  
## More Output

```
Estimate of p1-p2:	 -0.3952 
SE(p1.hat - p2.hat):	 0.1099
```

The statistic is

$$\hat{p_1}-\hat{p_2} = \dfrac{X_1}{n_1}-\dfrac{X_2}{n_2}$$

where 

* $X_1=$ # of gals that believe in extra-terrestrial life,
* $X_2=$ # of guys that believe in extra-terrestrial life,
* $n_1=$ # of gals in sample, $n_2=$ # of guys in sample.

This time $\hat{p_1}-\hat{p_2} = -0.3952.$

  
## Step 3 (1):  Report Interval

```
95% Confidence Interval for p1-p2:

lower.bound     upper.bound
-0.610510       -0.179812
```

## Step 3 (2):  Interpret Interval


* Based on the data at hand, we are 95% confident that $p_1-p_2$ is somewhere between -0.61 and -0.18.

Since $p_1$ was defined to be the proportion of gals and $p_2$ the proportion of guys, 

* negative values for $p_1-p_2$ indicate that the proportion of guys that believe in extra-terrestrial life is bigger than the proportion of gals.


## Step 4:  Conclusion/Answer


(Here is where we use the interval to answer the Research Question, as specifically as possible.)

* The confidence interval contains only negative values (corresponding to gals being less likely than guys to be believe in ET life).
* We can be pretty confident that GC gals are less likely to believe in ET-life than GC guys are.

## Another Example:  Union Membership

**Research Queston:**

>*In 1985, who was more likely to belong to a union:  a male worker or a female worker?*

## Descriptive Look

```{r}
sexUnion <- xtabs(~ sex + union, data = CPS85)
sexUnion
```

Check row percents:

```{r}
rowPerc(sexUnion)
```

## Inferential Work

Let

* $p_1=$ proportion of all male employees in 1985 who belonged to a union.
* $p_2=$ proportion of all female employees in 1985 who belonged to a union.

```{r eval = FALSE}
proptestGC(~ sex + union, data = CPS85,
           success = "Union", first = "M")
```

## Results

```
Estimate of p1-p2:	 0.121 
SE(p1.hat - p2.hat):	 0.03218 

95% Confidence Interval for p1-p2:

          lower.bound         upper.bound          
          0.057930            0.184086             
```


## Practice { .practice }

```{r eval=FALSE}
data(ucdavis1)
View(ucdavis1)
help(ucdavis1)
```

Let $p_1$ be the proportion of all UC-Davis females who prefer to sit in the front.

Let $p_2$ be the proportion of all UC-Davis males who prefer to sit in the front.

Make a 90%-confidence interval for $p_1-p_2$.

Hints:

* use the `success` argument to count people who prefer the Front
* use the `conf.level` argument for confidence levels other than 95%
* use the `first` argument to say which group should come first


## Example:  Summary Data


The function `proptestGC()` can be used with summary data as well:

Suppose we took two samples from two populations and:

* there were 23 successes out of 100 trials in the first sample;
* there were 33 out of 110 successes in the second sample.  

```{r eval=F}
proptestGC(x=c(23,33),n=c(100,110))
```

## Results

```
Descriptive Results:

        successes   n estimated.prop
Group 1        23 100           0.23
Group 2        33 110           0.30

Inferential Results:

Estimate of p1-p2:   -0.07 
SE(p1.hat - p2.hat):   0.06066 

95% Confidence Interval for p1-p2:

lower.bound         upper.bound          
-0.188899           0.048899
```

## Practice { .practice }

* In a random sample of 2000 adults in the year 1940, 800 had completed high school.
* In a random sample of 2500 adults in the years 1991, 1900 had completed high school.

Let $p_1$ be the proportion of all adults in 1991 who had completed high school.

Let $p_2$ be the proportion of all adults in 1940 who had completed high school.

Find a 95%-confidence interval for $p_1-p_2$.
