---
title: "DAR Draft Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(tigerstats)
```

# A Detail To Start With

## A Possible Outlier

```{r}
bwplot(~ wage, data = CPS85)
```

One person makes a lot more than everyone else!

## Who Is This?

```{r}
CPSRich <- subset(CPS85, wage > 40)
CPSRich
```

## Decide

Should you exlude the outlier?  (Maybe it's a typo.  For sure it will mess up graphs a bit!)

Or keep it?  Or maybe keep it in numerical summaries, but exclude it from graphs.

Make your choice and explain it to the reader in the Methods section.

## Removing the Outlier

If you decide to remove the outlier, you can do it with:

```{r}
CPS2 <- subset(CPS85, wage < 44)
```

Then use `CPS2`.


# Understanding the Assignment

## The Assignment

>"Your assignment involves using `CPS85` to investigate the question of whether women are the victims of wage discrimination.  If a man and a woman are alike in all respects---age, education, work experience, type of job, etc.---that might affect wage, does the man still make more money?"

Let's think about what this means, step-by-step.  Here is ONE POSSIBLE way  to "unpack" the assignment's meaning.

## A Thesis

You could have a thesis:

>*Women are the victims of sex discrimination.*

## Reason for the Thesis

>Women are being paid less than men.

## Back It Up! 

You can back it up with numbers:

```{r}
favstats(wage ~ sex, data = CPS85)[c("sex","mean","median","n")]
```

## Back It Up!

You can (and should) also back it up with a graph:

```{r eval = FALSE}
bwplot(wage ~ sex,
	data = CPS2,
	panel = function(..., box.ratio) {
		panel.bwplot(..., fill = NULL, box.ratio = 0.1)
		panel.violin(..., col = "burlywood", border = "grey60", alpha = 0.5, varwidth = FALSE)},
	xlab = "Sex",
	ylab = "wages ($/hour)")
```

Labels, captions, etc!

## The Graph

```{r echo = FALSE}
bwplot(wage ~ sex,
	data = CPS2,
	panel = function(..., box.ratio) {
		panel.bwplot(..., fill = NULL, box.ratio = 0.1)
		panel.violin(..., col = "burlywood", border = "grey60", alpha = 0.5, varwidth = FALSE)},
	xlab = "Sex",
	ylab = "wages ($/hour)")
```

## Objections

But there are objections to consider.

## The "Education" Objection

Someone could raise the following objection:

>*The wage gap between men and women is due not to sex discrimination, but rather to differences in education between men and women.*

## Understanding the Objection

The objection really involves two claims:

1. Men have more education than women.
2. If you were to compare men and women who have the same level of education, then the wage gap would vanish.

You need to explain this to your reader!

## Responding to the Objection

You have two choices:

* You could attack Claim 1 by showing that men don't have more education than women.
* You could attack Claim 2 by "correcting" for education and showing that the wage gap persists.

It's OK to do both!

## Attacking Claim 1

To attack Claim 1, you need to examine the relationship between **educ** and **sex**.

## Attacking Claim 1 (Graphical)

One possibility:

```{r eval = FALSE}
densityplot( ~ educ,
	data = CPS2,
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	xlab = "education level (years completed)")
```

## The Graph

```{r echo = FALSE}
densityplot( ~ educ,
	data = CPS2,
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	xlab = "education level (years completed)")
```

## Attacking Claim 1 (Numerical)

You should complement this with a numerical analysis:

```{r}
favstats(educ ~ sex, data = CPS85)[c("sex","median","mean","n")]
```

## Attacking Claim 2

You could also deal with the objection by attacking Claim 2:  "correct" for **educ** and see that the wage gap persists.

## Graphical Attack

**wage** and **educ** are both numerical.  So one possibility is to look at a scatter plot, but group by **sex**.

## The Attack Code

```{r eval = FALSE}
xyplot(wage ~ educ,
	data = CPS2,
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	type = c("p","smooth"),
	pch = 20,
	xlab = "Education (years)",
	ylab = "Wages ($/hr)")
```

## The Graph

```{r echo = FALSE}
xyplot(wage ~ educ,
	data = CPS2,
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	type = c("p","smooth"),
	pch = 20,
	xlab = "Education (years)",
	ylab = "Wages ($/hr)")
```


## Numerical Attack

One possible approach:  "cut" **educ** into groups, then break down by the groups.

```{r eval = FALSE}
EdGrp <- with(CPS85, cut(educ, breaks = 4))
favstats(wage ~ sex + EdGrp, 
         data = CPS85)[c("sex.EdGrp","median","mean","n")]
```

## The Table

```{r echo = FALSE}
EdGrp <- with(CPS85, cut(educ, breaks = 4))
favstats(wage ~ sex + EdGrp, 
         data = CPS85)[c("sex.EdGrp","median","mean","n")]
```

## Choices

You can attack the "Education" Objection at either point, or at both points.  The choice is up to you.

You can use graphs or numbers, or both.  You should use both if it seems that one approach alone does not conclusively make your point.

Use your judgement.

## Many Other Objections:

* the "age" objection
* the "experience" objection
* the "geography" objection
* the "marital status" objection

Some are more plausible than others, but you should consider them all.

## Further Suggestions

* You could include your thesis in the Introduction.
* In the Methods section you could:
    * Provide the basic argument for your thesis.
    * Then consider just one objection, and *explain in detail* what claims it involves.
    * Then *explain* how each claim could be attacked.
* Then in Results you could deal with each objection, one per sub-section.
* Some other variables don't make for plausible objections, but correcting for them can lead to special insight.  When/if this happens, include your analysis of that variable in your report, under Discussion/Conclusion.

# Graphs and Tables

## Pretty, and with Captions!

* Graphs should have captions.  (Use `fig.cap`.)
* When a graph has a good caption it does not need a title.
* Use the caption to explain technical points of a graph (i.e., that the black dots in a box-and-whisker plot represent medians).
* Conclusions drawn from the graph usually belong in the surrounding text, not in the caption itself.
* Tables should be done with `kable()` as much as possible.
* They should have captions, using the `caption` argument.

## Graph/Table Placement

You have two choices:

* You can put all graphs and tables in a separate, final section called **Appendix**.  If you have both graphs and tables there, then it should have a subsection for tables and another subsection for graphs.
* Or, you can try to keep graphs and tables very close to the places in the report where they are discussed.

## Tips on Keeping Graphs Close

* Use the `fig.height` and `fig.width` chunk options.  Smaller graphs are less likely to get moved far away from where you want them.
* You can make page breaks by inserting `\newpage` (have one blank line before and one blank line after each such insertion).

# Final Advice

## Read the Directions, Use the Resources ...

* Read the Assignment Again
* Read the Supporting Documents:
    * The Sample Report
    * The Rubrics
    * The SAT slides
    * Watch the R Markdown Video


## ... and Put in the Effort!

* Start revising now, not two days before deadline
* Go to the Writing Center
* Use the Spell Checker (Edit -> Check Spelling) ...
* ... and then keep checking your spelling and grammar!
* Ask when you are stuck on anything
* But don't bother to ask at the lat minute!




