---
title: "DAR Final Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(tigerstats)
library(knitr)
```

# Methods

## A Possible Outlier

```{r}
bwplot(~ wage, data = CPS85)
```

One person makes a lot more than everyone else!

## Who Is This?

```{r}
CPSRich <- subset(CPS85, wage > 40)
CPSRich
```

## Decide

Some folks decided to keep the outlier, thinking there was no reason the woman was not "real."

Some kept her in numerical summaries, but not in graphs (so as not to waste graph space).

## Removing the Outlier

If you decided to remove the outlier, you could have done it with:

```{r}
CPS2 <- subset(CPS85, wage < 44)
```

## Confouding Factors

At some point in methods it was important to:

* show that men were making more money than women
* but point out that confounding factors (rather than sex) could explain the wage
* and explain how one might investigate confounding factors (e.e., by correcting for them)

# Results

## Important Things

* Consider all possible confounding factors
* find graphs where pattern is easiest to see
* back up graphs with numerical summaries, especially when graphs are difficult to read


## Interesting Insights:  Marital Status

Marital Status is NOT a confounding factor, because it isn't related to **sex**:

```{r}
rowPerc(xtabs(~ sex + married, data = CPS2))
```

## Interesting Insights:  Marital Status

But watch what happens when you correct for marital status anyway:

```{r eval = FALSE}
favstats(wage ~ sex + married, data = CPS2)[c("sex.married","mean","n")]
```

```{r echo = FALSE}
kable(favstats(wage ~ sex + married, data = CPS2)[c("sex.married","mean","n")])
```


## Correct for Two (or More) at Once?

One person wondered if men had more education, after breaking down by sector:

```{r eval = FALSE}
favstats(educ ~ sex + sector, data = CPS2)[c("sex.sector", "mean", "n")]
```

```
   sex.sector     mean  n
1  F.clerical 12.84211 76
2  M.clerical 13.28571 21
3     F.const      NaN  0
4     M.const 11.15000 20
5     F.manag 14.25000 20
6     M.manag 14.79412 34
7     F.manuf 10.66667 24
8     M.manuf 11.47727 44
```

## Continued ...

```
9     F.other 12.83333  6
10    M.other 11.72581 62
11     F.prof 15.28846 52
12     M.prof 15.98113 53
13    F.sales 12.76471 17
14    M.sales 13.57143 21
15  F.service 11.65306 49
16  M.service 11.52941 34
```

Questionable whether men are significantly more educated after taking sector into account, but still, could we correct for **educ** and **sector** simultaneously?

##  One Way

```{r}
CPSccmm <- subset(CPS2, 
  sector %in% c("clerical", "construction", "manag","manuf"))
CPSopss <- subset(CPS2, 
  sector %in% c("other", "prof", "sales","service"))
```

##

```{r echo = FALSE}
xyplot(wage ~ educ | sector,
	data = CPSccmm,
	layout = c(2,2),
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	type = c("p","smooth"), span = 1.5,
	pch = 20)
```

##

```{r echo=FALSE}
xyplot(wage ~ educ | sector,
	data = CPSopss,
	layout = c(2,2),
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	type = c("p","smooth"), span = 1.5,
	pch = 20)
```

# Conclusion/Discussion

## Correcting for "All at Once"

Are there ways to correct for all confounding variables (in the data set) at once?

## Population vs. Sample

Women make less than men after correcting for confounding factors, but this is a pattern we see in the sample.  Does it carry over to the population (i.e., is it statistically significant)?

## Wage Gap Shrinking?

```{r echo = FALSE}
xyplot(wage ~ age,
	data = CPS2,
	groups = sex,
	auto.key = list(
		space = "top",
		title = "Sex",
		cex.title = 1,
		columns = 2),
	type = c("p","smooth"),
	pch = 20)
```

Need a longitudinal study!

## Why the Wage Gap?

The wage gap persists to this day, but is it still due to outright wage discrimination?  Or are women less assertive than men when it comes to raises, promotions, etc?

These matters are currently being studied.


