---
title: "DAR Ch.9 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Plagiarism Issues

## Reminders (from Ch 4 DAR Feedback):

* You should leave the Introduction untouched (it sets the assignment)
* Use your own words in other sections.
    * Really, really use your own words!
    * Don't just change a word or two from each sentence in the example document!
* Advice:  read the example document in advance, but don't look at it while writing up your own assignment

**Copying is plagiarism!**

# Methods

## Variable Analysis

You must include this in your methods!

Two variables involved:

* **conc.situation**, a factor variable with two values:
    - `lost.money`
    - `lost.ticket`
* **conc.decision** a factor variable with two values:
    - `buy`
    - `not.buy`

We wonder if the way you lost the \$30 affects your decision, so best to pick **conc.situation** as the explanatory variable.

## Descriptive Methods

Two factor variables leads to:

* `barchartGC()` for a bar-chart (graphical descriptive)
* `xtabs() and rowPerc()` for two-ways tables (numerical descriptive)

## Inferential Methods

* Since response has two values (buy or not) we are interested in proportions.

* Since explanatory has two values (lost money vs. lost ticket), we will be interested in the difference of two proportions.

## Define Parameters (One Good Choice)

This was a randomized experiment.  (See Chapter 8 Part 2 website slides).  If we think the 267 are like a random sample from the population of all GC students, then we could define:

* $p_1 =$ the proportion of all GC students who would buy a ticket if they had lost \$30 in cash

* $p_2 =$ the proportion of all GC students who would buy another ticket if they had lost a previously-purchased $30 ticket

## Define Parameters (Another Good Choice)

If we think the 267 are NOT like a representative sample of the population of all GC students, then we could define:

* $p_1 =$ the proportion of all 267 students in the experiment who would buy a ticket if they had lost \$30 in cash

* $p_2 =$ the proportion of all 267 students in the experiment who would buy another ticket if they had lost a previously-purchased $30 ticket

Either way, we will make a confidence interval for $p_1-p_2$ using `proptestGC()`.

# Graphical Descriptive Results

## barchartGC()

Easiest to make a two-way table, then keep re-using it:

```{r echo=FALSE}
sitDec <- xtabs(~conc.situation+conc.decision,data=attitudes)
```


```{r eval=FALSE}
sitDec <- xtabs(~conc.situation+conc.decision,data=attitudes)
barchartGC(sitDec,type="percent",
           main="Concert Decision,\nby Situation on Survey Form",
           xlab="Scenario on Survey Form")
```

---

```{r echo=FALSE}
barchartGC(sitDec,type="percent",
           main="Concert Decision,\nby Situation on Survey Form",
           xlab="Scenario on Survey Form")
```

## Also Popular:  Flat Barcharts!

```{r eval=FALSE}
barchartGC(sitDec,type="percent",
           main="Concert Decision,\nby Situation on Survey Form",
           ylab="Scenario on Survey Form",
           flat=TRUE)
```

---

```{r echo=FALSE}
barchartGC(sitDec,type="percent",
           main="Concert Decision,\nby Situation on Survey Form",
           ylab="Scenario on Survey Form",
           flat=TRUE)
```

## Common Knitting Problem

Make sure that in your R Markdown document you create everyting you need to run a function BEFORE you ask R to do something with it in the document.

This will knit up:

```{r eval=FALSE}
sitDec <- xtabs(~conc.situation+conc.decision,data=attitudes)
barchartGC(sitDec)
```

This will NOT knit:

```{r eval=FALSE}
barchartGC(sitDec)  #asking for a barchart of sitDec before making sitDec
# then later on ...
sitDec <- xtabs(~conc.situation+conc.decision,data=attitudes)
rowPerc(sitDec)
```



# Numerical Results

## Two Way Table and Row Percents


```{r}
sitDec
```


```{r}
rowPerc(sitDec)
```

# Inferential Results

## The R-Code

```{r eval=FALSE}
proptestGC(~conc.situation+conc.decision,data=attitudes,
           success="buy")
```

* gives us a 95%-confidence interval, by default

## Short Form Also OK

```{r eval=FALSE}
proptestGC(~conc.situation+conc.decision,data=attitudes,
           success="buy",verbose=FALSE)
```


## Some Output (Longer Version)

```
Estimate of p1-p2:   0.3081 
SE(p1.hat - p2.hat):	 0.05449 
```

* Our best guess is that $p_1-p_2$ is about 30.81%

* but this estimate could easily be off by 5.5% or so, by chance variation in which subjects were given which form

* still, $\hat{p}_1-\hat{p}_2$ is nearly six standard errors above 0

* our results are very unlikely to occur if the anchors have no effect!

## Confidence Interval (Both Versions)

```
95% Confidence Interval for p1-p2:

          lower.bound         upper.bound          
          0.201340            0.414937        
```

Interpretation:  We are 95%-confident that $p_1-p_2$ is somewhere between 20% and 41.5%.

## Address the Research Question

If anchor has no effect on decision to buy, then $p_1-p_2 = 0$.  But the 95%-interval

$$(0.201340,0.414937)$$

lies entirely above 0.  This study provided VERY strong evidence that if you lose cash you are more likely to attend the concert than if you lose a ticket of the same cash value.

# Discussion

## Safety Check

* You should have taken two independent random samples from two populations, or done a completely randomized experiment.  (We did the latter.)

```{r echo=FALSE}
sitDec
```


* Number of successes and number of failures in both groups should be at least 10.
    * 106 buys, 24 non-buys in lost-money group (OK)
    * 70 buys, 68 non-buys in lost-ticket group (OK)