---
title: "DAR Ch.2 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Methods

## Variable Analysis

You must include this in your methods!

Two variables involved:

* **position** (a factor variable, explanatory)
* **pushups** (a numerical variable, response)

This leads to:

* density plots, or
* histograms, or
* box plots
* and `favstats()` (for numerical summaries)

# Graphical Results

## Density Plots A

Some folks chose density plots, as in class slides:

```{r eval=FALSE}
densityplot(~pushups|position,data=pushups,
            main="Pushups in Two Minutes, by Position",
            xlab="Number of Pushups Done",
            layout=c(1,2))
```

## The Graph


```{r echo=FALSE}
densityplot(~pushups|position,data=pushups,
            main="Pushups in Two Minutes, by Position",
            xlab="Number of Pushups Done",
            layout=c(1,2))
```

## Density Plots B

Some folks "grouped", as taught in website slides and Course Notes:

```{r eval=FALSE}
densityplot(~pushups,data=pushups,
            groups=position,
            main="Pushups in Two Minutes, by Position",
            xlab="Number of Pushups Done",
            auto.key=TRUE)
```

## The Graph

```{r echo=FALSE}
densityplot(~pushups,data=pushups,
            groups=position,
            main="Pushups in Two Minutes, by Position",
            xlab="Number of Pushups Done",
            auto.key=TRUE)
```

## Box Plots A

Some folks followed the procedure in class slides:

```{r eval=FALSE}
bwplot(pushups~position,data=pushups,
       main="Pushups in Two Minutes, by Position",
       xlab="Type of Position",
       ylab="Number of Pushups Done")
```

## The Graph

```{r echo=FALSE}
bwplot(pushups~position,data=pushups,
       main="Pushups in Two Minutes, by Position",
       xlab="Type of Position",
       ylab="Number of Pushups Done")
```

## Box Plots B

Other folks discovered that you could lay the box plots flat:

```{r eval=FALSE}
bwplot(position~pushups,data=pushups,
       main="Pushups in Two Minutes, by Position",
       ylab="Type of Position",
       xlab="Number of Pushups Done")
```

## The Graph

```{r echo=FALSE}
bwplot(position~pushups,data=pushups,
       main="Pushups in Two Minutes, by Position",
       ylab="Type of Position",
       xlab="Number of Pushups Done")
```

## Histograms A

Some folks blindly imitated the code for in the example report:

```{r eval=FALSE}
histogram(~pushups|position,data=pushups,
      main="Pushups in Two Minutes, by Position",
      xlab="Number of Pushups Done",
      layout=c(1,2),width=1)
```

## Not So Good!  (Too "Spikey")

```{r echo=FALSE}
histogram(~pushups|position,data=pushups,
      main="Pushups in Two Minutes, by Position",
      xlab="Number of Pushups Done",
      layout=c(1,2),width=1)
```

## Better Histograms

Exercise good judgement, modify the rectangle widths:

```{r eval=FALSE}
histogram(~pushups|position,data=pushups,
      main="Pushups in Two Minutes, by Position",
      xlab="Number of Pushups Done",
      layout=c(1,2),width=5)
```

## A Better Graph

```{r echo=FALSE}
histogram(~pushups|position,data=pushups,
      main="Pushups in Two Minutes, by Position",
      xlab="Number of Pushups Done",
      layout=c(1,2),width=5)
```

# Numerical Results

## favstats()


```{r}
favstats(pushups~position,data=pushups)
```

In your discussion, focus primarily on:

* means, or
* medians, or even
* how the middle 50% for Skill players lies entirely above the middle 50% for Line players

But not on mins or maxes!


## A Poor Choice

Don't try `xtabs()`:

```{r}
xtabs(~position+pushups,data=pushups)
```

`xtabs()` is for factor variables!

# Conclusion

## Inference to a Population

Many had questions about whether the results in the sample might apply to a larger population:

* all GC football players?
* all college football players?
* all GC football players a few years ago, but maybe not today?

Good questions.

## What About Weight?

The weight of each player was recorded, but **weight** was not involved in the Research Question.

Some folks wondered whether **weight** might be related to **pushups**, maybe even be what *causes* the number of push-ups you can do.

## Studying Weight and Pushups

We could do that with Chapter 4 methods:

```{r eval=FALSE}
xyplot(pushups~weight,data=pushups,
       main="Pushups Done, by Weight",
       xlab="weight of player (lbs)",
       ylab="number of pushups in two minutes",
       pch=19,type=c("p","r"))
```

## The Graph

```{r echo=FALSE}
xyplot(pushups~weight,data=pushups,
       main="Pushups Done, by Weight",
       xlab="weight of player (lbs)",
       ylab="number of pushups in two minutes",
       pch=19,type=c("p","r"))
```

## Grouping

You can even break down by **position**:

```{r eval=FALSE}
xyplot(pushups~weight|position,data=pushups,
       main="Pushups Done, by Weight",
       xlab="weight of player (lbs)",
       ylab="number of pushups in two minutes",
       pch=19,type=c("p","r"),
       layout=c(2,1))
```

## The Graph

```{r echo=FALSE}
xyplot(pushups~weight|position,data=pushups,
       main="Pushups Done, by Weight",
       xlab="weight of player (lbs)",
       ylab="number of pushups in two minutes",
       pch=19,type=c("p","r"),
       layout=c(2,1))
```





