---
title: "DAR Ch.4 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(tigerstats)
```

# Methods

## Education and Wages

Is a worker's education related to his/her wages?

For the question on education and wages:

* **educ** is numerical
* **wage** is numerical

This leads to:

* scatter plot (`xyplot()`)
* a linear model (`lmGC()`)

## Education and Sex

Do men and women differ in how much education they have?

* **sex** is a factor
* **educ** is numerical

This leads to:

* `favstats()`
* `densityplot()` (or bwplots or histograms)

## Experience and Wages

Is a worker's number of years of work experience related to his/her wages?

* **exper** is numerical
* **wage** is numerical

This leads to:

* scatter plot (`xyplot()`)
* a linear model (`lmGC()`)

## Experience and sex

Do men and women differ in how much education they have?

* **sex** is a factor
* **exper** is numerical

This leads to:

* `favstats()`
* `densityplot()` (or bwplots or histograms)

## An Outlier!

Some students noticed an outlier:

```{r eval=FALSE}
bwplot(~wage,data=CPS85,
            main="Wages",
            xlab="wage (dollars per hour)")
```

## The Graph


```{r echo=FALSE}
bwplot(~wage,data=CPS85,
            main="Wages",
            xlab="wage (dollars per hour)")
```

## Who is This?

```{r}
bigBucks <- subset(CPS85,wage > 40)
bigBucks
```

Seems unlikely!  Could there have been a data-entry error?

## Removing the Outlier

Some folks chose to remove the outlier:

```{r}
CPSsmall <- subset(CPS85,wage <= 40)
```

Then they did all of their analysis with `CPSsmall`.

Their results weren't much different than if they had kept the outlier.


# Results

## Is Education a Confounding Factor?

If so, it would be:

* related to **sex**, and
* help to cause **wage**

## Education and Wages

```{r eval=FALSE}
xyplot(wage~educ,data=CPS85,
       xlab="years of education",
       ylab="wage (dollars per hour)",
       main="Education and Wages",
       type=c("p","r"))
```

## The Scatterplot

```{r echo=FALSE}
xyplot(wage~educ,data=CPS85,
       xlab="years of education",
       ylab="wage (dollars per hour)",
       main="Education and Wages",
       type=c("p","r"))
```

More education could cause higher wages!

## Education and Sex

Do guys have more education than gals?

```{r}
favstats(educ~sex,data=CPS85)
```

Apparently not!

## Education and Sex

```{r eval=FALSE}
densityplot(~educ,data=CPS85,
            groups=sex,
            xlab="years of education",
            plot.points=FALSE,
            auto.key=TRUE)
```

## The Graph

```{r echo=FALSE}
densityplot(~educ,data=CPS85,
            groups=sex,
            xlab="years of education",
            plot.points=FALSE,
            auto.key=TRUE)
```

Not much difference!

## Is Experience a Confounding Factor?

If so, it would be:

* related to **sex**, and
* help to cause **wage**

## Experience and Wages

```{r eval=FALSE}
xyplot(wage~exper,data=CPS85,
       xlab="years of experience",
       ylab="wages (dollars perhour)",
       main="Experience and Wages",
       type=c("p","r"))
```

## The Scatterplot

```{r echo=FALSE}
xyplot(wage~exper,data=CPS85,
       xlab="years of experience",
       ylab="wage (dollars per hour)",
       main="Experience and Wages",
       type=c("p","r"))
```

Not clearly a positive relationship, here!

## A Deeper Look

You can break down by sex:


```{r eval=FALSE}
xyplot(wage~exper|sex,data=CPS85,
       xlab="years of experience",
       ylab="wage (dollars per hour)",
       main="Experience and Wages",
       type=c("p","r"))
```

## The Graph

```{r echo=FALSE}
xyplot(wage~exper|sex,data=CPS85,
       xlab="years of experience",
       ylab="wage (dollars per hour)",
       main="Experience and Wages",
       type=c("p","r"))
```

Experience and wages are related for men, but not for women!

## Experience and Sex

Do guys have more experience than gals?

```{r}
favstats(exper~sex,data=CPS85)
```

Apparently not!

## Experience and Sex

```{r eval=FALSE}
densityplot(~exper,data=CPS85,
            groups=sex,
            xlab="years of experience",
            plot.points=FALSE,
            auto.key=TRUE)
```

## The Graph

```{r echo=FALSE}
densityplot(~exper,data=CPS85,
            groups=sex,
            xlab="years of experience",
            plot.points=FALSE,
            auto.key=TRUE)
```

Not much difference!

## Correcting for Education

Let's predict wages for a man and a woman who have the SAME level of education.

First, make separate linear model for each sex:

```{r}
educWageMale <- lmGC(wage~educ,
                     data=subset(CPS85,sex=="M"))
educWageFemale <- lmGC(wage~educ,
                     data=subset(CPS85,sex=="F"))
```

## Six Years of Education

```{r}
predict(educWageMale,x=6)
predict(educWageFemale,x=6)
```

The guy is predicted to make more!

## Fifteen Years of Education


```{r}
predict(educWageMale,x=15)
predict(educWageFemale,x=15)
```

The guy is also predicted to make more!

## Correcting for Experience

Let's predict wages for a man and a woman who have the SAME level of experience.

First, make separate linear model for each sex:

```{r}
experWageMale <- lmGC(wage~exper,
                     data=subset(CPS85,sex=="M"))
experWageFemale <- lmGC(wage~exper,
                     data=subset(CPS85,sex=="F"))
```

## Ten Years of Experience

```{r}
predict(experWageMale,x=10)
predict(experWageFemale,x=10)
```

The guy is predicted to make more!

## Forty Years of Experience


```{r}
predict(experWageMale,x=40)
predict(experWageFemale,x=40)
```

The guy is also predicted to make more!


# Discussion

## The "Education" Theory

The wage difference between the sexes cannot be explained as being due to guys having more education:

* guys and gals had about the same amount of education.

## The Experience Theory

The wage difference between the sexes cannot be explained as being due to guys having more experience:

* guys and gals had about the same amount of experience;
* experience wasn't clearly related to wages, anyway!

## Correcting for Education

```{r echo=FALSE}
par.settings <- list(
  superpose.symbol = list(col = c("blue", "red"), fill = c("blue", "red")),
  superpose.line = list(col =c("blue", "red")))                                                                                                                       
xyplot(wage~educ,data=CPS85,groups=sex,auto.key=list(space="right"),
       type=c("p","r","g"),pch=19,par.settings=par.settings,
       main="Wages vs. Education, by Sex",
       xlab="years of education",
       ylab="wages (dollar per hour")
```

## Correcting for Experience

```{r echo=FALSE}
par.settings <- list(
  superpose.symbol = list(col = c("blue", "red"), fill = c("blue", "red")),
  superpose.line = list(col =c("blue", "red")))                                                                                                                       
xyplot(wage~exper,data=CPS85,groups=sex,auto.key=list(space="right"),
       type=c("p","r","g"),pch=19,par.settings=par.settings,
       main="Wages vs. Expereince, by Sex",
       xlab="years of experience",
       ylab="wages (dollar per hour")
```





