---
title: "DAR Ch.4 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Plagiarism Issues

## One More Reminder:

* You should leave the Introduction untouched (it sets the assignment)
* Use your own words in other sections.
    * Really, really use your own words!
    * Don't just change a word or two from each sentence in the example document!
* Advice:  read the example document in advance, but don't look at it while writing up your own assignment

**Copying is plagiarism!**

# Methods

## Variable Analysis

You must include this in your methods!

Two variables involved:

* **Weight** (a numerical variable)
* **ChugTime** (a numerical variable)

Later on you have to predict a chug-time from a known weight, so best to pick **Weight** as the explanatory variable.

## Methods Chosen

Two numerical variables leads to:

* `xyplot()` for a scatter plot
* `lmGC()` for a linear regression model

# Graphical Descriptive Results

## xyplot()

```{r eval=FALSE}
xyplot(ChugTime~Weight,data=chugtime,
       main="Time to Chug, by Weight",
       xlab="weight (pounds)",
       ylab="chuggin time (sec)",
       pch=19,type=c("p","r"))
```

---

```{r echo=FALSE}
xyplot(ChugTime~Weight,data=chugtime,
       main="Time to Chug, by Weight",
       xlab="weight (pounds)",
       ylab="chuggin time (sec)",
       pch=19,type=c("p","r"))
```

## Address the Research Question

People did this well!


# Numerical Results

## The Linear Model

```{r eval=FALSE}
chugWeight <- lmGC(ChugTime~Weight,data=chugtime)
chugWeight
```


## Correlation

```
Correlation coefficient r =  -0.7879 
```

This is fairly near -1, confirming a moderately strong negative relationship between **Weight** and **ChugTime**.

## Regression Line

```
ChugTime = 13.2975 + -0.0458 * Weight 
```

* Intercept = 13.2975
    * a zero-pound person is predicted to chug in 13.2975 seconds.  This makes no sense!
* Slope = -0.0458
    * For every extra pound of weight, the predicted chug-time decreases by 0.0458 seconds.
    * If two guys differ by 10 pounds in their weight, the heavier guy is predicted to chug in 0.458 fewer seconds.
    * etc. ...
    
## R^2

```
R^2 (unadjusted):  	R^2 = 0.6208
```

* The weights vary
* the chug-times vary
* About 62% of the variation is weights is accounted for by variation in the chug-times

## Prediction

```{r eval=FALSE}
predict(chugWeight,x=170)
```

You get a prediction of 5.513 seconds.

Your prediction could be off by 1.1234 or so (residual standard error s).

**Note**:  Actually, s is a bit low as a standard error for prediction.  Later you may learn "better" standard errors.

## A Problem

Some people made this model:

```{r}
weightChug <- lmGC(Weight~ChugTime,data=chugtime)
weightChug
```

## How to Predict Chugtime?

You can't do this:

```{r}
predict(weightChug,x=170)
```

The `x` is a weight, not a chug-time!



## What They Thought

They looked at their regression equation:

$$Weight = 248.0874 - 13.5563 \times ChugTime$$

They thought:

* "Hmm, maybe I can solve it for **ChugTime**!"

## What They Did

They plugged in 170 for the weight, then solved for ChugTime:

$$170 = 248.0874 - 13.5563 \times ChugTime \\
13.5563 \times ChugTime = 248.0874 - 170 \\
ChugTime = \frac{248.0874 - 170}{13.5563} \\
\approx 5.76$$

This is not 5.513 (the correct prediction).

The regression line for `weightChug` is not the same as the regression line for `chugWeight`!

# Discussion

## Confounding Variable

A possible confounding variable is **cross-sectional area of throat**:

* It's a third variable.
* It is associated with **weight** (bigger guys tend to weigh more and to have bigger throats), but is not *caused* by weight (gaining wight doesn't make your throat wider)
* It helps cause chug-time (when your throat is big, you can pour stuff down it faster!)





   
   

