---
title: "DAR Ch.10 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(abd)
require(tigerstats)
```

# Plagiarism Issues

## Reminders (from Chs 4,9 DAR Feedback):

* You should leave the Introduction untouched (it sets the assignment)
* Use your own words in other sections.
    * Really, really use your own words!
    * Don't just change a word or two from each sentence in the example document!
* Advice:  read the example document in advance, but don't look at it while writing up your own assignment

**Copying is plagiarism!**

# Methods

## Variable Analysis

You must include this in your methods!

One variable involved:

* `percent.clockwise`, a numerical variable.

(It's a number, for each dolphin!)

## Descriptive Methods

One numerical variable leads to:

* `densityplot()` for a  (graphical descriptive)
    - with such a small sample size, a box-plot or histogram won't be very informative
* `favstats()` (numerical descriptive)

## Inferential Methods

* `percent.clockwise` is numerical

* so we could be interested in one population mean.


## Define Parameter, Choose Methods

Define  $\mu$ to be the mean percentage of sleep-time spent swimming clockwise, for ALL Southern Hemisphere dolphins.

* we'll make a 95%-confidence interval for $\mu$, and
* perform a two-sided test to see if the data provide strong evidence that $\mu$ is bigger than 50%.


# Graphical Descriptive Results

## densityplot()



```{r eval=FALSE}
densityplot(~percent.clockwise,data=Dolphins,
            to=100,
            xlab="percent of time swimming counter-clockwise",
            main="Dolphins in the Southern Hemisphere")
```

* For such a small sample, `bwplot()` and `histogram()` might not deliver very informative graphs.  (But I did not count off if you used one of them.)
* `from` and `to` arguments are useful when you know the possible bouinds of the data

---

```{r echo=FALSE}
densityplot(~percent.clockwise,data=Dolphins,
            to=100,
            xlab="percent of time swimming counter-clockwise",
            main="Dolphins in the Southern Hemisphere")
```


# Numerical Results

## favstats()


```{r}
favstats(~percent.clockwise,data=Dolphins)
```



# Inferential Results

## Hypotheses

Do Southern Hemisphere dolphins spend a majority of their sleep-time swimming clockwise?

* $H_0:  \mu = 50$%

* $H_a:  \mu \neq 50$%

## The R-Code

```{r eval=FALSE}
ttestGC(~percent.clockwise,data=Dolphins,mu=50)
```


## Some Output

```
Estimate of mu:   88.21 
SE(x.bar):	 2.861 
```

* Our best guess is that $\mu$ is about 88.21%

* but this estimate could easily be off by 2.861% or so, by chance variation in the sampling process

* still, $\bar{x}$ is MANY standard errors above 0

* our results are very unlikely to occur if Southern Hemisphere dolphins don't spend most of their sleep-time swimming clockwise!

## Confidence Interval

```
95% Confidence Interval for mu:

          lower.bound         upper.bound          
          81.446788           94.978212
```

Interpretation:  We are 95%-confident that $\mu$ is somewhere between 81.4% and 99.8%.

## Address the Research Question

All of the values in the confidence interval were well above 50%.

We can rule out the idea that Southern Hemisphere dolphins spend, on average, less than half of their sleep-time swimming clockwise.

## Test Statistic and P-Value

```
Test Statistic:  	t = 13.36  
P-value:		P = 3.091e-06 
```

* The sample mean was 13.36 standard errors above what $H_0$ expected!

$$3 \times 10^{-6} = \frac{3}{1000000}$$

If Southern Hemisphere dolphins spend only half of their sleep-time swimming clockwise, then there is only a 3 in one million chance of getting a test statistic at least as far from 0 as the one we actually got!

## Decision and Conclusion

Since $p < 0.05$, we'll reject $H_0$.

This study provided very strong evidence that the mean percentage of sleep-time spent swimming clockwise, for all Southern Hemisphere dolphins, was greater than 50%.

# Discussion

## Safety Check

Most important part:  You should have taken a simple random sample from the population!  But ...

* ... nothing said here about how the sample was taken!  (What if the dolphins are all from one herd/family?)

* In fact (Google the title of the article to find it online), the article says:  "The present study reports the results of several nights of observation of eight dolphins held at Seaworld in Durban, South Africa."  Not a random sample at all!!

## More Safety Check

Less Important Part:  If we don't KNOW that the population is normally distributed, then either:

* we must have a large sample, or
* we should have good reason to believe that the population is not far from normal

That's why when the sample is small we look carefully at a graph of the sample. 

---

```{r echo=FALSE}
densityplot(~percent.clockwise,data=Dolphins,
            to=100,
            xlab="percent of time swimming counter-clockwise",
            main="Dolphins in the Southern Hemisphere")
```

But if the sample is very small (e.g, 8) even a symmetric, unimodal graph does not provide much evidence that the population itself is close to normal.  

# Alternative Analysis

## Turn data into yes/no

For each dolphin, record whether or not the dolphin spent more than half of its sleep-time swimming clockwise.

The we could be interested in one proportion $p$, the proportion of all Southern Hemisphere dolphins that spend more than half their time swimming clockwise.

## Hypotheses

* $H_0:  p = 0.50$

* $H_a:  p \neq 0.50$

```{r eval=FALSE}
binomtestGC(x=8,n=8,p=0.50,graph=TRUE)
```

## Some Results

```
Estimate of p:   1 
SE(p.hat):	 0 
```

The standard error doesn't really make sense when there are no failures.  Ignore it.

## More Results

```
95% Confidence Interval for p:

          lower.bound         upper.bound          
          0.630583            1.000000
```

Interval lies entirely above 0.5.  Looks bad for the Null!

## P-Value

```{r echo=FALSE}
pbinomGC(7,region="above",size=8,prob=0.5,graph=TRUE)
```

##  Our Judgements

We would reject $H_0$ and say that the data provided strong evidence that a majority of Southern Hemisphere dolphins spend more than half of their sleep-time swimming clockwise ...

... except that we still think we don't have a random sample of dolphins!

No fancy statistical trick gets us around this problem!

* For practice, we perform the inferential procedure.

* In real life, you would not perform the descriptive procedures, and you would explain to the reader why inferential procedures are not appropriate in this study.

