---
title: "DAR Ch.2 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Methods

## Variable Analysis

You must include this in your methods section! (See example document.)

For each question, two variables are involved, for example:

* **race** (a factor variable, explanatory)
* **wages** (a numerical variable, response)

This leads to:

* density plots, or
* histograms, or
* box plots
* and `favstats()` (for numerical summaries)

# Results

## union and Wage:  Graphs

Some folks chose density plots, as in class slides:

```{r eval=FALSE}
densityplot(~wage|union,data=CPS85,
            main="Hourly Wage, by Union Status",
            xlab="wage (dollars per hour)",
            layout=c(1,2))
```

## The Graph


```{r echo=FALSE}
densityplot(~wage|union,data=CPS85,
            main="Hourly Wage, by Union Status",
            xlab="wage (dollars per hour)",
            layout=c(1,2))
```

## Density Plots B

Some folks "grouped", as taught in website slides and Course Notes:

```{r eval=FALSE}
densityplot(~wage,data=CPS85,
            groups=union,
            main="Hourly Wage, by Union Status",
            xlab="wage (dollars per hour)",
            auto.key=TRUE)
```

## The Graph

```{r echo=FALSE}
densityplot(~wage,data=CPS85,
            groups=union,
            main="Hourly Wage, by Union Status",
            xlab="wage (dollars per hour)",
            auto.key=TRUE)
```



## Histograms

You've been leanring about histograms in Course Notes and Online Quizzes.  Some folks used them:


```{r eval=FALSE}
histogram(~wage|union,data=CPS85,
      main="Hourly Wage, by Union Status",
      xlab="wage (dollars per hour)",
      layout=c(1,2), width=2
      )
```

## The Graph

```{r echo=FALSE}
histogram(~wage|union,data=CPS85,
      main="Hourly Wage, by Union Status",
      xlab="wage (dollars per hour)",
      layout=c(1,2), width=2,
      )
```

## Union and Wage:  Numerical

```{r}
favstats(wage~union,data=CPS85)
```

Unionized workers earn more on average (10.80/hr, as compared to 8.64/hr for non-unionized workers).


## (Option) Sex and Wage

```{r}
favstats(wage~sex,data=CPS85)
```

## (Another Option) Sector and Wage

```{r}
favstats(wage~sector,data=CPS85)
```

## Graphs for Sector and Wage

```{r echo=FALSE}
bwplot(wage~sector,data=CPS85,
       main="Hourly Wage, by Sector",
       ylab="wage (dollars per hour)",
       xlab="sector of employment"
       )
```

But when you knit the labels may get jammed together!

## Solution 1

You could use histograms or density plot, e.g.:

```{r eval=FALSE}
densityplot(~wage|sector,data=CPS85,
            main="Hourly Wage, by Sector",
            xlab="wage (dollars per hour)",
            layout=c(1,8))
```

## Graph

```{r echo=FALSE}
densityplot(~wage|sector,data=CPS85,
            main="Hourly Wage, by Sector",
            xlab="wage (dollars per hour)",
            layout=c(1,8))
```

## Solution 2 (nice trick)

You can flip the bwplot:

```{r eval=FALSE}
bwplot(sector~wage,data=CPS85,
       main="Hourly Wage, by Sector",
       xlab="wage (dollars per hour)",
       ylab="sector of employment"
       )
```

## The Graph

```{r echo=FALSE}
bwplot(sector~wage,data=CPS85,
       main="Hourly Wage, by Sector",
       xlab="wage (dollars per hour)",
       ylab="sector of employment"
       )
```

# Discussion/Conclusion

------

The strongest relationships were:

* between wages and union status (mean 10.80 vs. 8.44)
* between wages and sex (mean 9.99 vs. 7.88)

Some differences in wages among sectors were also impressive!





