---
title: "DAR Ch.10 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(tigerstats)
require(tigerData)
```


# Methods

## Departure and Being Honked At

Research Question:  Do people takes longer to vacate a parking spot if a waiting car honks at them?

Variable Analysis:

* **horn** (explanatory, factor, two values)
* **time** (response, numerical)

Notice:  this was a randomized experiment, because researchers chose which subjects to honk at.

## Departure and Being Honked At

Research Question:  Who takes longer to vacate a parking spot:  a male or a female?

Variable Analysis:

* **sex** (explanatory, factor, two values)
* **time** (response, numerical)

Notice:  this was an observational study, because researchers did not assign subjects their sex!

## Methods

For these two Research Questions:

* graphical:  `densityplot()`
* numerical:  `favstats()`
* inferential:  `ttestGC()`

## Departure and Car Value

Research Question:  Do drivers of expensive cars take longer to vacate their spot?

Variable Analysis:

* **carval** (explanatory, numerical)
* **time** (response, numerical)

Notice:  this was an observational study, because researchers just measured the value of the car the subject was driving.

## Methods

For this Research Question:

* graphical:  `xyplot()`
* numerical:  `lmGC()`

# Results

## Horn and Time

```{r eval=FALSE}
densityplot(~time|horn,data=parkExp,
            main="Departure Time Based on\nBeing Honked at",
            xlab="departure time (seconds)",
            layout=c(1,2))
```

## The Graph

```{r echo=FALSE}
densityplot(~time|horn,data=parkExp,
            main="Departure Time Based on\nBeing Honked at",
            xlab="departure time (seconds)",
            layout=c(1,2))
```

## Numerical

```{r}
favstats(time~horn,data=parkExp)
```

## Inferential:  Parameters

Define:

$\mu_1 =$ mean time to depart for all shoppers, if all could be honked at by the waiting car.


$\mu_2 =$ mean time to depart for all shoppers, if all could be NOT honked at by the waiting car.


## Inferential:  Safety Check

This was a randomized experiment.

Both samples were right-skewed but the samples sizes (79 honked at, 77 not) were pretty large.

So we are probably safe.

## Inferential:  Code

```{r eval=FALSE}
ttestGC(time~horn,data=parkExp,mu=0,
        first="yes_horn",graph=TRUE)
```

## Inferential:  Discussion


```
Estimate of mu1-mu2:   12.27 
SE(x1.bar - x2.bar):	 2.753 
```

$\mu_1-\mu_2$ is about 12.27 seconds, give or take about 2.753 seconds or so for chance variation in the assignment of subjects to their groups.


## Inferential:  Interval

```
95% Confidence Interval for mu1-mu2:

          lower.bound         upper.bound          
          6.834534            17.712180            
```

We are 95%-confident that $\mu_1-\mu_2$ is between 6.8 and 17.7 seconds.  This interval lies entirely above 0, so the data provides strong evidence that people who are honked at take longer to depart, on average.

## Inferential:  Test of Significance

Here are the hypotheses:

```
H_0:  mu1-mu2 = 0 
H_a:  mu1-mu2 != 0 
```

## Inferential:  Test Statistic and P-Value

```
Test Statistic:  	t = 4.458  
P-value:		P = 1.596e-05 
```

Our results are 4.458 standard errors above what the Null expected.  Looks bad for the Null!

There is about a 1.6 in 100,000 chance of getting a test statistic at least as far from 0 as the one we got, if $H_0$ is true!

## Inferential:  Wrap It Up

The P-value was tiny, so we will reject $H_0$.

This data provided very strong evidence that people take longer to depart, on average, when they are honked at.



## Sex and Time

```{r eval=FALSE}
densityplot(~time|sex,data=parkExp,
            main="Departure Time Based on\nSex of the Driver",
            xlab="departure time (seconds)",
            layout=c(1,2))
```

## The Graph

```{r echo=FALSE}
densityplot(~time|sex,data=parkExp,
            main="Departure Time Based on\nSex of the Driver",
            xlab="departure time (seconds)",
            layout=c(1,2))
```

## Numerical

```{r}
favstats(time~sex,data=parkExp)
```

## Inferential:  Parameters

Define:

$\mu_1 =$ mean time to depart for all female shoppers.


$\mu_2 =$ mean time to depart for all male shoppers.


## Inferential:  Safety Check

This was an observational study.  Hopefully the subjects were like a random sample from the population of all shoppers.

Both samples were right-skewed but the samples sizes (84 female, 72 male) were pretty large.

So we are probably safe.

## Inferential:  Code

```{r eval=FALSE}
ttestGC(time~sex,data=parkExp,mu=0,
        first="female",graph=TRUE)
```

## Inferential:  Discussion


```
Estimate of mu1-mu2:   4.143 
SE(x1.bar - x2.bar):	 2.941 
```

$\mu_1-\mu_2$ is about 4143 seconds, give or take about 2.941 seconds or so for chance variation in who got into the study.


## Inferential:  Interval

```
95% Confidence Interval for mu1-mu2:

          lower.bound         upper.bound          
          -1.668900           9.954773         
```

We are 95%-confident that $\mu_1-\mu_2$ is between -1.67 and 9.95 seconds.  This contains 0, so the data does NOT provide strong evidence that departure time is, on average, related to sex of the driver.

## Inferential:  Test of Significance

Here are the hypotheses:

```
H_0:  mu1-mu2 = 0 
H_a:  mu1-mu2 != 0 
```

## Inferential:  Test Statistic and P-Value

```
Test Statistic:  	t = 1.409 
P-value:		P = 0.161 
```

Our results are 1.4 standard errors above what the Null expected.  This is not so far off from what the Null expects!

There is about a 16% chance of getting a test statistic at least as far from 0 as the one we got, if $H_0$ is true!

## Inferential:  Wrap It Up

The P-value was $> 0.05$, so we will NOT reject $H_0$.

This data did NOT provide strong evidence that males and females differ, on average, in how long they take to depart.

## Car Value and Time to Depart

```{r eval=FALSE}
xyplot(time~carval,data=parkExp,
       main="Departure Tiem and Car Value",
       xlab="Blue Book value of driver's car (dollars)",
       ylab="time to depart (seconds)",
       type=c("p","r"))
```

## The Graph

```{r echo=FALSE}
xyplot(time~carval,data=parkExp,
       main="Departure Tiem and Car Value",
       xlab="Blue Book value of driver's car (dollars)",
       ylab="time to depart (seconds)",
       type=c("p","r"))
```

## Linear Model

```{r eval=FALSE}
lmGC(time~carval,data=parkExp)
```


## Output

```
Correlation coefficient r =  0.02252 

Equation of Regression Line:

   time = 36.3682 + 1e-04 * carval 

Residual Standard Error:	s   = 18.2914 
R^2 (unadjusted):		R^2 = 5e-04 
```

Correlation very small.  Essentially no relationship, here.

# Discussion

## Interaction between honking and sex?

Interesting new question:  Do men and women differ in how they respond to being honked at?

```{r eval=FALSE}
densityplot(~time|sex*horn,data=parkExp)
```

## The Graph

```{r echo=FALSE}
densityplot(~time|sex*horn,data=parkExp)
```

## Numerical Approach

```{r}
favstats(time~sex+horn,data=parkExp)[c(".group","mean")]
```

Guys and gals both take longer when being honked at, but the difference is about the same.

## Time and Confronting Car Status

Some folks mistook the final Research Question to be this:

Is time to departure related to the status of the car that is waiting?

## Quick Numerical Look


```{r}
favstats(time~ccstatus,data=parkExp)[c(".group","mean")]
```

Does not look like much of a difference, here.

## Interaction between ccstatus and sex?

But what if we brak down by ccstatus AND by sex?

```{r}
favstats(time~sex+ccstatus,data=parkExp)[c(".group","mean")]
```


Wow, big difference here between how men and women react to the different types of waiting car!

## Graphical Look

```{r eval=FALSE}
bwplot(time~ccstatus|sex,data=parkExp)
```

## The Graph

```{r echo=FALSE}
bwplot(time~ccstatus|sex,data=parkExp)
```




