---
title: "DAR Ch.3 Feedback"
author: "Homer White"
date: ""
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(mosaicData)
require(tigerstats)
```

# Plagiarism Issues

## Bear in Mind:

Last week when going over DARCh2 I warned you not to copy the words in the example document; you should use your own words instead.

You should leave the Intro untouched (see video for that advice), but you must use your own words in other sections.

This time around, I deducted !0% for plagiarism.  **Next time the deduction will be much, much greater!**

# Methods

## Variable Analysis

You must include this in your methods!

Two variables involved:

* **biggerfoot** (a factor variable)
* **domhand** (a numerical variable)

You could take either variable as the explanatory one.  Let's pick **biggerfoot**.

Two factor variables leads to:

* `barchartGC()` for graphs
* `xtabs()`, `rowPerc()` for numerical summaries
*  `chisqtestGC()` for inference

# Numerical Descriptive Results

## Tables A

```{r}
footHand <- xtabs(~biggerfoot+domhand,data=KidsFeet)
footHand
```

Then row percents:

```{r}
rowPerc(footHand)
```

## Follow Directions!!

From the directions:

"Use your numerical results to describe the pattern (if any) that you find in your data.  Your discussion should focus on using the data to address the Research Question."

## Address the Research Question

```{r echo=FALSE}
rowPerc(footHand)
```

One way to address the Research Question:

"The kids whose right foot was bigger were about four times more likely to be left-handed than were the kids whose left foot was bigger (35.29% vs. 9.09%).""


## Now With Variable Names

In last week's `tigerstats`, you had to tell the reader:

* rows say which foot is bigger
* columns say which hand is dominant

Now you've got the labels.

## Tables B

Some people learned about `mosaic`'s `tally()` function:

```{r}
tally(domhand~biggerfoot,data=KidsFeet,format="count")
```


```{r}
tally(domhand~biggerfoot,data=KidsFeet,format="percent")
```

Not bad, once you get used to the different way the formula works.

# Graphical Descriptive Results

## Barchart A

Some folks chose to make bar charts as in the in-class slides:

```{r eval=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger")
```

## The Graph

```{r echo=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger")
```

Then you must explain things to the reader!

## Barchart B

Others used the new arguments for `barchartGC()`:

```{r eval=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!")
```

## The Graph


```{r echo=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!")
```

## Barchart C

Percentage bar charts are fine, too:

```{r eval=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```

## The Graph


```{r echo=FALSE}
barchartGC(~biggerfoot+domhand,data=KidsFeet,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```


## Barchart D

Some people realized you could put a two-way table into `barchartGC()`:

```{r eval=FALSE}
footHand <- xtabs(~biggerfoot+domhand,data=KidsFeet)
barchartGC(footHand,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```

## The Graph

```{r echo=FALSE}
footHand <- xtabs(~biggerfoot+domhand,data=KidsFeet)
barchartGC(footHand,
           main="Dominant Hand, by Which Foot is Bigger",
           xlab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```


## Barchart E

Some people made a "flat" bar chart:

```{r eval=FALSE}
barchartGC(footHand,
           flat=TRUE,
           main="Dominant Hand, by Which Foot is Bigger",
           ylab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```

## The Graph

```{r echo=FALSE}
barchartGC(footHand,
           flat=TRUE,
           main="Dominant Hand, by Which Foot is Bigger",
           ylab="Which foot is bigger",
           sub="Bar colors indicate which hand is dominant!",
           type="percent")
```

This looks just like a table of row percents!

## Whatever You Do ...

Address the Research Question when you discuss the graph!

From the directions:

"Use your graphical results to describe the pattern (if any) that you find in your data.  Your discussion should focus on using the graph to address the Research Question."


# Inference

## What Does a Test of Significance Do?

* You see a pattern in your data.
* But the data are a product of chance (in the random sampling process, for instance)

So you wonder:

* Does your data provide strong evidence for a patter in the population?
* Or is it reasonable to believe that the pattern in your data was due just to chance?

The chi-square test helps you answer this question!

## State Your Hypotheses!

$H_0: $ In the population of all fourth grade kids, there is no relationship between which foot is bigger and which hand is dominant.


$H_a: $ In the population of all fourth grade kids, there IS a relationship between which foot is bigger and which hand is dominant.


## Chi-Square Test Code

Some people tried the regular chi-square test:

```{r eval=FALSE}
chisqtestGC(footHand)
```

Problem!  Expected cell counts are low:

```
Counts Expected by Null:
          domhand
biggerfoot    L     R
         L 4.51 17.49
         R 3.49 13.51

Warning message:
In chisqtestGC(footHand) : Some expected cell counts are low:
  the approximation of the P-value may be unreliable.
	Consider using simulation.
```

## Simulation A

You can always just set `simulate.p.value` to `TRUE`:

```{r eval=FALSE}
set.seed(3030)
chisqtestGC(footHand,
            simulate.p.value=TRUE,
            B=10000)
```

Don't forget to set your seed!

## Simulation B

Since the kids are a random sample from the population of all kids, we could instead use:

```{r eval=FALSE}
set.seed(3030)
chisqtestGC(~biggerfoot+domhand,data=KidsFeet,
            simulate.p.value="random",
            B=10000,
            graph=TRUE)
```

## Discuss Results!

From the directions:

"State hypotheses clearly.  Make sure that your discussion of the inferential test addresses all of the important features of the output such as test statistic, P-value, etc.  Make a decision about the Null Hypothesis.  Be sure to choose a routine that produces a P-value approximation on which you can rely."

## Outline of a Good Discussion

* The chi-square statistic was 4.0383.
* If the Null is correct, it should be about 1, give or take some for chance error in process of sampling kids for the study.
* The P-value is about 0.0434. 
* This means that if the Null is correct, then there was only a 4.34% chance of getting a chi-square statistic at least as big as the one we got.
* The P-value is less than 0.05, so we'll reject the Null.
* This study provided strong evidence (barely) for a relationship in the population between the two variables.

Note that all "steps" in a test of significance are addressed, including the conclusion.

## Interesting ...

P-value a bit lower with `simulate.p.value="random"` than with `simulate.p.value=TRUE`.

If you used `simulate.p.value=TRUE`, then you probably got a P-value a bit over 0.05, so you would not have rejected $H_0$.  Then your conclusion would have been a bit different:

"This study did not quite provide strong evidence for a relationship in the population between the two variables. Maybe we should do the study again, with a bigger sample.""

