---
title: "Describing Patterns 2"
author: "Homer White"
date: "Sunday, August 17, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```


# Median and Quartiles

## Justin Verlander:  Pitcher

```{r eval=FALSE}
data(verlander)
View(verlander)
help(verlander)
```

## favstats() Again

```{r}
favstats(~speed,data=verlander)
```

* Q1:  first Quartile.  About 25% are below this value.

* median:  50th percentile.  About 50% are below this value.

* Q3:  third Quartile.  about 75% are below this value.

* IQR = Q3 - Q1 = 96.4 - 85.3 = 11.1.  Middle 50% are spread this far apart.

## Boxplot for Speeds

```{r eval=FALSE}
bwplot(~speed,data=verlander)
```


## The Boxplot

```{r echo=FALSE}
bwplot(~speed,data=verlander)
```

The 58.9 mph pitch was an outlier!

## Practice 1 { .practice }

```{r eval=FALSE}
data(m111survey)
View(m111survey)
help(m111survey)
```

*Find the median and IQR of the fastest speeds driven.*

# Shapes of Numerical Distributions

## Unimodal, Left-Skewed


```{r echo=FALSE}
set.seed(12345)
dunileft <- rbeta(1000,shape1=10,shape2=1)
plot(density(dunileft),axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,at=F,labels=F)
axis(2,at=F,labels=F)
```

## Unimodal, Right-Skewed

```{r echo=FALSE}
set.seed(12345)
duniright <- rbeta(1000,shape1=1,shape2=10)
plot(density(duniright),axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,labels=F,at=F)
axis(2,labels=F,at=F)
```

## Bimodal, Right-Skewed


```{r echo=F}
set.seed(12345)
dbiright <- c(duniright,rnorm(400,mean=0.5,sd=0.1))
plot(density(dbiright),axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,labels=F,at=F)
axis(2,labels=F,at=F)
```

## Unimodal, Symmetric


```{r echo=FALSE}
set.seed(12345)
dunisymm <- rnorm(1000)
plot(density(dunisymm),axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,labels=F,at=F)
axis(2,labels=F,at=F)
```

This is often called "bell-shaped."

## Bimodal, Symmetric


```{r echo=FALSE}
set.seed(12345)
dbisymm <- rbeta(1000,shape1=0.25,shape2=0.25)
plot(density(dbisymm),
     axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,labels=F,at=F)
axis(2,labels=F,at=F)
```

## Bimodal, Symmetric


```{r echo=F}
set.seed(12345)
dbisymm2 <- rbeta(1000,shape1=0.25,shape2=0.25)
plot(density(dbisymm2,from=0,to=1),
     axes=F,xlab="",ylab="",main="",frame.plot=T,lwd=3,col="blue")
axis(1,labels=F,at=F)
axis(2,labels=F,at=F)
```

## Practice 2 { .practice }

```{r echo=FALSE}
densityplot(~speed,data=verlander,
            main="Verlander Pitches",
            xlab="speed (mph)",
            plot.points=F,
            subset=speed > 60)
```

Describe the shape:  symmetric or skewed?  Unimodal, bimodal, or trimodal?


# Mean and Standard Deviation

## Verlander Again

```{r}
favstats(~speed,data=verlander)
```

* Mean:  sum of values, divided by how many

* SD:  standard deviation.  Typical values are within a SD of the mean.

## 68-95 Rule

When the distribution is roughly bell-shaped:

* About 68% of the values are within one SD of the mean

* About 95% are within two SDs of the mean

* About 99.7% are within 3 SDs of the mean.

*Click the curve on the GC Stats website!*

## Verlander Four-Seam Fastballs

```{r}
verlanderFF <- subset(verlander,pitch_type=="FF")
favstats(~speed,data=verlanderFF)
```

The mean speed is about 96 mph.  The SD is about 2 mph.

## Verlander Fastballs

```{r echo=FALSE}
densityplot(~speed,data=verlanderFF,
            xlab="speed (mph)",
            plot.points=F)
```

Distribution is roughly bell-shaped.

## Practice 3 { .practice }

* About 68% of Justin Verlander's fastballs were between ? and ? (Find the numbers to fill in the ?'s.)

* About what percentage of the time did Verlander throw faster than 100 mph?

* Find a speed so that about 16% of his fastballs were slower than that speed.

# z-scores

## The z-score

$$ z = \frac{\textbf{value}-\textbf{mean}}{SD}.$$

This tells you how many SDs above or below the mean any given value is.

* Large positive z ($z > 2$) means the value was unusually large.
* Large negative z ($z < -2$) means the value was unusually small.



## Example z-score

Mean speed of Verlander's fastball = 96 mph.

SD of speeds is 2 mph.

He throws a 99 mph pitch.  *What is its z-score?*

$$ z= \frac{99-96}{2}=1.5.$$


## Practice 4 { .practice }

* Find the z-score of a 91 mph fastball.
* Was this fastball unusually slow?
* Find the z-score of a 95 mph fastball.  Was this pitch unusually slow?
