---
title: "Describing Patterns 1"
author: "Homer White"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```


# One Factor Variable

## Justin Verlander:  Pitcher

```{r eval=FALSE}
data(verlander)
View(verlander)
help(verlander)
```

## Research Question

>*What types of pitches does Justin Verlander throw most often?*

## Variable Analysis

We are interested in the variable **pitch_type**.

It's a factor variable.

## Numerical Method:  xtabs()

```{r}
xtabs(~pitch_type,data=verlander)
```

## Numerical Method:  rowPerc()

Useful to have a table of percentages, not just counts.

```{r}
rowPerc(xtabs(~pitch_type,data=verlander))
```

## Another Way

Store table in an object, then can print or re-use the object:

```{r}
verPitch <- xtabs(~pitch_type,data=verlander)
verPitch
rowPerc(verPitch)
```


## Graphical Method:  barchartGC()

```{r eval=FALSE}
barchartGC(~pitch_type,data=verlander,
           main="Justin Verlander's Pitches")
```

## The Result

```{r echo=FALSE}
barchartGC(~pitch_type,data=verlander,
           main="Justin Verlander's Pitches")
```


## Percentage Bar Charts:

```{r eval=FALSE}
barchartGC(~pitch_type,data=verlander,
           main="Justin Verlander's Pitches",
           type="percent")
```



## The Result

```{r echo=FALSE}
barchartGC(~pitch_type,data=verlander,
           main="Justin Verlander's Pitches",
           type="percent")
```

He likes to throw fastballs!

# Two Factor Variables

## Research Question

>*Does the type of pitch Verlander throws depend on the stance of the batter?*

## Variable Analysis

* **batter_hand**:  factor variable (left, right)

* **pitch_type**:  factor variable

Compare groups:

* pitches thrown to left-hand batters
* pitches thrown to right-hand batters

## Numerical Methods

We'll use `xtabs()` to make a two-way table, then look at row percentages:

```{r}
batPitch <- xtabs(~batter_hand+pitch_type,data=verlander)
batPitch
rowPerc(batPitch)
```


## Graphical Method:  Barcharts Again


```{r eval=FALSE}
barchartGC(~batter_hand+pitch_type,data=verlander,
           main="Verlander's Pitches, by Batter Stance",
           type="percent")
```

## The Result

```{r echo=FALSE}
barchartGC(~batter_hand+pitch_type,data=verlander,
           main="Verlander's Pitches, by Batter Stance",
           type="percent")
```

# One Numerical Variable


## Math 111 Survey Data

```{r eval=FALSE}
data(m111survey)
View(m111survey)
help(m111survey)
```

## Research Question

>*What is the typical fastest speed driven by GC students?*

Variable Analysis:

**fastest** is a numerical variable.

No `xtabs()`, no `barchartGC()`!

## Graphical Method:  Density Plot

```{r eval=FALSE}
densityplot(~fastest,data=m111survey,
            main="Fastest Speeds at GC",
            xlab="Fastest Speed Driven (mph)")
```

## The Result


```{r echo=FALSE}
densityplot(~fastest,data=m111survey,
            main="Fastest Speeds at GC",
            xlab="Fastest Speed Driven (mph)")
```


## Numerical Method:  favstats()

```{r}
favstats(~fastest,data=m111survey)
```


The typical student drives about 105.9 mph, give or take 20.9 mph or so.


# One Factor and One Numerical Variable

## Research Question

>*Who drives faster, on average:  a GC guy or a GC gal?*

Variable analysis:

* **sex** is a factor variable (male, female)
* **fastest** is numerical

Compare groups:

* speeds of the gals
* speeds of the guys

## Numerical Method:  favstats()

```{r}
favstats(fastest~sex,data=m111survey)
```

## Graphical Method:  bwplot()

```{r eval=FALSE}
bwplot(fastest~sex,data=m111survey,
       main="Fastest Speeds, by Sex",
       xlab="Sex",
       ylab="Fastest Speed (mph)")
```


## The Results

```{r echo=FALSE}
bwplot(fastest~sex,data=m111survey,
       main="Fastest Speeds, by Sex",
       xlab="Sex",
       ylab="Fastest Speed (mph)")
```

## Graphical Method:  Density Plots


```{r eval=FALSE}
densityplot(~fastest|sex,data=m111survey,
       main="Fastest Speeds, by Sex",
       xlab="Fastest Speed (mph)",
       layout=c(1,2))
```

## The Result


```{r echo=FALSE}
densityplot(~fastest|sex,data=m111survey,
       main="Fastest Speeds, by Sex",
       xlab="Fastest Speed (mph)",
       layout=c(1,2))
```

## Practice 1 { .practice }

>**Research Question:**  *How does the speed of a Justin Verlander pitch depend on the type of pitch that it is?*

Variable analysis:

* **pitch_type**: factor or numerical?
* **speed**:  factor or numerical?

What numerical method will you use?

What graphical method?

## Practice 2 { .practice }

>**Research Question:**  *Who is more likely to believe in love at first sight:  a GC guy or a GC gal?*

Variable analysis:

* **?**: factor or numerical?
* **??**:  factor or numerical?

What numerical method(s) will you use?

What graphical method?
