---
title: "Goodness of Fit: Pt.2"
author: "Homer White"
date: "November, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---


```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
require(abd)
```

# More Wild, Wild West

## Further Discussion

What's an "even-handed" way to measure the difference between:

* the table we observed

      vs.

* the table one expects if the Die is fair?

## Chi-Square Statistic

We could try:

$$\chi^2 = \sum_{\textbf{cells}} \frac{(\textbf{observed}-\textbf{expected})^2}{\textbf{expected}}.$$

(We used this back in Chapter 3.)

## Compute It

$$\chi^2 = \frac{(8-10)^2}{10}+\frac{(18-10)^2}{10}+\frac{(11-10)^2}{10} \\
+\frac{(7-10)^2}{10}+\frac{(9-10)^2}{10}+\frac{(7-10)^2}{10} \\
= 8.8$$

## What Are the Chances

Is this a "big" number?

If the die is really fair, what are the chances of getting a $\chi^2$-statistic at least as big as the one we actually got?

## Physical Simulation

Pair up:

* one person rolls fair die 60 times
* the other tallies the results
* compute the chi-square statistic with calculator or R

Were your results $\geq 8.8$?  What about the other students?

## More Simulations

Let's get more simulations with an app:

[Slow Chi-Square Simulation, Shinyapps Server](https://homer.shinyapps.io/SlowGoodness/)

or

[Slow Chi-Square Simulation, GC Server](http://rstudio.georgetowncollege.edu:3838/SlowGoodness/)

## Practice { .practice }

How many times simulations did you do?

What percentage of the time did you get above 8.8?

Do you think the data provide overwhelming evidence that the die was weighted?

## Another Situation

Define

* $p_1 =$ the proportion of all GC students who to sit in the front

* $p_2 =$ the proportion of all GC students who to sit in the middle

* $p_3 =$ the proportion of all GC students who to sit in the back

## In Many Gc Classrooms ...

* there is one "front" row
* there are two "middle" rows
* there is one "back" row

Do the preferences of GC students reflect the space available in each area?

## Hypotheses

* $H_0: p_1=0.25,p_2=0.5,p_3=0.25$

(This means that student preferences reflect available seats.)

* $H_a:$  at least one of the above proportions is wrong

## Our Data

```{r}
xtabs(~seat,data=m111survey)
```

## Practice { .practice }

* Go back to your app.
* Enter the Null Probabilities: 0.25, 0.5, 0.25
* Enter the observed Counts: 27,32,12
* Enter three good names for the values

1. What is the value of the chi-square statistic?
1. Perform some simulations.  How many did  you perform?
1. What percentage of the time did you get at least as big as the actual chi-square statistic?
1. Does the data provide strong evidence that people prefer one part of the classroom more than its number of seats would suggest?


