---
title: "Goodness of Fit: Pt.3"
author: "Homer White"
date: "November, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---


```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
require(abd)
```

# Chi-Square Test for Goodness of Fit


## chisqtestGC()

We can do the inferential work more quickly with `chisqtestGC()`.  For the Gambler Study:

```{r echo=FALSE}
set.seed(2020)
```


```{r eval=FALSE}
obs <- c(one=8,two=18,three=11,four=7,five=9,six=7)
nullProbs <- rep(1/6,times=6)
chisqtestGC(obs,p=nullProbs,
            simulate.p.value=TRUE, B=3000,
            graph=TRUE)
```

## The Output

```
Pearson's chi-squared test with simulated p-value 
   (based on 3000 resamples) 

      Observed counts Expected by Null Contr to chisq stat
one                 8               10                 0.4
two                18               10                 6.4
three              11               10                 0.1
four                7               10                 0.9
five                9               10                 0.1
six                 7               10                 0.9

Chi-Square Statistic = 8.8 
Degrees of Freedom of the table = 5 
P-Value = 0.123 

```
## Graph of Resamples

```{r echo=FALSE,results='hide'}
obs <- c(one=8,two=18,three=11,four=7,five=9,six=7)
nullProbs <- rep(1/6,times=6)
chisqtestGC(obs,p=nullProbs,
            simulate.p.value=TRUE, B=3000,
            graph=TRUE)
```

## The Seating Study

```{r echo=FALSE}
set.seed(3030)
```


```{r eval=FALSE}
chisqtestGC(~seat,data=m111survey,
            p=c(0.25,0.5,0.25),
            simulate.p.value=TRUE,B=3000,
            graph=TRUE)
```

## The Results

```
Pearson's chi-squared test with simulated p-value 
   (based on 3000 resamples) 

         Observed counts Expected by Null Contr to chisq stat
1_front               27            17.75                4.82
2_middle              32            35.50                0.35
3_back                12            17.75                1.86


Chi-Square Statistic = 7.0282 
Degrees of Freedom of the table = 2 
P-Value = 0.03 
```

# The Chi-Square Curve "Approximation"

## Facts About the Chi-Square Statistic

If you have a large enough sample and $H_0$ is true, then

* the expected value is about the degrees of freedom $df$

* the degrees of freedom is the number of cells - 1

* the standard deviation is about $\sqrt{2 \times df}$

## Really?

Try the seating study again:

```{r echo=FALSE}
set.seed(3030)
```


```{r}
seatResults <- chisqtestGC(~seat,data=m111survey,
            p=c(0.25,0.5,0.25),
            simulate.p.value=TRUE,B=3000,
            graph=TRUE)
```

To see the results:

```{r eval=FALSE}
seatResults
```

## Investigate the Simulations

```{r}
favstats(~sims,data=seatResults)[c("mean","sd","n")]
```

Sure enough:

* the mean of the simulations was about 2, the $df$
* the SD of the sims was about 2 (the  $\sqrt{2 \times df}$ figure )

## Thinking with the Mean and the SD


So for the seating study, if the Null is right then we would expect the chi-square statistic to be about 2, give or take

$$\sqrt{2 \times 2} = 2$$

or so.  It came out to 7.0282 (quite a few SDs above the expected value), so chances of getting this or more were small.

## How Big is "Big Enough"

If the expected counts in each cell is at least five, then our sample is DEFINITELY "big enough."

In fact, in that case the chi-square statistic behaves much like a "chi square" density curve with degrees of freedom equal to:

* number of cells -1

## Comparison:  Seating

Let's compare our chi-square statistic simulations with a chi-square curve that has 2 degrees of freedom:

```{r eval=FALSE}
densityplot(~sims,data=seatResults,plot.points=F,from=0,xlim=c(0,10),
            main="Simulations vs. Chi-Square Curve, df=2",
            xlab="Simulations of the chi-square statistic",
            sub="Simulations in blue, Chi-Square Curve in black",
            panel=function(x,...) {
              panel.mathdensity(dmath=dchisq,args=list(df=2),
                  lwd=5,col="black",lty=1,alpha=0.5,...)
              panel.densityplot(x,col="blue",lwd=2,...)
            })
```


## The Graph

```{r echo=FALSE}
densityplot(~sims,data=seatResults,plot.points=F,from=0,xlim=c(0,10),
            main="Simulations vs. Chi-Square Curve, df=2",
            xlab="Simulations of the chi-square statistic",
            sub="Simulations in blue, Chi-Square Curve in black",
            panel=function(x,...) {
              panel.mathdensity(dmath=dchisq,args=list(df=2),
                  lwd=5,col="black",lty=1,alpha=0.5,...)
              panel.densityplot(x,col="blue",lwd=2,...)
            })
```

## Comparison:  Gambler's Die

In the study of the gambler's die:

* there were six cells (one for each side of the die)

* so $df = 6-1=5$

* So if the die is far, then we expect the chi-square statistic to be about 5 ...

* ... give or take $\sqrt{2 \times 5} \approx 3.16$ or so, for chance variation in the rolling of the die

## What do we get?

```{r echo=FALSE}
set.seed(4040)
```


```{r}
obs <- c(one=8,two=18,three=11,four=7,five=9,six=7)
nullProbs <- rep(1/6,times=6)
dieResults <- chisqtestGC(obs,p=nullProbs,
            simulate.p.value=TRUE, B=3000,
            graph=TRUE)
favstats(~sims,data=dieResults)[c("mean","sd","n")]
```

Pretty close!

## Compare With Chi-Square Curve

Let's compare our simulations to a chi-square curve with 5 degrees of freedom:


```{r eval=FALSE}
densityplot(~sims,data=dieResults,plot.points=F,from=0,xlim=c(0,15),
            main="Simulations vs. Chi-Square Curve, df=5",
            xlab="Simulations of the chi-square statistic",
            sub="Simulations in blue, Chi-Square Curve in black",
            panel=function(x,...) {
              panel.mathdensity(dmath=dchisq,args=list(df=5),
                  lwd=5,col="black",lty=1,alpha=0.5,...)
              panel.densityplot(x,col="blue",lwd=2,...)
            })
```

## The Graph

```{r echo=FALSE}
densityplot(~sims,data=dieResults,plot.points=F,from=0,xlim=c(0,15),
            main="Simulations vs. Chi-Square Curve, df=5",
            xlab="Simulations of the chi-square statistic",
            sub="Simulations in blue, Chi-Square Curve in black",
            panel=function(x,...) {
              panel.mathdensity(dmath=dchisq,args=list(df=5),
                  lwd=5,col="black",lty=1,alpha=0.5,...)
              panel.densityplot(x,col="blue",lwd=2,...)
            })
```


## The Chi-Square Curve Approximation

We have seen that if your sample size is "big enough", it's OK to just use a chi-square curve to approximate the P-value:

```{r eval=FALSE}
chisqtestGC(~seat,data=m111survey,
            p=c(0.25,0.5,0.25),
            graph=TRUE)
```

## Summary

* The Chi-Square Test for Goodness of Fit is an inferential procedure for studying one factor variable

* It helps you determine whether the or not the factor variable is distributed in a specific way (as stated by the Null Hypothesis)

* If expected cell counts are big enough (at least five in each cell) then you can use a chi-square curve to approximate the P-value

* If not, just use simulation

* (You can always use simulation!)

# Another Example

## Picking "at Random"

```{r eval=FALSE}
data(pennstate1)
View(pennstate1)
help(pennstate1)
```

Subjects were asked to pick a whole number "at random" from 1 to 10.  The choice is recorded as the variable **RandNumb**.

## Descriptive Work

```{r}
choices <- xtabs(~RandNumb,data=pennstate1)
choices
```

```{r eval=FALSE}
barchartGC(choices,xlab="number chosen",
           main='Students Choosing "Randomly"')
```

## The Barchart

```{r echo=FALSE}
barchartGC(choices,xlab="number chosen",
           main='Students Choosing "Randomly"')
```

## A Test of Significance

Does the study provide strong evidence that people don't pick numbers randomly, when asked to pick "at random"?

* $H_0$:  Each number has a 10% chance to be picked

* $H_a:$ Some numbers were more likely to be picked than others were.

## Setting up the Code

We need to specify the null hypothesis in the `p` argument.

Suggestion:  use the `rep()` function:

```{r}
PennNullProbs <- rep(0.1,10)
PennNullProbs
```


## Practice { .practice }

Run the chi-square test for goodness of fit.  (Hint:  set `p=PennNullProbs`)

What do you conclude?

## Want Less Output?

You can set the argument `verbose` to `FALSE`:

```{r eval=FALSE}
chisqtestGC(~seat,data=m111survey,p=c(0.25,0.5,0.25),
            verbose=FALSE)
```


