---
title: "Introduction to R"
author: "Homer White"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css: slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```


# R as Calculator

## Basic Arithmetic

Operation|Sign|
|---|---|
addition|+|
subtraction|-|
multiplication|\*|
division|/|

<br>
Parentheses are used to clarify order of operations.

## Examples

```{r}
2+2
```

```{r}
2*3+7
```


```{r}
(23.7-20)/1.5
```

## Square Roots, Powers

Use the `sqrt()` function:

```{r}
sqrt(9)
```

To raise to a power, use "^".  To get $4^3$:

```{r}
4^3
```


## Practice 1 { .practice }

Use R to find:

$$2+7-3 \times 5,$$

$$\frac{25.5-20.2}{3},$$

and finally:

$$\sqrt{\frac{3.1^2}{25}+\frac{5.2^2}{36}}.$$

# Packages

## A System of Packages

* Packages extend the basic functionality of R.

* Some are already "loaded" and ready for use.

* Others must be loaded by you, the user (once per R session).

## Important Packages

To load them, use the `library()` function:

```{r eval=FALSE}
library(tigerstats)
```

You can also use `require()`, or just make sure the box next to the package name is checked.

Once loaded, the package stays loaded until you end your R session.

## Practice 2 { .practice }

Check the box next to `tigerstats`.  Then check to make sure that:

* `mosaic`
* `mosaicData`
* `abd`
* `tigerstats`

are all loaded.

# Data Frames

## Learning about a Dataset

`m111survey` comes from the `tigerstats` package.

```{r eval=FALSE}
data(m111survey)
View(m111survey)
help(m111survey)
```

## Data Frame Structure

* Each row stands for an individual (also ocalled an **observation**)

* The columns give *variables* that were measured for the individuals.

When data is missing, you see an NA.

Variables can be:

* Factor variables (values not naturally numbers)
* Numerical variables (integer and real number values)

## Structure

You can learn about the variables quickly using the `str()` function:

```{r}
str(m111survey)
```

## Structure

You can also click the arrow next to the data frame in the Global Environment.

Often you can tell just by viewing the data frame, or by looking at the Help file on the data frame.

## Practice 3 { .practice }

* How many students were in the MAT 111 survey?
* For each variable in `m111survey`, say whether the variable is a factor variable or a numerical variable.
* What does the variable **love_first** record, for each student?
* How many did students did not answer the question about their ideal height?

# Simple R Functions

## Making a "List" with c()

A "list" of four numbers:

```{r}
c(1,7,3,2)
```

You can store it in an object and see it any time you want

```{r}
myNumbers <- c(1,7,3,2)
myNumbers
```


## Making Lists with rep()

`rep()` repeats a number as many times as you want.

```{r}
rep(3,times=5)
rep(1/6,6)
someNumbs <- rep(3,10)
someNumbs
```


## Making Lists with seq()

Suppose you want all the odd numbers from 3 through 27:

```{r}
seq(3,27,by=2)
```

Or the numbers from -2 to 3, going in steps of 0.5:

```{r}
seq(-2,3,by=0.5)
```

## Practice 4 { .practice }

Use R to make a list of the numbers $1,5,-2$ and $7$.

Use R to make a list 100 fives.

Use R to make a list of the numbers from 100 to 220, going up in steps of 10.

# Simulation (optional)

## Coin Flipping

Flip a fair coin 16 times, and count the number of heads:

```{r eval=FALSE}
nflip(16)
```

```{r echo=FALSE}
library(parallel)
set.seed(2020)
nflip(16, quiet=TRUE)
```
I got 7 heads, but your results may vary from this!


## Repetition

```{r eval=FALSE}
do(10)*nflip(16)
```

```{r echo=FALSE}
library(parallel)
do(10)*nflip(16)
```

## Store Your Results

Let's do it lots of times, and store the results:

```{r eval=FALSE}
mySims <- do(500)*nflip(16)
head(mySims,n=5)
```


```{r echo=FALSE}
library(parallel)
mySims <- do(500)*nflip(16)
head(mySims,n=5)
```

## Make Histogram of Results

```{r eval=FALSE}
histogram(~nflip,data=mySims,
          xlab="Number of Heads",
          breaks=seq(-0.5,16.5,by=1),
          type="count"
          )
```

## Histogram of the Simulations

```{r echo=FALSE}
histogram(~nflip,data=mySims,
          xlab="Number of Heads",
          breaks=seq(-0.5,16.5,by=1),
          type="count"
          )
```

Does it seem likely to get 15 or more heads?

## Tallying the Simulations

```{r}
tally(~nflip,data=mySims)
```

How many times out of the 500 did we get 10 or more heads?  Is 10 or more an unusual result?

