---
title: "Probability in Sampling: Pt.1"
author: "Homer White"
date: "October, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(mosaic)
library(tigerstats)
```

# Definitions

## Parameters and Statistics

  >A *parameter* is a number associated with a population.
  

* It does not depend on chance
* but usually is unknown
  
  >A *statistic* is a number we compute from the sample data


* It can be known (we have the data)
* but it depends on sample (and so depends on chance)


## Estimators

We often use a statistic to *estimate* a parameter. Examples:

| Parameter | Estimator |
|:---------:|:----------:|
| population mean $\mu$ | sample mean $\bar{x}$ |
| population SD $\sigma$ | sample SD $s$ |
| population median | median of sample |
| population Q3 | Q3 of sample |

We hope that usually

$$\textbf{estimator} \approx \textbf{parameter}$$

## The "Basic Five" Parameters

| Parameter | Estimator |
|:---------:|:----------:|
| one mean $\mu$ | sample mean $\bar{x}$ |
| one proportion $p$ | sample proportion $\hat{p}$ |
| difference of two means $\mu_1-\mu_2$ | $\bar{x}_1-\bar{x}_2$ |
| difference of two proportions $p_1 - p_2$ | $\hat{p}_1-\hat{p}_2$ |
| mean of differences $\mu_d$ | sample mean of differences $\bar{d}$ |

# One Mean

## Research Question


>*Is the mean height in `imagpop` more than 66 inches?*

## Parameter of Interest


One population mean.  Let's define it:

  >Let $\mu =$ the mean height of all people in `imagpop`.

## The Estimator


The mean of the sample, $\bar{x}$.

This is a random variable, and it has an expected value and a standard deviation:

$$EV(\bar{x}) = \mu \\
SD(\bar{x}) = \frac{\sigma}{\sqrt{n}}$$

where

* $\sigma$ is the SD of the population
* $n$ is the sample size

## Example

* A population of males has a mean height $\mu$ of 72 inches.

* The standard deviation $\sigma$ of the population is 3 inches.

* You pick one male at random from the population.

How tall do you expect him to be?  Give or take about how much?

## Solution

Let $X$ be the height of the randomly-picked male.  When you sample just one guy, his height IS the sample mean $\bar{x}$!

$EV(X) = \mu = 72$, so we expect the guy to be about 72 inches tall.

$SD(X) = \sigma/\sqrt{n} = 3/\sqrt{1} = 3$, so give-or-take figure is 3 inches.

We expect the guy to be around 72 inches tall, give or take about 3 inches for chance variation in the sampling process.

## Example

Same population of guys:

* $\mu = 72$
* $\sigma = 3$

This time you pick $n=4$ guys at random and compute the mean $\bar{x}$ of their heights.

What do you expect the sample mean to be?  Give or take about how much?

## Solution

$EV(\bar{x}) = \mu = 72$, so we expect the guy to be about 72 inches tall.

$SD(\bar{x}) = \sigma/\sqrt{n} = 3/\sqrt{4} = 3/2 = 1.5$

We expect sample mean to be around 72 inches, give or take about 1.5 inches for chance variation in the sampling process.

## Practice { .practice }

Same population of guys:

* $\mu = 72$
* $\sigma = 3$

This time you pick $n=25$ guys at random and compute the mean $\bar{x}$ of their heights.

What do you expect the sample mean to be?  Give or take about how much?

## Moral of the Story

If you  do random sampling, then:

* as sample size $n$ goes up ...
* chance variation goes down!

The bigger your sample size, the more likely it is that $\bar{x}$ will be close to $\mu$!

## We Apologize ...

... for the mess:  there are THREE "standard deviations" to keep track of:

* $s$, the SD of the sample (says how much a typical data value differs from the mean of the sample)

* $\sigma$, the SD of the population (says how much a typical member of the population differs form the mean of the population)

* The SD of a random variable (says how much a random variable is liable to differ from its expected value)


# One Proportion


## Research Question


>What proportion of all college students are math majors?

## Parameter of Interest


One population proportion.  Let's define it:

Let $p =$ the proportion of all college students who are math majors.

## The Estimator


The sample proportion:

$$\hat{p}=\frac{\textbf{number in sample who are math majors}}{\textbf{sample size}}.$$

This is a random variable, and it has an expected value and a standard deviation:

$$EV(\hat{p}) = p \\
SD(\hat{p}) = \sqrt{\frac{p(1-p)}{n}}$$

where $n$ is the sample size.

## Example

Suppose that 10% of all college students are math majors.

Suppose you take a random sample of 900 college students.

About what percentage of the students in your sample will be math majors?  Give or take about how much?

## Solution

$EV(\hat{p})=p = 0.10$, or 10%.

$n=900$, so $SD(\hat{p})$ is

```{r}
sqrt(0.1*(1-0.1)/900)
```

So you expect about 10% of the sample to be math majors, give or take 1% or so.

## Practice { .practice }

Suppose that 45% of all registered voters in KY favor Alison Lundergren Grimes in the Senate race.

Suppose you take a random sample of 635 registered voters.

About what percentage of the students in your sample will favor Alison?  Give or take about how much?

# Two Means

## Research Question

>*We wonder how much taller the GC guys are, on average, than GC gals.*



## Parameter of Interest

Divide the GC population into two separate populations:

* all of the females, and
* all of the males.

The parameters are:

  >$\mu_1$ = mean height of all GC males, and

  >$\mu_2$ = mean height of all GC females

We are interested in the value of $\mu_1-\mu_2$:

* if $\mu_1-\mu_2 >0$, then guys are taller, on average
* if $\mu_1-\mu_2 <0$, then gals are taller, on average


## The Estimator


* We take a simple random sample of size $n_1$ of guys, computing sample mean $\bar{x}_1$.
* We take an *independent* simple random sample of size $n_2$ of gals, computing sample mean $\bar{x}_2$.

The estimator is $\bar{x}_1 - \bar{x}_2$, the difference of sample means.

$$EV(\bar{x}_1 - \bar{x}_2) = \mu_1-\mu_2 \\
SD(\bar{x}_1-\bar{x}_2)=\sqrt{\frac{\sigma_1^2}{n_1}+\frac{\sigma_2^2}{n_2}}$$

where $\sigma_1$ and $\sigma_2$ are the population SDs.

# Two Proportions

## Research Question

*Who is more likely to believe in love at first sight:  a GC gal or a GC guy?*

## Parameters

Divide the GC population into two separate populations:

* all of the females, and
* all of the males.

The parameters are:

  >$p_1$ = proportion of all GC females who believe in love at first sight, and

  >$p_2$ = proportion of all GC males who believe in love at first sight
  
## Parameters

We are interested in $p_1-p_2$.

* if $p_1-p_2 >0$, then GC gals are more likely to believe than GC guys are;
* if $p_1-p_2 <0$, then it's the other way around!

## The Estimator


* We take a simple random sample of size $n_1$ of gals, computing sample proportion $\hat{p}_1$.
* We take an *independent* simple random sample of size $n_2$ of guys, computing sample proportion $\hat{p}_2$.

The estimator is $\hat{p}_1 - \hat{p}_2$, the difference of sample proportions.

$$EV(\hat{p}_1 - \hat{p}_2) = p_1-p_2 \\
SD(\hat{p}_1 - \hat{p}_2) = \sqrt{\frac{p_1(1-p_1)}{n_1}+\frac{p_2(1-p_2)}{n_2}}$$

# Mean of Differences

## Research Question

*Do Tidarren spiders run faster after one pedipalp has been amputated?*

```{r eval=FALSE}
data(SpiderSpeed)
View(SpiderSpeed)
help(SpiderSpeed)
```


## The Parameter

This was a repeated-measures design, so we are interested in the differences.

Let $\mu_d$ be the mean difference in speed (after minus before) for ALL Tidarren spiders.

(Not just the 16 in the sample!)

If $\mu_d > 0$, then on average spiders run faster after amputation!

## Estimator


* We take a simple random sample of size $n$ for the population
* record ideal height and actual height for each person
* compute the difference (ideal minus actual) for each person
* compute the mean of these differences (written $\bar{d}$)


$$EV(\bar{d}) = \mu_d \\
SD(\bar{d})=\frac{\sigma_d}{\sqrt{n}}$$

where $\sigma_d$ is the SD of the differences for the population.

# Tips for Identifying the Parameter

## Keywords


  >*On average*, do adults in the U.S. make more than 38,000 dollars per year?

"On average" often indicates a mean.

Parameter is:

  >$\mu =$ mean income of all adults in the U.S..
  
## Keywords


  >Do a *majority* of people in the U.S. favor capital punishment?
  
Words like "majority" and "minority" often indicate an interest in a proportion.

The parameter is:

  >$p =$ proportion of all people in the U.S. who favor capital punishment.
  
## Variable Analysis

When the Research Question involves two variables:

* identify each variable
* if relevant, decide which is explanatory and which is response
* classify each one (factor?  numerical?)
* for each factor, count how many levels it has


## Variable Analysis:  Example 1


Research Question:

  >*On average, who drives faster at GC:  a guy or a gal?*
  
Variable Analysis:

* One variable is **sex**
    * it is the explanatory variable
    * it is a factor
    * it has two values:  "male" and "female"
* The other variable is **fastest**:
    * it is the response variable
    * it is numerical
    
## Variable Analysis:  Example 1


If

* explanatory variable is a factor with two values, and
* response variable is numerical

then we are probably interested in the difference of two means.

(In this Research Question, the keyword "average" further supports this idea.)

## Variable Analysis:  Example 1


So we define

  >$\mu_1 =$ mean fastest speed ever driven, for all males at GC.
  
and

  >$\mu_2 =$ mean fastest speed ever driven, for all females at GC.
  
We are interested in $\mu_1-\mu_2$.

## Variable Analysis:  Example 2


Research Question:

  >*At GC, who is more likely to believe in love at first sight:  a guy or a gal?*
  
Variable Analysis:

* One variable is **sex**
    * it is the explanatory variable
    * it is a factor with two values ("male","female")
* The other variable is **love_first**:
    * it is the response variable
    * it is a factor with two values ("no","yes")
    

## Variable Analysis:  Example 2


If

* explanatory variable is a factor with two values, and
* response variable is a factor with two values

then we are probably interested in the difference of two proportions.

(Proportions can stand for probabilities, so in this Research Question, the keyword "likely" further points toward proportions.)

## Variable Analysis:  Example 2


So we define

  >$p_1 =$ proportion of all males at GC who believe in love at first sight.
  
and

  >$p_2 =$ proportion of all females at GC who believe in love at first sight.
  
We are interested in $p_1-p_2$.

## Variable Analysis:  Example 3


Research Question:

  >*On average, do people at GC want to be taller than they actually are?*
  
Variable Analysis:

* One variable is **height**
    * it is numerical
* The other variable is **ideal_height**:
    * it is numerical

Explanatory/Response distinction does not apply here:  we have repeated measures instead.


## Variable Analysis:  Example 3


When

* you are in a repeated measures or matched pairs situation, and
* both variables are numerical

then you are probably interested in the mean of differences.

## Variable Analysis:  Example 3


So we are interested in

  >$\mu_d =$ mean difference (ideal height minus actual height) for all students at Georgetown College.
  