---
title: "Probability in Sampling: Pt.2"
author: "Homer White"
date: "March, 2015"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# The Central Limit Theorem


## Center and Spread


The estimators for the Basic Five Parameters are all random variables, so they all have a center (EV) and a spread (SD).

|  Estimator | Center | Spread |
|:--------:|:----------:|:-----------:|
|$\bar{x}$|$\mu$|$\frac{\sigma}{\sqrt{n}}$|
|$\bar{x}_1-\bar{x}_2$|$\mu_1-\mu_2$|$\sqrt{\frac{\sigma_1^2}{n_1}+\frac{\sigma_2^2}{n_2}}$|
|$\hat{p}$|$p$|$\sqrt{\frac{p(1-p)}{n}}$|
|$\hat{p}_1-\hat{p}_2$|$p_1-p_2$|$\sqrt{\frac{p_1(1-p_1)}{n_1}+\frac{p_2(1-p_2)}{n_2}}$|
|$\bar{d}$|$\mu_d$|$\frac{\sigma_d}{\sqrt{n}}$|


## Estimating One Mean


To investigate the shape of the distribution of $\bar{x}$, try:

[http://homer.shinyapps.io/CentralLimit](http://homer.shinyapps.io/CentralLimit)



## The Pattern


As sample sizes increase, shape of the distribution of $\bar{x}$ looks more and more bell-shaped.


**No matter what the underlying population looks like!!**

But:

* the more skewed the population is, and
* the more far-out its outliers are

the bigger the sample needs to be before $\bar{x}$ starts looking normal.

## Central Limit Theorem (CLT)


No matter how the population is distributed, as sample size $n$ increases, the distribution of $\bar{x}$ gets closer and closer to:

$$norm(\mu,\frac{\sigma}{\sqrt{n}}).$$

## Another Way  to Put It

No matter how the population is distributed, as sample size $n$ increases, the distribution of

$$Z=\frac{\bar{x}-\mu}{\sigma/\sqrt{n}}$$

gets closer and closer to:

$$norm(0,1),$$

the *standard normal* distribution.

## How Big is Big Enough?

Many folks say $n \geq 30$ is "big enough" for $\bar{x}$ to look normal.

But it's just a rule of thumb!

## The CLT Carries Over ...


... to the other four of the Basic Five estimators!

As long as sample sizes are "big enough", their shape will be approximately normal!

## How Big is "Big Enough"?


Rules of Thumb:

|  Estimator | Big Enough |
|:--------:|:----------:|
|$\bar{x}$|$n \ge 30$|
|$\bar{x}_1-\bar{x}_2$|$n_1 \ge 30,n_2 \ge 30$|
|$\hat{p}$|$np \ge 10,n(1-p) \ge 10$|
|$\hat{p}_1-\hat{p}_2$| all of $n_1p_1,n_1(1-p_1),n_2p_2,n_2(1-p_2) \ge 10$|
|$\bar{d}$|$n \ge 30$|


# Probability With Estimators

## Imagine ...


... that you are a powerful being (a Greek god/goddess, maybe).

* You know everything about the present
* so you have complete information on all populations
* so you can instantly find any population parameter you like
* but you do NOT know the future

## Imagine ...


You see a poor mortal, a statistician, about to take a simple random sample of size $n$ from a population.

The following is all true:

* You know population mean $\mu$,
* you know population SD $\sigma$,
* and you know that $\bar{x} \approx norm(\mu,\sigma/\sqrt{n}).$
* You don't know what $\bar{x}$ will be,
* but you can compute probabilities for it to lie in various ranges.

## Example

Recall `imagpop`:

```{r eval=FALSE}
data(imagpop)
View(imagpop)
help(imagpop)
```

## Example

  > A statistician is about to take a SRS of size $n=25$ from `imagpop` and compute $\bar{x}$, the sample mean of the heights of the 25 selected individuals.
  
  >*What is the probability that the sample mean will exceed 68.3 inches?*
  
In other words, what is:

$$P(\bar{x} > 68.3)?$$

## Solution


First of all, we use our god-like powers (and R!) to find:

* mean $\mu$ and
* standard deviation $\sigma$

of the heights in the population:

```{r}
favstats(~height,data=imagpop)[c("mean","sd")]
```

## Solution


So we know that:

*  $EV(\bar{x}) = 67.53$, and
*  $SD(\bar{x}) = 3.907/\sqrt{25}=0.7814$.

We know the center and the spread of $\bar{x}$!

## Solution


How about the shape of $\bar{x}$?  

* Sample was size $n=25$,
* a bit less than the "cut-off" of 30 where CLT kicks in.

Better use our god-like powers to make density plot of the population:

```{r eval=FALSE}
densityplot(~height,data=imagpop,
            xlab="Height (inches)",
            main="Imagpop Heights",
            plot.points=FALSE)
```




## The Population


```{r echo=FALSE}
densityplot(~height,data=imagpop,
            xlab="Height (inches)",
            main="Imagpop Heights",
            plot.points=FALSE)
```

Looks fairly bell-shaped!

## Solution


So even though $n=25 < 30$, we still figure that

$$\bar{x} \approx norm(67.53,0.7814).$$

So for $P(\bar{x} > 68.3)$, go for:

```{r eval=FALSE}
pnormGC(bound=68.3,region="above",
        mean=67.53,sd=0.7814,
        graph=TRUE)
```

## Solution

```{r echo=FALSE}
pnormGC(bound=68.3,region="above",
        mean=67.53,sd=0.7814,
        graph=TRUE)
```


## Practice { .practice }

In a population of college students, the mean fastest speed ever driven is 110 mph.  The SD of the fastest speeds is 10 mph.

Herbert will sample 64 college students at random and compute the sample mean $\bar{x}$.

Approximately what is the probability that $\bar{x}$ will be less than 107 mph?

## Example

49% of all registered Deomocrats in the United States support Bernie Sanders for President.

Harriet plans to take a random sample of 1600 registered Democrats.  She will compute $\hat{p}$, the proportion of them that support Sanders.

Approximately what is

$$P(0.47 < \hat{p} < 0.51)?$$

## Solution

Think about center, spread and shape of $\hat{p}$:

* Center:  $EV(\hat{p})=p=0.49$

* Spread:  $SD(\hat{p}) = \sqrt{p(1-p)/n}$, which is:

```{r}
sqrt(0.49*(1-0.49)/1600)
```

The sample percentage should be around 49%, give or take 1.25% or so for chance variation.

## Solution

Shape of distribution of $\hat{p}$:

* $np = 1600 \times 0.49 = 784 \geq 10$

* $n(1-p) = 1600 \times 0.51 = 816 \geq 10$

So by the CLT, $\hat{p} \approx$ normal!

## Solution

So you can use `pnormGC()` to find the chance:

$$P(0.47 < \hat{p} < 0.51) =$$

```{r eval=FALSE}
pnormGC(c(0.47,0.51),region="between",
        mean=0.49,sd=0.0124975,
        graph=TRUE)
```

---

```{r echo=FALSE}
pnormGC(c(0.47,0.51),region="between",
        mean=0.49,sd=0.0124975,
        graph=TRUE)
```

## Moral

When you use random sampling, polls can be pretty accurate!

* There are millions of registered Democrats
* Harriet is sampling only 1600 of them
* But there is about an 89% chance that her estimate is within 2% of the truth!

# 68-95 Rule, Again


## 68-95 Rule:  Review


If the distribution of a numerical variable is roughly bell-shaped, then

* about 68% of the values are within one SD of the mean
* about 95% are within 2 SDs of the mean
* about 99.7% are within 3 SDs of the mean


## 68-95 Rule for Probability


If the probability distribution of a random  variable $X$ is approximately normal, then

* there is about a 68% chance that $X$ will turn out to be within one SD of its EV
* there is about a 95% chance that $X$ will turn out to be within two SDs of its EV
* there is about a 99.7% chance that $X$ will turn out to be within three SDs of its EV


## Application


You are about to take a simple random sample from a population.

* The mean $\mu$ is 50.
* The standard deviation $\sigma$ of the population is 6.
* The size of your sample will be $n=36$.

Approximately what is

$$P(\bar{x} > 52)?$$

## Solution


* $EV(\bar{x})=\mu=50$
* $SD(\bar{x})=\sigma/\sqrt{n}=6/\sqrt{36}=1$
* $n=36 \ge 30$, so by CLT $\bar{x} \approx norm(50,1)$


52 is 2 SDs above the EV. 

---

```{r, echo=FALSE}
QuickEmpRuleGraphs <- function(mean=0,sd=1,xlab="x",range=1,shade="between") {

  curve(dnorm(x,mean=mean,sd=sd),from=mean-4*sd,to=mean+4*sd,xlab=xlab,ylab="density",
       main=paste("68-95 Rule:\nmean =",mean,", SD =",sd),axes=FALSE)
 DesiredLabels <- rep("",9)
 DesiredLabels[5] <- as.character(mean)
 DesiredLabels[5-range] <- as.character(mean-sd*range)
 DesiredLabels[5+range] <- as.character(mean+sd*range)
 axis(side=1,at=mean+sd*seq(-4,4,by=1),labels=DesiredLabels)
 
 if (shade=="between")  {
   x.coords <- seq(from=mean-sd*range,to=mean+sd*range,
                   length.out=500)
   y.coords <- dnorm(x.coords,mean=mean,sd=sd)
   x.coords <- c(mean-sd*range,x.coords,mean+sd*range)
   y.coords <- c(0,y.coords,0)
   polygon(x.coords,y.coords,col="lightblue")
   percs <- c("~68%","~95%","~99.7%")
   text(x=mean,y=0.45/(sqrt(2*pi)*sd),labels=percs[range],cex=4)
 } else {
   x.coords.left <- c(mean-4*sd,
                      seq(from=mean-4*sd,to=mean-sd*range,length.out=300),
                      mean-sd*range)
   x.coords.right <- c(mean+sd*range,
                       seq(from=mean+sd*range,to=mean+4*sd,length.out=300),
                       mean+4*sd)
   y.coords.left <- c(0,dnorm(seq(from=mean-4*sd,to=mean-sd*range,length.out=300),mean=mean,sd=sd),0)
   y.coords.right <- c(0,
                       dnorm(seq(from=mean+sd*range,to=mean+4*sd,length.out=300),mean=mean,sd=sd),
                       0)
   polygon(x.coords.left,y.coords.left,col="lightblue")
   polygon(x.coords.right,y.coords.right,col="lightblue")
   x.txt <- mean-sd*c(2.3,3,3.55)
   y.txt <- c(0.4,0.25,0.15)/(sqrt(2*pi)*sd)
   percs <- c("~16%","~2.5%","~0.15%")
   text(x.txt[range],y.txt[range],percs[range],cex=4)
   x.txt.right <- mean+sd*c(2.3,3,3.55)
   text(x.txt.right[range],y.txt[range],percs[range],cex=4)
                            
 }

}  #end EmpRuleGC

QuickEmpRuleGraphs(mean=50,sd=1,xlab="Sample Mean",range=2,shade="between")
```

## Think "Outside"


```{r echo=FALSE}
QuickEmpRuleGraphs(mean=50,sd=1,xlab="Sample Mean",range=2,shade="outside")
```

## Solution


So,

$$P(\bar{x} > 52) \approx 2.5\%.$$


## Use the Right SD!


Use the SD for the random variable you are working with, NOT the SD of the population.

$$SD(\bar{x})=\frac{\sigma}{\sqrt{n}}=\frac{6}{\sqrt{36}}=1,$$

$$SD(\bar{x}) \neq \sigma=6$$

## Practice { .practice }

In a population of high school students, the mean fastest speed ever driven is 106 mph.  The SD of the fastest speeds is 12 mph.

Herbert will sample 36 college students at random and compute the sample mean $\bar{x}$.

Approximately what is the probability that $\bar{x}$ will be between 104 and 108 mph?
