---
title: "Sampling and Surveys: Pt.1"
author: "Homer White"
date: "September, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(mosaic)
library(tigerstats)
```

# Population vs. Sample


## Population and Sample

Definitions:

* The *population* is the set of all items of interest.
  
* The *sample* is the subset of the population for which we have data.
  
## Parameters

Another definition:

* A *parameter* is a number that you could compute if you knew the entire population.
  
Suppose the population is **all adults**.

Examples of parameters:

* The mean height $\mu$ of the population
* The standard deviation $\sigma$ of the heights
* The proportion $p$ of the population that plays tennis

## Statistics

Another definition:


* A *statistic* is a number that you can compute from your sample data.
  
Suppose that we have a sample from the population of all adults.  Examples of statistics:

* The sample mean $\bar{x}$ of the heights
* The sample standard deviation $s$ of the heights
* The proportion $\hat{p}$ of the sample that plays tennis
    * Note:  $\hat{p}$ is the number $X$ in the sample who play tennis, divided by the sample size $n$.
    
## The Basic Ideas


* The parameters are fixed, but we don't know them
* The statistics depend on the sample
* We use statistics to estimate parameters.  That is, we hope that:
    * $\bar{x} \approx \mu$
    * $s \approx \sigma$
    * $\hat{p} \approx p$
 
 
# How to Sample

## Important Thought

For the approximations to be good, the sample should be *representative* of the population.
  
So we should employ methods of sampling for which the sample is **likely** to be representative of the population.

*How do we get a representative sample?*

## One Try:  a Volunteer Sample

Should we let potential subjects choose whether or not to be in the sample?

(This is called a *volunteer* sample.)

**Example**:  When conducting an opinion survey on food in the Cafe, you leave forms at the entrance for people to fill out.


## There is a Problem!


The volunteers might differ from the general population in some important way.

(In our example, the students who take the time to fill out the survey might have stronger opinions---one way or another---than those who don't bother.)
  
## Another Try:  Researchers Decide Sample


Should the researchers decide who gets into the sample?

Example:  *Quota* sampling in the 1948 U.S. presidential elections.

## Problem Here, Too!


When researchers use their own judgement to decide on the sample, they could (intentionally or unintentionally) choose an unrepresentative sample.

In the 1948 quota sampling, pollsters ended up interviewing "approachable" folks, who turned out to be wealthier than average, thus biasing poll results toward the Republican candidate.

## Wait a Minute ...


* If you can't let subjects decide whether to be in the sample,
* and you can't let researchers decide who should be in the sample ...

... then what should decide who gets into the sample?

## Surprising Answer


  *Let chance decide who gets into the sample!*
  
We should use some form of *random sampling*.  That is, we should use chance in a controlled, quantifiable way.

This is called **random sampling**.

## Simple Random Sampling


There are many types of random sampling.  The one we will think about the most is *simple random sampling.*

Suppose you are planning to take a sample of size $n$ from a population. If you take the sample so that

* every set of $n$ subjects in the population has the same chance to be the sample selected
  
then you are doing *simple random sampling* (SRS).

## SRS is Like ...


Having a box full of tickets, one for each member of the population.

*  You randomly pick out one ticket ...
* and set it aside ...
* then randomly pick out another ticket
* and set it aside ...
* ... and so on until you have drawn $n$ tickets.

(You draw $n$ tickets at random from the box, **without** replacement.)

## "Is it SRS?"

Suppose you have 4 rows in a class.  Each row has six people.

You need a sample of size 4.

You pick one person at random, from each row.

*Is this simple random sampling?*

## No!

Not every set of size four has the same chance to be the sample selected.

For example, a set like this:

* two people from row 1,
* one person from row 3 and
* one person from row 4

Has NO chance to be the sample selected!

## "Is it SRS?" { .practice }

You have a population of 1000 people, all lined up by order of their height.  You need a sample of size 10.

You pick a number at random from 1 to 100.  (Say you get 17.)

Then you pick people numbered:

17, 117, 217, 317, ..., 917

Is this procedure simple random sampling?

## Another Way to Think of It

In SRS:

* every individual in the population has the same chance to be in the sample;
* knowing some members of the sample doesn't tell you anything "extra" about who else is or is not in the sample.


## SRS Works Amazingly Well ...

... especially when the sample size $n$ is large.

Try this app:

```{r eval=FALSE}
require(manipulate)
SimpleRandom2(n=100) #n specifies the sample size
```

*In this course we will assume that all random sampling is SRS, unless otherwise stated.*

## Random Sampling Methods

SRS is not the only good random sampling method.  Some other useful methods are:

* *stratifed sampling*
* *cluster sampling*

To compare SRS with these two methods, try the Shiny app **Sampling Methods**:

* [http://homer.shinyapps.io/SamplingMethods](http://homer.shinyapps.io/SamplingMethods)