---
title: "Student Success on the SAT"
author: "Homer White"
date: "February, 2016"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---


```{r include=FALSE}
library(tigerstats)
```

# SAT and School Expenditure

## Research Question:

>*How does expenditure on schools affect student performance?*

```{r eval = FALSE}
data(sat)
View(sat)
help(sat)
```


## Teacher Salary and SAT Scores

Looking at relationship between:

* **salary** (numerical)
* **sat** (numerical)

```{r eval = FALSE}
xyplot(sat ~ salary, data = sat,
       type = c("p","r"))
```

##

```{r echo = FALSE}
xyplot(sat ~ salary, data = sat,
       type = c("p","r"))
```

Hmm, not what we would expect!

## Expenditure per Student and SAT


Looking at relationship between:

* **expend** (numerical)
* **sat** (numerical)

```{r eval = FALSE}
xyplot(sat ~ expend, data = sat,
       type = c("p","r"))
```

##

```{r echo = FALSE}
xyplot(sat ~ expend, data = sat,
       type = c("p","r"))
```

Again, not what we would expect!

## How about smaller Class sizes?

Looking at relationship between:

* **ratio** (numerical)
* **sat** (numerical)

```{r eval = FALSE}
xyplot(sat ~ ratio, data = sat,
       type = c("p","r"))
```

##

```{r echo = FALSE}
xyplot(sat ~ ratio, data = sat,
       type = c("p","r"))
```

Again, not what we would expect!

## So it appears that ...

... in order to raise ACT scores we should:

* pay teachers less
* spend less on the students
* pack 'em in like sardines!

This can't be right!  What's going on?

## Percentage in State Taking the SAT

**frac** is the percentage of seniors in the state that take the SAT test.

How does this relate to student performance?

```{r eval = FALSE}
xyplot(sat ~ frac, data = sat,
       type = c("p","r"))
```

##

```{r echo = FALSE}
xyplot(sat ~ frac, data = sat,
       type = c("p","r"))
```

High frac goes with low act!

## frac and salary

```{r eval = FALSE}
xyplot(frac ~ salary, data = sat,
       type = c("p","r"))
```

##

```{r echo = FALSE}
xyplot(frac ~ salary, data = sat,
       type = c("p","r"))
```

States with high **frac** also pay their teachers more!

## Confounding Variable

**frac** is a *confounding variable*:

* it's related to the explanatory variable **salary**
* It helps to cause the response variable **sat**

We need to "correct" for **frac**.

##

```{r eval = FALSE}
Percentage <- equal.count(sat$frac, number = 3, overlap = 0.1)
xyplot(sat ~ salary | Percentage,
	data = sat,
	layout = c(3,1),
	type = c("p","r"),
	pch = 19,
	main = "Salary and Mean SAT,\nby percentage taking SAT",
	xlab = "Mean Teacher Salary ($1000s)",
	ylab = "Mean SAT Score")
```

##

```{r echo = FALSE}
Percentage <- equal.count(sat$frac, number = 3, overlap = 0.1)
xyplot(sat ~ salary | Percentage,
	data = sat,
	layout = c(3,1),
	type = c("p","r"),
	pch = 19,
	main = "Salary and Mean SAT,\nby percentage taking SAT",
	xlab = "Mean Teacher Salary ($1000s)",
	ylab = "Mean SAT Score")
```

## Another Tool:  Grouping

You can also group by **frac**.  It's a numerical variable, though, so you have to break it up first:

```{r eval = FALSE}
sat$fracGroup <- cut(sat$frac, 
                     breaks = 3, 
                     labels = c("low","middle","high"))
sat
```


##

```{r echo = FALSE}
sat$fracGroup <- cut(sat$frac, 
                     breaks = 3, 
                     labels = c("low","middle","high"))
xyplot(sat ~ salary,
	data = sat,
	groups = fracGroup,
	auto.key = list(
		space = "top",
		title = "Group",
		cex.title = 1,
		columns = 3),
	type = c("p","smooth"),
	pch = 19)
```



## Practice { .practice }

Make a graph to examine the relationship between ratio and SAT, but correct for the percentage of students who take the SAT.






