---
title: "Tests of Significance: Pt.3"
author: "Homer White"
date: "April, 2016"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---


```{r include=FALSE}
library(tigerstats)
```

# One Proportion

## Research Question


Are a majority of Georgetown College Students female?
  
We'll investigate this question inferentially with the `m111survey` data.

## Parameter and Hypotheses


Let

* $p =$ proportion of all GC students who are female
  
Our hypotheses are:

* $H_0: p=0.50$
  
* $H_a:  p > 0.50$ (one-sided, for variety!)
  
Note:  0.50 is $p_0$, the Null's belief about $p$.
  
## The Code


```{r eval=FALSE}
binomtestGC(~sex,data=m111survey,
        success="female",
        p=0.5,alternative="greater",
        graph=TRUE)
```

Notes:

* We set the argument `p` to $p_0$, the Null's belief about $p$.
* `graph=TRUE` will give us a graph of the P-value

## Some Results


* $\hat{p}=40/71=0.5634$
* $SE(\hat{p})=0.0589$
* If Null is correct, number of successes $X \sim binom(71,0.5)$
* No test statistic given!
* (But you can think of the 40 successes as the test statistic)
* $P$-value is $P(X \geq 40 \vert p=0.5)$

## Graph of P-Value


```{r echo=FALSE}
pbinomGC(bound=39,region="above",size=71,prob=0.5,graph=TRUE)
```

## Rest of the Analysis


* $P = 0.1712$
* Interpretation of P-value:  "If only 50% of all GC students are female, then there is about a 17.12% chance of getting 40 or more females in a sample of 71, as we did in this study."
* Our results are not so unlikely, if $H_0$ is true.
* Since $P > 0.05$, we do not reject $H_0$.
* This study did not provide strong evidence that a majority of GC students are female.


## Another Example


A random sample of 2500 registered voters shows that 1325 of them favor the Affordable Care Act.

*Do a majority of all registered voters in the U.S. favor the Act?*
  
## Parameter and Hypotheses


Let

* $p =$ proportion of all registered voters in the U.S. who favor the Act
  
Our hypotheses are:

* $H_0: p=0.50$
  
* $H_a:  p \neq 0.50$ (two-sided)
  
## The Code


We have summary data.

* argument `x` is set to the number of successes
* argument `n` is set to the sample size

```{r eval=FALSE}
binomtestGC(x=1325,n=2500,p=0.50)
```

## The Results


* We had 1325 voters approve of the Act (this is like a test statistic).
* $\hat{p}=1325/2500=0.53$
* $SE(\hat{p}) \approx 0.01$
* Our results are about 3 SE's above what $H_0$ expected.
* $P = 0.0029$
* Since $P < 0.05$, we reject $H_0$.
* We have strong evidence that a majority of registered voters in the U.S. favor the Act.

## Practice { .practice }

In the 2014 Major League season, Detroit Tiger Miguel Cabrera had a batting average of .313 on 611 at-bats.  Can we conclude from this data that in 2014 his chance of getting a hit was more than 30%?

Define $p$ to be the probability that Cabrera gets a hit when he is at bat.

* $H_0: p = 0.3$
* $H_a: p \neq 0.3$

Perform the two sided test with `binomtestGC()`, using the summary data:

* What should you enter for `n`, the number of trials?
* What should you enter for `x`, the number of successes
* What should you enter for `p`?
* Can you conclude that Cabrera's chance was better than 30%?

# Tests for Two Proportions


## Capital Punishment:  GSS 2002

```{r eval=FALSE}
data(gss02)
View(gss02)
help(gss02)
```

The variable **cappun** says whether or not the respondent favors the death penalty.

## The Tally

```{r}
xtabs(~cappun,data=gss02)
```

899 out of 1308 (68.7%) approved of the death penalty.

## Capital Punishment:  GSS 2012

```{r eval=FALSE}
data(gss2012)
View(gss2012)
help(gss2012)
```

```{r}
xtabs(~cappun,data=gss2012)
```

1183 out of 1824 (64.8%) approved.

## Research Question


*In the U.S. population, did support for capital punishment decline between 2002 and 2012?*
  
**Notice**:  This is a question about the relationship between two factor variables:

* Explanatory variable is the **year under study** (2002,2012).
* Response variable is **opinion about capital punishment** (favor, oppose).
* Both variables have two values,
* so we are interested in $p_1-p_2$.
  
## Parameters and Hypotheses


Let

* $p_1$ = the proportion of all U.S. adults in 2002 who support capital punishment

* $p_2$ = the proportion of all U.S. adults in 2012 who support capital punishment

Then our hypotheses are:

* $H_0$:  $p_1 - p_2 = 0$

* $H_a$:  $p_1 - p_2 \neq 0$
  
## The Code


We have summary data:

|Year|$x$|$n$|
|:---:|:----:|:----:|
|year 2002|899|1308|
|year 2012|1183|1824|

<br>

```{r eval=FALSE}
proptestGC(x=c(899,1183),
           n=c(1308,1824),
           p=0,graph=TRUE)
```

## Estimator and SE


```
Estimate of p1-p2:   0.03873 
SE(p1.hat - p2.hat):	 0.01701
```
* $H_0$ expected $\hat{p}_1-\hat{p}_2$ to be about 0
* We actually got $\hat{p}_1-\hat{p}_2 = 0.03873$
* This is a bit more than two SEs above what $H_0$ expected

## Confidence Interval


95%-confidence interval for $p_1-p_2$ is

```
lower       upper          
0.005399    0.072069
```

Observe that 0 (Null's belief about  $p_1-p_2$) is below this interval.

## Test Statistic


```
Test Statistic:    z = 2.277
```

$$z=\frac{\hat{p}_1-\hat{p}_2}{SE(\hat{p}_1-\hat{p}_2)}=\frac{\textbf{Observed Difference}}{\textbf{Standard Error}}$$


* "$z$-score style" (yet again!)
* Tell you how many SEs $\hat{p}_1-\hat{p}_2$ is above or below what $H_0$ expected

## P-value

Statistical theory says that if $H_0$ is true then:

$$z \approx norm(0,1)$$

R uses this fact to approximate the P-Value.

## P-Value


```
P-value:  P = 0.0228
```

Interpretation:  If support for the death penalty remained constant from 2002 to 2012, then there is only about a 2.28% chance of getting a test statistic at least as far from 0 as the one we actually got.

## Graph of the P-Value

```{r echo=FALSE}
pnormGC(c(-2.277,2.277),region="outside",mean=0,sd=1,graph=TRUE)
```

  
## Decision and Conclusion


* Since $P < 0.05$, we will reject $H_0$.
* The GSS data provided strong evidence that support for the death penalty declined a bit between 2002 and 2012.

## Astrology and Political Preference

* The [Chapman University Survey on American Fears](http://www.chapman.edu/wilkinson/research-centers/babbie-center/survey-american-fears.aspx) recently published results of a survey of 2500 U.S. Adults.

* 14% of Democrats surveyed said that they believed in astrology

* 8% of Republicans surveyed said that they believed in astrology

## Guess at a Table

Here is a guess at the numbers in the survey:

|Party affiliation|believe in astrology|do not believe|
|:---:|:----:|:----:|
|Democratic|81|500|
|Republican|36|412|

<br>

Do these results provide strong evidence that Democrats are more likely to believe in astrology than Republicans are?

## Parameters and Hypotheses

Define:

* $p_1 =$ the proportion of all Democrats who believe in astrology

* $p_2 =$ the proportion of all Republicans who believe in astrology

Let's test:

* $H_0$:  $p_1 - p_2 = 0$

* $H_a$:  $p_1 - p_2 \neq 0$

## Practice { .practice }

|Party affiliation|believe in astrology|do not believe|
|:---:|:----:|:----:|
|Democratic|81|500|
|Republican|36|412|

<br>

Use `proptestGC()` to perform the two-sided test, from the summary data:

* What should you enter for `x`?
* What should you enter for `n`?
* What should you enter for `p`?
* Can you conclude that Democrats are more likely than Republicans to believe in astrology?

## Thought Question

So there appears to be a relationship, in the U.S. population, between party affiliation and belief in astrology.

* Does your party affliation affect what you believe about astrology?
* Do you choose your party partly on the basis of your beliefs about astrology?
* Or is there a confounding factor at work, here?

## Love at First Sight


* At Georgetown College, who is more likely to believe in love at first sight:  a female or a male?

## Descriptive Work (From Sample)


```{r}
SexLove <- xtabs(~sex+love_first,data=m111survey)
SexLove
rowPerc(SexLove)
```

  
## Inferential Work


Let

* $p_1$ = the proportion of all GC females who believe in love at first sight

* $p_2$ = the proportion of all GC males who believe in love at first sight

Then our hypotheses are:

* $H_0$:  $p_1 - p_2 = 0$

* $H_a$:  $p_1 - p_2 \neq 0$
  
## The Code


```{r eval=F}
proptestGC(~sex+love_first,data=m111survey,
        success="yes",p=0,graph=TRUE)
```

Note same old formula for studying relationship between two factor variables:

$$\sim Explanatory + Response$$

## Descriptive Results


```
##        yes  n estimated.prop
## female  18 40         0.4500
## male     8 31         0.2581
```

Notice:

* fewer than 10 successes for the males.
* `proptestGC()` delivers a warning

## What to Do?


We are dealing with two factor variables, so could use:

```{r eval=FALSE}
chisqtestGC(~sex+love_first,data=m111survey,
        simulate.p.value="random",
        B=3000,graph=TRUE)
```

* this gives us a test of significance
* but (sadly) no confidence interval for $p_1-p_2$

## Group Order in proptestGC()


If you want to be sure that a certain group comes first, then use the argument `first`:

For example:

```{r eval=FALSE}
proptestGC(~sex+extra_life,data=m111survey,
        first="male",
        success="yes",p=0)
```

## Want Less Output?


Just set `verbose` to `FALSE`:

```{r eval=FALSE}
proptestGC(~sex+extra_life,data=m111survey,
        first="male",
        success="yes",p=0,
        verbose=FALSE)
```