---
title: "Tips and Tricks"
author: "Homer White"
date: "December, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---


```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
library(abd)
```


# Getting Data Into R

## Importing Data (on Server somewhere)

Use Import Dataset:

* From Text File ...

Example:  `m111surveyFA13.csv`

## Importing Data From the Web

Use Import Dataset:

* From Web URL ...

Example:

* Go to [http://harvardforest.fas.harvard.edu:8080/exist/xquery/data.xq?id=hf124](http://harvardforest.fas.harvard.edu:8080/exist/xquery/data.xq?id=hf124)
* find the data link
* right-click and copy the link location
* paste it into dialog box

## Importing into an R Markdown Document

use `read.csv()`:

```{r eval=FALSE}
newSurvey <- read.csv("/mat111/Additional_Datasets/m111surveyFA13.csv",header=TRUE)
drs<-read.csv("http://harvardforest.fas.harvard.edu/data/p12/hf124/hf124-01-browse.csv",
                  header=TRUE)
```

## Uploading Data

* In the Files pane, find the Upload button
* press it, then find the file on your computer

Example:

* I sent you an email with a csv file attached
* download the file onto your computer
* upload it into your Home directory on the Server
* import it

# Sending Data to Others

## Get It Off the Server

* In the Files pane, check the file you want to send
* Find the More button
* choose Export
* the file will go to the Downloads file on your computer
* save it where you wish
* or send it as an email attachment, etc.


# Handling Excel Files

## Excel Can be Messy!

* Download and open the Excel Workbook that I sent you by email.  Note the messiness.

* Save it somewhere on your computer.

* Upload it to the Server, to your Home directory.

## The readxl Package

This package is great for reading parts of excel files:

```{r eval=FALSE}
require(readxl)
```

## Reading a specific worksheet

Say that you want the data on the second sheet.  Then run:

```{r eval=FALSE}
carinfo <- read_excel("datasets.xlsx",sheet=2)
View(carinfo)
```


## Skipping lines

In the first sheet, the first two lines are junk.  We can skip them:

```{r eval=FALSE}
irisinfo <- read_excel("datasets.xlsx",sheet=1,skip=2)
```

## A List of Sheets

If you wnat a reminder about which sheet is which, you can get a list of sheet names, in order:

```{r eval=FALSE}
excel_sheets("datasets.xlsx")
```

```
[1] "iris"     "mtcars"   "chickwts" "quakes"  
```









