---
title: "Two Numeric Variables: Pt.1"
author: "Homer White"
date: "August, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(mosaicData)
library(tigerstats)
```

# Graphical Methods

## Swimming World Records

```{r eval=FALSE}
library(mosaicData)
data(SwimRecords)
View(SwimRecords)
help(SwimRecords)
```

**Research Question**:  How have 100m records varied over the years?

## Variable Analysis

Variables involved:

* **year** explanatory
* **time** response

Their types:

* **year** numerical
* **time** numerical

## Are Old Methods Useful?

With numerical variables we have used:

* `densityplot()`
* `bwplot()`
* `histogram()`
* `favstats()`

Try this:

```{r eval=FALSE}
histogram(~time|year,data=SwimRecords)
```


## Weirdness

```{r echo=FALSE}
histogram(~time|year,data=SwimRecords)
```


## The Problem

* explanatory variable **year** is numerical, not a factor, and
* it has LOTS of values!

We need a new graphical method to study the relationship between two numerical variables.

## A Scatter Plot

```{r echo=FALSE}
xyplot(time~year,data=SwimRecords,
       main="100m Swim Records",
       xlab="year",
       ylab="time (seconds)",
       pch=19)
```


## How to Make a Scatter Plot

Use the `xyplot()` function:

```{r eval=FALSE}
xyplot(time~year,data=SwimRecords,
       main="100m Swim Records",
       xlab="year",
       ylab="time (seconds)",
       pch=19)
```

Basic Form:

$$xyplot(response \sim explanatory, data = \ldots)$$

# Relationship Terminology

## Positive Linear

```{r echo=FALSE}
xyplot(height~father,data=Galton,
       subset=sex=="M",
       main="Father and Son Heights",
       xlab="father's height (inches)",
       ylab="son's height (inches)",
       pch=19,type=c("p","r"))
```

Rising cloud:  the taller the father, the taller the son!

## Negative Linear

```{r echo=FALSE}
set.seed(2040)
x <- rnorm(100,mean=50,sd=10)
y <- -2*x+rnorm(100,sd=5)
xyplot(y~x,pch=19,type=c("p","r"))
```

Falling cloud:  bigger x's go with smaller y's!

## Curvilinear

```{r echo=FALSE}
set.seed(2040)
x <- rnorm(100,mean=50,sd=10)
y <- (x-50)^2+rnorm(100,sd=20)
xyplot(y~x,pch=19)
```

There's a relationship, but it's not linear!

## A Bit Curvilinear

```{r echo=FALSE}
xyplot(gasbill~temp,data=Utilities,
       main="Temperature and Gas Bill",
       xlab="temperature (deg Fahrenheit)",
       ylab="gas bill (dollars)",
       pch=19)
```

See the cloud "leveling off"?

## No Relationship

```{r echo=FALSE}
set.seed(3010)
x <- rnorm(100,mean=50,sd=10)
y <- rnorm(100,mean=50,sd=10)
xyplot(y~x,pch=19)
```

As you move to the right, cloud neither rises nor falls.

# Examples

## Beetles and Stumps

```{r eval=FALSE}
data(stumps)
View(stumps)
help(stumps)
```

**Research Question**:  Are there more larvae cluster in plots where there are more cottonwood stumps?

## Practice {.practice}

* **stumps** numerical
* **larvae** numerical

Which is explanatory?

Which is the response?

Make the scatter plot.

Describe the relationship between number of larvae and number of stumps.

## Speed and Fuel Efficiency

```{r eval=FALSE}
data(fuel)
View(fuel)
help(fuel)
```

Research Question:  How does the speed at which the Ford Escort is driven affect its fuel efficiency?

## Practice { .practice }

* **speed** numerical
* **efficiency** numerical

Which is explanatory? Which is the response?

Make the scatter plot.

Describe the relationship between number of speed and fuel efficiency.

Units for **efficiency** are "liters per 100 km".  Does a high number for **efficiency** represent good or poor fuel efficiency?

# Groups in Scatter Plots

## Dealing With Groups: 1

Use a `groups` argument, with a key:

```{r eval=FALSE}
xyplot(time~year,data=SwimRecords,
       main="100m Swim Records",
       xlab="year",
       ylab="time (seconds)",
       pch=19,
       groups=sex,
       auto.key=TRUE)
```

## The Result


```{r echo=FALSE}
xyplot(time~year,data=SwimRecords,
       main="100m Swim Records",
       xlab="year",
       ylab="time (seconds)",
       pch=19,
       groups=sex,
       auto.key=TRUE)
```

## Dealing with Groups:  2

```{r eval=FALSE}
data(TenMileRace)
View(TenMileRace)
help(TenMileRace)
```

Let's look at the relationship between **age** (explanatory) and **net** time (response), for both men and women.

## First Pass

```{r eval=FALSE}
xyplot(net~age,data=TenMileRace,
       groups=sex,
       auto.key=TRUE)
```

## Sexes are "Too Mixed Up"

```{r echo=FALSE}
xyplot(net~age,data=TenMileRace,
       groups=sex,
       auto.key=TRUE)
```


## Plot in Separate Panels

Make a scatter plot for each sex:

```{r eval=FALSE}
xyplot(net~age|sex,data=TenMileRace,
       main="Age and Race Time, by Sex",
       xlab="age (years)",
       ylab="10k time (seconds)",
       layout=c(2,1))
```

## The Result

```{r echo=FALSE}
xyplot(net~age|sex,data=TenMileRace,
       main="Age and Race Time, by Sex",
       xlab="age (years)",
       ylab="10k time (seconds)",
       layout=c(2,1))
```

The guys appear to run faster.  Hard to tell if age make a difference.

## Verlander Over the Plate

```{r eval=FALSE}
help(verlander)
```


* **px** horizontal distance of ball from center of plate (feet)
* **pz** vertical distance of ball above plate (feet)

**Research Question**:  How are these variables related?

## First Try

```{r eval=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19)
```

## Result


```{r echo=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19)
```


Over-plotting! (15,307 points sitting on top of each other.)

## Second Try

```{r eval=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10)
```

With `alpha` set to 0.10, you need 10 points stacked on each other to get full color!

## Result

```{r echo=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10)
```

## Third Try

Let's group by type of pitch:

```{r eval=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10,
       groups=pitch_type,
       auto.key=list(space="right"))
```

## The Result

```{r echo=FALSE}
xyplot(pz~px,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10,
       groups=pitch_type,
       auto.key=list(space="right"))
```

Hard to tell!

## Fourth Try

Let's plot in separate panels for each pitch type:

```{r eval=FALSE}
xyplot(pz~px|pitch_type,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10)
```

## The Result

```{r echo=FALSE}
xyplot(pz~px|pitch_type,data=verlander,
       main="Verlander over the Plate",
       pch=19,
       alpha=0.10)
```

## Practice { .practice }

For which types of pitch is there a relationship between **px** and **pz**?

How would you describe the relationships you see (if any)?