---
title: "Two Factor Variables: Pt.2"
author: "Homer White"
date: "Saturday, August 30, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```



# The Chi-Square Test

## Inferential Procedure

Suppose we are studying the relationship between two factor variables.

We might find a relationship in our sample.

**Inferetial Question**:  Did it happen by chance, or do we have good evidence that there is a relationship in the population?

## Chi-Square Test

We can answer that question using the *chi-square test*.

There are "steps" in our reasoning process.

## Step One:  State Hypotheses

(Say we are studying sex and seating preference at GC.)

Null Hypothesis:

$H_0:$  Sex and seating preference are not related in the GC population.

Alternative Hypothesis:

$H_a:$  Sex and seating preference ARE related in the GC population.

## Step Two:  The Code

```{r eval=FALSE}
chisqtestGC(~sex+seat,data=m111survey,graph=TRUE)
```

Give it a try!

##  Some Output

The two-way table comes "for free":

```
Observed Counts:
        seat
sex      1_front 2_middle 3_back
  female      19       16      5
  male         8       16      7
```
  
## More Output

Expected counts also come "for free":

```
Counts Expected by Null:
        seat
sex      1_front 2_middle 3_back
  female   15.21    18.03   6.76
  male     11.79    13.97   5.24
```

The more difference there is between the observed and the expected counts, the worse it looks for the Null hypothesis!

## More Output

```
Contributions to the chi-square statistic:
        seat
sex      1_front 2_middle 3_back
  female    0.94     0.23   0.46
  male      1.22     0.29   0.59

Chi-Square Statistic = 3.734 
```

The "contributions" add up to the chi-square statistic.  The bigger it is, the more the observed results differ from what the Null expected to see!

## More Output

```
Degrees of Freedom of the table = 2 
```

If the Null is correct, then the chi-square statistic should work out to around 2, give or take some for chance error.

## Step Three

In Step Three of our reasoning, we report the test statistic and the P-value:

The test statistic was 3.734.

The machine says that the P-value is 0.1546.

## About the P-Value

* It's based on a mathematical approximation method.
* Before we could make computers do simulations, it's all people had to go by!

**Interpretation** (always try to say this): "If sex and seating preference are not related, then there is about a 15.46% chance of getting a chi-square statistic at least as big as the one we got in our study."

## Graph of the P-value

```{r echo=FALSE}
invisible(pchisqGC(3.734,region="above",df=2,graph=T))
```

The curve approximates what you would get using simulation.

## Step Four:  Decision

Since the P-value is fairly large, we won't reject the Null Hypothesis.

People often use 0.05 as a "cut-off" value:

* if $P < 0.05$, they reject $H_0$
* If $P \ge 0.05$, they don't reject $H_0$ (though they might be suspicious about it).

## Step Five:  Conclusion

We say how much evidence the study provided for $H_a$:

"This study did not provide very strong evidence that sex and seating preference are related in the GC population."

# More Examples

## Love at First sight

Research Question:

>*In the GC population, do the two sexes differ in their beliefs about love at first sight?*

Null Hypothesis:

$H_0:$  Sex and belief in love at first sight are not related in the GC population.

Alternative Hypothesis:

$H_a:$  These two variables ARE related in the GC population.

## Practice { .practice }

Try it out yourself:

```{r eval=FALSE}
chisqtestGC(~sex+love_first,data=m111survey,
            graph=TRUE)
```

What's the test statistic?

About how big should it be if the Null is correct?

What's the P-value?

Do you think we have strong evidence for a relationship in the population, or could the pattern in the data be due just to chance?

## Race and Guns

The General Social Survey, from the year 2002:

```{r eval=FALSE}
data(gss02)
View(gss02)
help(gss02)
```

Research Question:

>*In the U.S. population, do the races differ in how likely they are to own a gun?*

Variable analysis:

* **race** (factor, explanatory)
* **owngun** (factor, response)

## Look at the data

```{r}
raceGun <- xtabs(~race+owngun,data=gss02); raceGun
rowPerc(raceGun)
```


## Hypotheses

Null Hypothesis:

$H_0:$  Race and gun ownership are not related in the US population.

Alternative Hypothesis:

$H_a:$  Race and gun ownership ARE related in the US population.

## Practice { .practice }

Try it yourself!  run the code needed for a chi-square test.

What's the test statistic?

About how big should it be if the Null is correct?

What's the P-value?

Do you think we have strong evidence for a relationship in the population, or could the pattern in the data be due just to chance?

# Simulation

## Ledge-Jumping

A study (Great Britain, 1970's) of people who threatened to commit suicide by leaping off of a high structure.  Usually a crowd would gather to watch.

```{r eval=FALSE}
data(ledgejump)
View(ledgejump)
help(ledgejump)
```

Research Question:  *Is there a relationship between the weather and the behavior of the crowd?*

## Descriptive Work

```{r}
weBe <- xtabs(~weather+crowd.behavior,data=ledgejump)
weBe
```


```{r}
rowPerc(weBe)
```

Looks like the crowd baited more often in warm weather!

## Inferential Question

But is this due to a relationship between weather and crowd behavior?

Or could it be due just to chance variation in all of the other factors that affect crowd behavior?

## Hypotheses

Null Hypothesis:

$H_0:$  Weather and crowd behavior are not related.

Alternative Hypothesis:

$H_a:$  Weather and crowd behavior ARE related.

## The Code

```{r eval=FALSE}
chisqtestGC(~weather+crowd.behavior,data=ledgejump)
```

<br>

Uh-oh! There is a **warning**!

## The Problem

If the expected cell counts are not big enough ... 

... then the math approximation for the P-value may not be reliable!

(each cell  > 5 is usually OK)

## The Solution

We can always fall back on simulation!

```{r eval=FALSE}
chisqtestGC(~weather+crowd.behavior,data=ledgejump,
            simulate.p.value="fixed",
            B=3000)
```

## Setting a Seed

In data analysis reports, it's useful to get the same P-value whenever you perform the test.  You can accomplish this with the `set.seed()` function:

```{r eval=FALSE}
set.seed(2020)  #Just pick a number
chisqtestGC(~weather+crowd.behavior,data=ledgejump,
            simulate.p.value="fixed",
            B=3000)
```

Run the code chunk several times.  Does the P-value change?


## Practice { .practice }

Try simulation on the sex and seating-preference study:

```{r eval=FALSE}
chisqtestGC(~sex+seat,data=m111survey,
            simulate.p.value="random",
            B=3000)
```

Compare the P-value to the P-value you got earlier, without using simulation:  are they about the same, or very different?


# (Optional) Types of Simulation

## "Random Sampling" Type

Suppose your data came from a random sample you took from some larger population.  Then when you simulate, you might want to use the "random" option:

```{r eval=FALSE}
chisqtestGC(~sex+seat,data=m111survey,
            simulate.p.value="random",
            B=3000,graph=TRUE)
```


(Students were a random sample from the population.  In the re-sampling, the number of guys and gals in the re-sampling will NOT be fixed at 31 and 40.)

## "Fixed" Type

Sometimes the tally for the explanatory variable is fixed.  Then when you simulate, you might want to use the "fixed" option:

```{r eval=FALSE}
chisqtestGC(~weather+crowd.behavior,data=ledgejump,
            simulate.p.value="fixed",
            B=3000,graph=TRUE)
```

(The 21 incidents were not a random sample from a larger population.  There were 9 in cold weather and 12 in warm weather, and that's that.  Chance was at play only in how each incident "turned out".)

## "Double-fixed" Type

In this type of random simulation, the tallies are fixed for both the explanatory and response variables.

`simulate.p.value=TRUE` gives you this type of re-sampling.

This type of simulation is ideal for randomized experiment (see Chapter 6).

But you can use it when you aren't sure what type of simulation to do.

