---
title: "Two Factor Variables: Pt.1"
author: "Homer White"
date: "Saturday, August 30, 2014"
output:
  ioslides_presentation:
    incremental: no
    mathjax: default
    widescreen: yes
    standalone:  true
    css:  slide_styles.css
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Explanatory and Response

## Two Variables

Often a Research Question involves two variables.

Example:  For the `m111survey` data, we might ask:

>*Do males and females differ in their seating preferences?*

Two variables involved:

* **sex** (a factor variable)
* **seat** (a factor variable)


## Explanatory and Response

Sometimes we consider one of the variable to be *explanaotry*, the other to the be the *response* variable.

The explanatory variable is one that

* we think might cause or explain the values of the response, OR
* be used to predict the response

Maybe here we think:

* **sex** is explanatory
* **seat** is the response

# Two Factor Varibles:  Methods

## Methods We Use

To study the relationship between two factor variables, we can make:

* two-way tables (`xtabs()`)
* bar charts (`barchartGC()`)

Code:

$$function(\sim explanatory + response, data = \ldots)$$

## Two-Way Table

```{r}
sexSeat <- xtabs(~sex+seat,data=m111survey)
sexSeat
```

There are:

* $19+16+5=40$ females
* $8+16+7=31$ males

## Row Percents

To correct for the differing group sizes, compute row percents:

```{r}
19/40
```

47.5% of the females preferred the front.

```{r}
8/31
```

25.8% of the males preferred the front.

## Practice:  Row Percents { .practice }

```{r echo=FALSE}
sexSeat
```

What percentage of females preferred the back?

What percentage of males preferred the back?

## Using rowPerc()

```{r}
rowPerc(sexSeat)
```

## Describing the Relationship

```{r echo=FALSE}
rowPerc(sexSeat)
```

"In the sample, sex and seating preference are related:  for example, females were more likely to prefer the front than males were (47.5% vs. 25.81%)."

<br>

## Describing the Relationship

* Make row percents
* Compare row percentages down columns (*not across a row!!*)
* If you see one or more big differences, then:
    * the two variables are related
    * in other words, the groups differ in their response
    
## Another Example

In the `mat111survey` data:

>*Who is more likely to believe in love at first sight:  a guy or a gal?*

Variable Analysis

* **sex** (factor, explanatory)
* **love_first** (factor, response)

We can make a two-way table and some row percents to study the Research Question!

## Practice { .practice }

Make the two-way table.

Make row percents.

Describe the relationship (if any) between the two variables.

# Inference

## From Sample to Population

In the sample, we saw that **sex** and **seat** were related:  males were less likely than females to prefer the front, more likely than females to prefer the back.

But are these two variables related *in the GC population*?

After all, the sample was picked randomly from the population.

Could the pattern we saw in the sample be due just to chance?  Or is it so strong that we have to conclude there is a relationship in the population?

## Back to Buzz

In Dr. Bastian's experiment, we saw that Buzz pushed the correct button 15 out of 16 times.

But could those results be due just to chance variation in the guessing process?  Or are they so extreme that we have to conclude that he can communicate with Doris?

## Two Hypotheses

Null Hypothesis:

$H_0:$  Sex and seating preference are not related in the GC population.

Alternative Hypothesis:

$H_a:$  Sex and seating preference ARE related in the GC population.

## Back to Buzz

Null Hypothesis:

$H_0:$  Buzz's chance of picking correct door is 50%. (No communication with Doris.)

Alternative Hypothesis:

$H_a:$  Buzz's chance of picking correct door is MORE THAN 50%. (Some communication with Doris.)

## Attacking the Dolphin Question

Null Hypothesis expects 8 correct picks (give or take some for chance error).

Buzz got 15.  (This is our *test statistic*)

We found P-value (chance of 15 or more correct if Null Hypothesis is correct).

The P-value helped us decide if we had strong evidence that Buzz could communicate.

## Attacking Sex and Seat

Find what sort of two-way table the Null Hypothesis expects.

Find a number that measures how different the actual table is from what the Null expected.  (This will be our test statistic.)

Find the P-value (chance of getting a test statistic at least as big as the one we got, if the Null is correct.)

Use the P-value to decide if we have strong evidence for a relationship between **sex** and **seat** in the GC population.

## Expected Counts

Observed table:

```{r echo=FALSE}
sexSeat
```

Null expects to see a table like this:

```{r}
expCounts(sexSeat)
```

Give or take some for chance error in the sampling process!

## Expected Counts

The Null says there is no relationship in the population.

So the Null expects essentially no relationship in the sample.

```{r}
rowPerc(expCounts(sexSeat))
```

(Well, it expects a small relationship, because of chance variation in the sampling process)

## Test Statistic

We need a test statistic that will measure how much the observed table differs from the one the Null expects.

$$\chi^2=\sum_{\text{all cells}} \frac{(\text{Observed}-\text{Expected})^2}{\text{Expected}}.$$

## Chi-Square Statistic

This works out to:

$$\chi^2 = \frac{(19-15.21)^2}{15.21}+\frac{(8-11.79)^2}{11.79} \\
+\frac{(16-18.02)^2}{18.02}+\frac{(16-13.97)^2}{3.97} \\
+\frac{(5-6.76)^2}{6.76}+\frac{(7-5.24)^2}{5.24} \\
\approx 3.73$$

```{r echo=FALSE}
expected <- expCounts(sexSeat)
chisqStat <- round(sum((sexSeat-expected)^2/expected),2)
```


(We'll always make R compute this number.)

## P-value

Is our Chi-square statistic an unusually big number?

What's the chance of getting 3.73 or more, if the Null is correct?

## Back to Buzz

To find the P-value for Buzz, we simulated.

We imagined that Buzz was just guessing.  Then we had him repeat the experiment many, many times.

## Simulation

We can do the same thing!

* Imagine that there is no relationship between sex and seating preference in a big population.
* Take a random sample from that imaginary population many, many times.
* Make a two-way table of our results.
* Compute the chi-square statistic.
* Keep track of how often that statistic exceeds the 3.73 that we actually got in our study.



## Repeat Many Times

[http://homer.shinyapps.io/SlowAssoc](http://homer.shinyapps.io/SlowAssoc)

## Now Decide

Think about your simulations.

Is is all that unlikely to get 3.73 or more?

Did the `m111survey` study provide strong evidence for a relationship between sex and seating preference?

## Practice { .practice }

```{r eval=FALSE}
ChisqSimSlow(~sex+love_first,data=m111survey)
```

How many times did you re-sample?

What percentage of the time did the re sampled chi-square statistic exceed the chi-square statistic in the actual study?

Do you think there is overwhelming evidence that in the GC population the two sexes differ in whether they believe in love at first sight?
