---
title: "Moodle Questions"
author: "Rebekah Robinson"
date: "Tuesday, June 17, 2014"
output: word_document
---

```{r include=FALSE}
require(knitr)
require(vcd)
require(tigerstats)
```


PART 1
# Baseball Scatterplot Code

Before starting this question, make sure to require the necessary package and load the appropriate dataset.  Run the following code.

```{r eval=FALSE}
require(vcd)
data(Baseball)
View(Baseball)
help(Baseball)
```

This data frame contains observations for 322 baseball players on 25 variables using data from 1986.  **At bats**, `atbat86`, count the number of official plate appearances by a hitter.   A **hit**, `hits86`, is credited to a hitter is they successfully make it to first base after they hit the ball into fair territory.  The scatterplot is shown below.

```{r,echo=FALSE}
xyplot(hits86~atbat86,data=Baseball,main="At Bats vs. Hits in 1986", ylab="1986 Player Hits",xlab="1986 Player At Bats",pch=20,col="blue")
```

Which of the following lines of code created this scatterplot?

```{r eval=FALSE}
xyplot(hits86~atbat86, data=Baseball)
```

```{r eval=FALSE}
xyplot(atbat86~hits86,data=Baseball)
```

```{r eval=FALSE}
xyplot(hits86~atbat86, groups=league86, data=Baseball, auto.key=TRUE)
```

```{r eval=FALSE}
xyplot(atbat86~hits86, groups=league86, data=Baseball, auto.key=TRUE)
```

```{r eval=FALSE}
xyplot(hits86~atbat86|league86, data=Baseball)
```

```{r eval=FALSE}
xyplot(atbat86~hits86|league86, data=Baseball)
```


Just by looking at the scatterplot, which of the following is the *best* description of the association?

# Association Matching


Match each of the following scatterplots with the best description of the type of association between the variables.

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=0
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=0.8
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=-0.6
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
x<-seq(0,20,by=0.1)
y<-dexp(x,rate=0.5)
realy<-y++rnorm(201,0,0.03)
plot(x,realy,xlim=c(0,20),ylim=c(-0.1,0.5),pch=19,col="blue", cex=0.5,ylab="y")
```

# KidsFeet Scatterplot Grouped by Sex

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children was the length, (*length*), and width, (*width*), of their longest foot.  Based on the scatterplot shown below, there appears to be a positive linear association between the length and width of a fourth grader's biggest foot that holds even when the child's sex is accounted for.  Which of the following lines of code created this scatterplot?

```{r echo=FALSE}
xyplot(width~length,groups=sex,data=KidsFeet,auto.key=TRUE,pch=19,xlab="Length of Biggest Foot (in cm)", ylab="Width of Biggest Foot (in cm)",main="Length vs. Width of Biggest Foot")
```


```{r eval=FALSE}
xyplot(width~length,groups=sex,data=KidsFeet, auto.key=TRUE)
xyplot(length~width, groups=sex, data=KidsFeet, auto.key=TRUE)
xyplot(width~length|sex,data=KidsFeet)
xyplot(length~width|sex, data=KidsFeet)
xyplot(width~length,groups=sex,data=KidsFeet)
xyplot(length~width, groups=sex, data=KidsFeet)
xyplot(width~length, data=KidsFeet)
xyplot(length~width, data=KidsFeet)
```

# KidsFeet Parallel Scatterplots by Sex

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children was the length, (*length*), and width, (*width*), of their longest foot.  Based on the scatterplot shown below, there appears to be a positive linear association between the length and width of a fourth grader's biggest foot that holds even when the child's sex is accounted for.  Which of the following lines of code created this scatterplot?

```{r echo=FALSE}
xyplot(width~length|sex,data=KidsFeet,pch=19,xlab="Length of Biggest Foot (in cm)", ylab="Width of Biggest Foot (in cm)",main="Length vs. Width of Biggest Foot")
```


```{r eval=FALSE}
xyplot(width~length,groups=sex,data=KidsFeet, auto.key=TRUE)
xyplot(length~width, groups=sex, data=KidsFeet, auto.key=TRUE)
xyplot(width~length|sex,data=KidsFeet)
xyplot(length~width|sex, data=KidsFeet)
xyplot(width~length,groups=sex,data=KidsFeet)
xyplot(length~width, groups=sex, data=KidsFeet)
xyplot(width~length, data=KidsFeet)
xyplot(length~width, data=KidsFeet)
```



PART 2
# Baseball Correlation

Before starting this question, make sure to require the necessary package and load the appropriate dataset.  Run the following code.

```{r eval=FALSE}
require(vcd)
data(Baseball)
View(Baseball)
help(Baseball)
```

This data frame contains observations for 322 baseball players on 25 variables using data from 1986.  **At bats**, `atbat86`, count the number of official plate appearances by a hitter.   A **hit**, `hits86`, is credited to a hitter is they successfully make it to first base after they hit the ball into fair territory.  The scatterplot is shown below.

```{r,echo=FALSE}
xyplot(hits86~atbat86,data=Baseball,main="At Bats vs. Hits in 1986", ylab="1986 Player Hits",xlab="1986 Player At Bats",pch=20,col="blue")
```

Just by looking at the scatterplot, which of the following is the most likely value for the correlation coefficient, $r$?

+1
0
-1
+0.25
+1.8
-0.009



# Correlation Coefficient Matching

Match the following scatterplots with it's most likely value for the correlation coefficient, $r$.

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=-0.2
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```


```{r echo=FALSE, message=FALSE}
require(MASS)
rho=+0.6
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=0.008
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=-0.95
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=1
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```

```{r echo=FALSE, message=FALSE}
require(MASS)
rho=-1
varcovar<-cbind(c(1,rho),c(rho,1))
rpoints<-mvrnorm(n=500,mu=c(0,0),Sigma=varcovar)
r<-round(cor(rpoints)[1,2],3)
x<-rpoints[,1]
y<-rpoints[,2]
plot(x,y,cex=0.5,col="blue",pch=19)
```


# KidsFeet Regression Analysis Code

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

Which of the following lines of code resulted in the following plot and regression analysis?  **Hint:  More than one choice may be correct.**

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet,graph=TRUE)
```

```{r eval=FALSE}
lmGC(width~length,data=KidsFeet,graph=TRUE)
lmGC(width~length,data=KidsFeet)
lmGC(length~width,data=KidsFeet,graph=TRUE)
lmGC(length~width,data=KidsFeet)

KidsFeetMod<-lmGC(length~width,data=KidsFeet,graph=TRUE)
KidsFeetMod

KidsFeetMod<-lmGC(width~length,data=KidsFeet,graph=TRUE)
KidsFeetMod

KidsFeetMod<-lmGC(length~width,data=KidsFeet)
KidsFeetMod

KidsFeetMod<-lmGC(width~length,data=KidsFeet)
KidsFeetMod
```

# KidsFeet Regression Line Slope Interpretation

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question*: How is the width of a fourth-grade student's foot related to the length?

Which of the following sentences interprets the **slope** of the regression line in terms of how the length of a student's foot changes as the width increases?

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

For every 1 centimeter increase in the width of a fourth-grader's foot, the predicted length increases by 1.6576 centimeters.

For every 1 inch increase in the width of a fourth-grader's foot, the predicted length increases by 1.6576 inches.

For every 1 inch increase in the length of a fourth-grader's foot, the predicted width increases by 1.6576 inches.

For every 1 centimeter increase in the width of a fourth-grader's foot, the predicted length decreases by 1.6576 centimeters.

For every 1 inch increase in the width of a fourth-grader's foot, the predicted length increases by 9.8172 inches.

# KidsFeet Regression Line Intercept

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

Which of the following is the $y$-intercept of the regression line?

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

9.8172
0.6411
1.6576
1.0248
0.411


# KidsFeet Predictions

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

Which of the following is the $y$-intercept of the regression line?

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

Use the following regression analysis to answer the following question:


What is the predicted foot length (in centimeters) for a fourth grader with a foot that is 8.8 centimeters wide?  Answers are all rounded to the nearest tenth.

```{r echo=FALSE, fig.height=5,fig.width=5}
mymod<-lmGC(length~width,data=KidsFeet)
mymod
```

24.4
24.8
-4.8
9.8
24.2

# KidsFeet Residuals

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

Allison is a fourth-grade student.  Suppose the width of a Allison's dominant foot is 9.25 centimeters wide.  Run the following lines of code below to see what the regression line predicts the length of her foot to be, in centimeters.  

```{r eval=FALSE}
KidsFeetMod<-lmGC(length~width,data=KidsFeet)
predict(KidsFeetMod,9.25)
```

Suppose that Allison's foot is really 26.5 centimeters long.  Which of the following numbers is the **residual** difference between the actual length and the predicted length of Allison's foot?

1.3497
-1.3497
17
-17
15.6503
-15.6503



# Explanatory and Response Matching

For the following research question, which variable should be the explanatory variable and which should be the response variable?

*Research Question:*  For MAT 111, how are midterm exam scores related to final exam scores?

# KidsFeet Squared Correlation

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

The $r^2$ statistic for these data is 0.411.  Which of the following sentences is the correct interpretation of this statistic?

41.1% of the children in the study have wider than average feet.

41.1% of the variation in the length of children's feet is explained by the linear relationship between the width and the length of a child's foot.

41.1% of the variation in the width of children's feet is explained by the linear relationship between the width and the length of a child's foot.

For every additional centimeter wide a student's foot is, the length of their foot increases by 0.411.

If the width of a child's dominant foot is 0 centimeters, the predicted length would be 0.411 centimeters.

# Regression Analysis Matching

A researcher gathered data on 13 college-aged males chugging a 12-ounce can of a certain beverage.  The researcher recorded the weight (**Weight**), in pounds, and the chugtime (**ChugTime**), in seconds, of each subject.  

The regression analysis, along with a plot, is shown below.

```{r echo=FALSE}
lmGC(ChugTime~Weight,data=chugtime,graph=TRUE)
```

Match each of the following numbers with their correct description.

-0.7879
The numerical description of the strong negative relationship between a college male's weight and the amount of time it takes for him to chug a 12 ounce beverage.  

13.3
A college-aged male weighing 0 pounds is predicted to take 13.2975 seconds to chug a 12 ounce beverage.

-0.0458
For every one pound increase in the weight of a college male, the predicted chug time of a 12 ounce beverage decreases by 0.0458 seconds.

1.123
A measurement of the variation in the data from the regression line that is affected by a change in scale of the response variable.

0.6208
The proportion of variation in the chug times that is explained by the weights of the college males.  


PART 3

# KidsFeet Error 1

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?

```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

True or False?  Sam is a senior in high school whose dominant foot measures 10.8 centimeters wide.  The regression line shown above could be used to make a prediction about how long Sam's foot is.

# KidsFeet Error 2

True or False?  For school children, foot size is correlated with reading skills.  Thus, learning new words makes children's feet grow bigger. 

# KidsFeet Error Matching

The *KidsFeet* dataframe contains data collected on 39 fourth grade students in Ann Arbor, MI, in October 1997.  Two of the measurements taken on the children were the length in centimeters, (*length*), and width in centimeters, (*width*), of their longest foot.  This data could be used to answer the following

*Research Question:*  How is the width of a fourth-grade student's foot related to the length?


```{r echo=FALSE, fig.height=5,fig.width=5}
lmGC(length~width,data=KidsFeet)
```

Match the following statements with the *type of error* they commit.

Extrapolation:  Sam is a senior in high school whose dominant foot measures 10.8 centimeters wide.  The regression line shown above could be used to make a prediction about how long Sam's foot is.

Association does not imply causation:  For school children, foot size is correlated with reading skills.  Thus, learning new words makes children's feet grow bigger. 

# SAT Misinterpretations

Data was gathered on the SAT (a college entrance exam) for each of the fifty states in the U.S.  A research was interested in answering the following 

*Research Question:*  How is a state's average annual teacher salary, in thousands of dollars, related to that state's average student performance on the math portion of the SAT?

Study the following regression analysis.

```{r echo=FALSE}
lmGC(math~salary,data=sat,graph=TRUE)
```

Which of the following statements is an **incorrect** interpretation of the analysis?  There may be more than one correct answer.

The correlation between average annual teacher salary and average SAT math subscore is -0.4013. (Correct)

Average annual teacher salary and average SAT math subscore are negatively correlated. (Correct)

States that pay lower teacher salaries tend to produce students to score higher on the math portion of the SAT.  (Correct)

Paying teachers lower salaries causes students to perform better on the math portion of the SAT. (Incorrect)

If a state that currently pays an average teacher salary of $40,000 per year would raise teacher pay by $10,000 next year, the average math subscore on the SAT would go down by 27 points. (Incorrect)

This analysis provides evidence that paying teachers $100,000 per year would be detrimental to student's math ability.  An average teacher salary of $100,000 would result in an average math subscore of 332 on the SAT. (Incorrect)

# Influential Observation

Which of the following placement of the red point in the scatterplots below would have the most influence over the regression line?

```{r echo=FALSE}
offloc <- 5
middle <- c(0,0)
offside <- c(offloc,0)
rho <- -0.7
n<- 100
big <- offloc
xmax <- big+1
xmin <- -xmax
ymax<- big+1
ymin <- -ymax
varcovar <- cbind(c(1,rho),c(rho,1))
testval <- offloc+1

while (testval>offloc)  {
  rpoints <- MASS::mvrnorm(n=n,mu=c(0,0),Sigma=varcovar)
  x <- rpoints[,1]
  y <- rpoints[,2]
  testval <- max(abs(c(x,y)))        
  }

plot(rpoints,pch=16,cex=0.5,
    xlim=c(xmin,xmax),
    ylim=c(ymin,ymax),
    xlab="x",ylab="y")
points(x=middle[1],y=middle[2],col="red",pch=16,cex=1.5)
```

```{r echo=FALSE}
offloc <- 5
middle <- c(-5,-5)
rho <- -0.7
n<- 100
big <- offloc
xmax <- big+1
xmin <- -xmax
ymax<- big+1
ymin <- -ymax
varcovar <- cbind(c(1,rho),c(rho,1))
testval <- offloc+1

while (testval>offloc)  {
  rpoints <- MASS::mvrnorm(n=n,mu=c(0,0),Sigma=varcovar)
  x <- rpoints[,1]
  y <- rpoints[,2]
  testval <- max(abs(c(x,y)))        
  }

plot(rpoints,pch=16,cex=0.5,
    xlim=c(xmin,xmax),
    ylim=c(ymin,ymax),
    xlab="x",ylab="y")
points(x=middle[1],y=middle[2],col="red",pch=16,cex=1.5)
```

```{r echo=FALSE}
offloc <- 5
middle <- c(0,5)
rho <- -0.7
n<- 100
big <- offloc
xmax <- big+1
xmin <- -xmax
ymax<- big+1
ymin <- -ymax
varcovar <- cbind(c(1,rho),c(rho,1))
testval <- offloc+1

while (testval>offloc)  {
  rpoints <- MASS::mvrnorm(n=n,mu=c(0,0),Sigma=varcovar)
  x <- rpoints[,1]
  y <- rpoints[,2]
  testval <- max(abs(c(x,y)))        
  }

plot(rpoints,pch=16,cex=0.5,
    xlim=c(xmin,xmax),
    ylim=c(ymin,ymax),
    xlab="x",ylab="y")
points(x=middle[1],y=middle[2],col="red",pch=16,cex=1.5)
```

# Prediction

We'll look at the `Galton` data frame from the `mosaicData` package:

```{r eval=FALSE}
data(Galton)
View(Galton)
help(Galton)
```

Here's a scatterplot of the heights of the mothers and the heights of their children:

```{r echo=FALSE}
xyplot(height~mother,data=Galton,pch=19,
       type=c("p","r"),
       xlab="Mother's height (inches)",
       ylab="Child's height (inches)")
```


We would like to use the data to predict the height of a child from the height of her mother.  As described in class and in the Course Notes, use the `predict()` function to predict the height of a child (the time of Galtons' study) if the child's mother was known to be 64 inches tall.  Record your answer (without units) as a decimal number, correct to at least two decimal places.

66.73

```{r}
motherMod <- lmGC(height~mother,data=Galton)
predict(motherMod,x=64)
```



# Give or Take

This question refers to the `Galton` data.

For the height of a child whose mother was known to be 64 inches tall, what is the prediction standard error?   Record your answer (without units) as a decimal number, correct to at least two decimal places.

3.513

# A 95%-Prediction Interval

This question refers to the `Galton` data.

Find a 95%-prediction interval the height of a child whose mother was known to be 64 inches tall.  Record the upper bound of your answer (without units) as a demcimal number, correct to at least two decimal places.

```{r}
motherMod <- lmGC(height~mother,data=Galton)
predict(motherMod,x=64,level=0.95)
```

# A Cubic Fit

The following R command will import a new data frame called `fc`:

```{r eval=FALSE}
fc <- read.csv("/mat111/Additional_Datasets/fuelConsumption.csv")
```

```{r echo=FALSE}
fc <- read.csv("../MiscDataSources/fuelConsumption.csv")
```


Here is the data:

```{r}
fc
```

The variable **fuel** gives the mean number of gallons used per car in the U.S., for each of the given years.  we would like to use the data to predcit the mean number fo gallons used per car in the year 1998, the first year not in the data.

We make and check a model with a linear fit:

```{r}
lmGC(fuel~year,data=fc,check=TRUE)
```


Then we make and check a model with a third-degree (cubic) fit:

```{r}
polyfitGC(fuel~year,data=fc,degree=3,check=TRUE)
```

Choose the code below that gives the most reliable prediction. along with an explanation of why that particular prediction is better.

Option A:

fc1 <- lmGC(fuel~year,data=fc)
predict(fc1,x=1998)

This is better because the regression line is used to predict y values from knonw x values.


Option B:

fc1 <- lmGC(year~fuel,data=fc)
predict(fc1,x=1998)

This is better because the regression line is used to predict y values from knonw x values.

Option C:

fc3 <- polyfitGCGC(fuel~year,data=fc,degree=3)
predict(fc3,x=1998)

This is better because the regression line strays outside the 95%-confidence band around the loess curve, whereas the cubic fit stays within the band.


Option D:

fc3 <- polyfitGCGC(year~fuel,data=fc,degree=3)
predict(fc3,x=1998)

This is better because the regression line strays outside the 95%-confidence band around the loess curve, whereas the cubic fit stays within the band.
