---
title: "Moodle Questions Chapter 1"
author: "Homer White"
date: "Sunday June 22, 2014"
output: word_document
---

# R as a Calculator

Which of the following bits of R-code will return the number 3, when entered into the Console?  (More than one answer may be correct:  try them all before committing to your answers!)

```{r eval=FALSE}
sqrt(9)
```


```{r eval=FALSE}
(2+3*2142)/1000 - 3.428
```


```{r eval=FALSE}
sqrt(sqrt(81))
```


```{r eval=FALSE}
2+7/3
```


# c(), rep() and seq()

Match each of the following bits of R-code with its output.

```{r}
seq(1,13,by=2)
```


```{r}
myNumbers <- seq(1,13,by=2)
myNumbers
```


```{r}
rep(5,times=3)
```

```{r}
rep(3,times=5)
```


```{r}
stoneAgeFriends <- c("fred","barney")
rep(stoneAgeFriends,times=3)
```


# Making a List of Numbers

Which of the following R commands will produce a list of numbers that begins with 50 and goes up through 90, counting by fives?

(That is, you want the numbers 50,55,60,65,70,75,80,85, and 90.)

```{r eval=FALSE}
seq(from=50,to=90,by=5)
```


```{r eval=FALSE}
seq(from=50,to=90)
```

```{r eval=FALSE}
rep(50,90,times=5)
```

```{r eval=FALSE}
c(50,55,60,65,70,75,80,85,90)
```

