---
title: "Your Title Here"
author: "Your Name Here"
date: "The Date Here"
output: html_document
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Introduction

The data for this project `KidsFeet`, is found in package `mosaic`:

```{r eval=FALSE}
data(KidsFeet)
View(KidsFeet)
help(KidsFeet)
```

The data are the result of an observational study of 39 fourth grade students in Ann Arbor, MI in October 1997.  (See the article "Wider Shoes for Wider Feet?", *Journal of Statistics Education*, 14(1), linked to [here](www.amstat.org/publications/jse14n1/datasets.meyer.html)).

We will assume that these 39 children are like a simple random sample from all fourth-grade children.

The researchers who collected the data had a number of interesting questions in mind.  Here we will focus on only one question that is suggested by the data:

>*Is there a relationship, in fourth-grade-age children, between which foot is wider and which hand is the dominant hand?*

# Methods

<Begin with a variable analysis.  Based on that analysis, choose numerical and graphical descritpive methods to examine the research question for your data,and choose an appropriate inferential method as well.>

# Results


## Descriptive Results


### Numerical Descriptive Results

<Use your numerical results to describe the patter (if any) that you find in your data.  Your discussion shoudl focus on using th data to address the Research Question.>

## Graphical Results

<Use your graphical results to describe the patter (if any) that you find in your data.  Your discussion shoudl focus on using th data to address the Research Question.>

## Inferential Results

<State hypotheses clearly.  Make sure that your discussion of inferential test addresses all of the important features of the output such as test statistic, P-value, etc.  Make a decision about the Null Hypothesis.  Be sure to choose a routine that produces a P-value approximation on which you can rely.>

# Discussion and Conclusion

