---
title: "Choose a Good Title!"
author: "Put Your Name Here!"
date: "put the date here!"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Introduction

Can you tell, from how much a fellow weighs, how fast he can a drink a can of pop or beer?  The package `tigerstats` contains the data frame `chugtime`:

```{r eval=FALSE}
data(chugtime)
View(chugtime)
help(chugtime)
```

This data was gathered by students who challenged several of their male friends to chug down a certain beverage--of the alcoholic variety, most likely.  The students also weighed each subject in the study.  The Research Question was:

  >*What is the relationship, if any,  between the weight of a college-aged male and the time it takes him to drink a can of beverage?*
  
In what follows we would like to analyze the researcher's data to see if we can answer their question

# Methods

<Decide on your methods:  begin with a variable analysis, then then choose your methods based on the outcome of that nalysis.  Make sure to choose at least one graphical method and at least one numerical method.>


# Results

<Note that the Results ection is divided into two sub-sections:  one for your graphical results and one for your numerical results.>

## Graphical Results

<Graphical results go hereProvide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Also, include a regression line in your graph.  Then discuss your graph, with particular emnphasis on specific features of the graph that help to answer your research question.>

## Numerical Results

<Numerical results go here.  This is where you will make your linear model.  Describe all important features of the output.  Provide a practical interpretation of the slope of the regression line.  Also, use the regression line to predict the chug-time of a person who weighs 170 pounds.  By how much is that prediction liable to be in error?>

# Discussion and Conclusion

<Discussion and conclusion here.  If you found a relationship between weight and chug-time, consider whether that relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  It might help to look at Chapter 6 before finishing this part of your report.>