---
title: "DA"
author: "Your Name Here"
date: "Friday, March 13 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# One Mean

Same population of guys as in the slides:

* $\mu =72$
* $\sigma =3$

This time you pick n=25 guys at random and compute the sample mean x-bar of their heights.


## Part a

What do you expect the sample mean to be? 


## Part b

Give or take about how much?  If you need to use a "calculator", write the necessary computations in the code chunk below:

```{r}

```

# One Proportion



Suppose that 45% of all registered voters in KY favor Alison Lundergren Grimes in the Senate race.

Suppose you take a random sample of 635 registered voters.

## Part a

About what percentage of the students in your sample will favor Alison? 

## Part b

Give or take about how much?  You will need to use R as a calculator.  Put your computations in the code chunk below:

```{r}

```

Turn your answer into a percentage:  othee people may find a percentage easier to understand!


