---
title: "DA"
author: "Your Full Name Here"
date: "01/30/2015"
output: pdf_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```



# Inference for Sex and Belief in Love at First Sight

Try the simulation app for the question about sex and belief in love at first sight:


```{r eval=FALSE}
require(manipulate)
ChisqSimSlow(~sex+love_first,data=m111survey)
```

How many times did you re-sample?



What percentage of the time did the re sampled chi-square statistic exceed the chi-square statistic in the actual study?



Do you think there is overwhelming evidence that in the GC population the two sexes differ in whether they believe in love at first sight?


# chisqtestGC() on the sex and love at first sight Question

Here's the code for a chi-square test to see if sex and belief in love at first sight are related in the GC population.  Run the code:

```{r}
chisqtestGC(~sex+love_first,data=m111survey,
            graph=TRUE)
```

Now look at the output and answer these question:

What's the test statistic?



About how big should it be if the Null is correct?



What's the P-value?


# Race and Gun Owndership

Are race and gun owndership related in the U.S. population?  In the code chunk below, insert the code needed to use chisqtestGC() to investigate this question.  **Tip**:  copy-paste and then modify the code from the previous problem.


```{r}

```


Looking at the output, answer the following questions.

What's the test statistic?



About how big should it be if the Null is correct?



What's the P-value?



Do you think we have strong evidence for a relationship in the population, or could the pattern in the data be due just to chance?




Do you think we have strong evidence for a relationship in the population, or could the pattern in the data be due just to chance?


Extra Practice:  Knit this document to pdf.  Does it knit correctly?  Did you put your own name at the top?




