---
title: "DA"
author: "Your Name Here"
date: "Friday, April 10 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```


# Spider Speeds (one last time!)

Does losing a pedipalp affect the sprinting speed of a Tidarren spider?

Let $\mu_d$ be the mean difference in speed (after amputation minus before) for all Tidarren spiders.

$H_0:  \mu_d = 0$

$H_a:  \mu_d \neq 0$


## Part a


Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~jiffrating-greatvaluerating,
      data=labels, mu=0)
```


## Part b

What's the value of the test statistic?


By how many standard errors does the sample mean of differences $\bar{d}$ differ from 0?


What's the P-value?


If amputation makes no difference in speed, on average, then what's the chance of getting a test statistic at least as far form 0 as the one we actually got?


What should we do with the Null Hypothesis this time?  Reject it or not?


Do you think that the data provided strong evidence that amputation makes a difference in speed, on average?



# Sagebrush Crickets

What do you think? Should we trust the results of the t-test for this example? Or should we ask a statistician for advice on how to analyze the data inferentially?  Just write a few words about which way you are leaning, and why.




