---
title: "DA"
author: "Your Full Name Here"
date: "01/26/2015"
output: pdf_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```


# Love at First Sight

Now we will look at the `m111survey` data, and ask the Research Question:

>*Who is more likely to believe in love at first sight:  a GC gal or a GC guy?*

## Practice Part a

What are the two variables form the `m111survey` data frame that are involved in this Research Question?




Which one would you say is the explanatory variable?



Which one would you say is the response variable?



For each variable, say what type of variable it is:  factor or numerical.



## Practice Part b

Why would it not be correct to make density plots to study this Research Question?



## Practice Part c

The following code makes a two-way table of the responses about love at first sight, broken down by the sex of the respondent:

```{r}
sexLove <- xtabs(~sex+love_first,data=m111survey)
sexLove
```

Run this code.  How many people in the study were males who believe in love at first sight?




## Practice Part c

Insert code below to make row percents for the two-way table stored as `sexLove`:

```{r}

```



What percentage of the females believe in love at first sight?



What percentage of the males believe in love at first sight?



## Practice Part d

Here is some code (from the slides) that makes a bar-chart to study the relationship between batter hand and type of pitch thrown by Justin Verlander:

```{r eval=FALSE}
barchartGC(~batter_hand+pitch_type,data=verlander,
           main="Verlander's Pitches, by Batter Stance",
           type="percent")
```


In the chunk below, write the code to make a bar-chart to study the relationship between **sex** and **love_first** in the `mat111survey` data.  (**Tip**:  Copy and paste the code from the previous chunk, and then make the necessary changes.)

```{r}

```


# Medians and Boxplots

## Pratice a

Here's the code to make `favstats()` on the speeds of Justin Verlander's pitches:

```{r eval=FALSE}
favstats(~speed,data=verlander)
```


In the code chunk below, enter the code to make `favstats()` for the fastest speeds driven by the students in the `m111survey` data.

```{r}

```


What is the median of the fastest speeds?


The first quartile?


The third quartile?


The interquartile range (IQR)?



## Practice b

Here's the code to make a box plot of the fastest speeds, broken down by sex:

```{r}
bwplot(fastest~sex,data=m111survey,
       main="Fastest Speed at GC, by Sex",
       xlab="sex",
       ylab="speed (mph)")
```


Make your own code chunk below, and inside it put the code needed for a box-plot of the speed of Justin Verlander's pitches, broken down by the type of pitch he threw.  **Tip**:  After you insert eh chunk, copy the code above and paste it into the chunk, then modify it so that it does what you want.





# Kim Kardashian

Ten thousand people were asked to give their "temperature rating" of Kim Kardashian, on a scale from 0 to 100.  0 means you don't like her at all, and 100 means you like her very much.  Here is some code for a density plot of the ratings:

```{r}
densityplot(~kkardashtemp,data=imagpop,
            main="Kim Kardashian Ratings",
            xlab="rating",
            from=0,to=100,
            plot.points=FALSE)
```


Is the distribution symmetric or skewed?


Is the distribution unimodal, bimodal, or neither?


True or False:  Most of the ratings were between 40 and 60.



# Justin Verlander Again

## Practice Part a

About 68% of Justin Verlander's four-seam fast were between ? and ??  (Give the two speeds.)



## Practice Part b

About what percentage of the time were his four-seam fastballs faster than 100 miles per hour?


## Practice Part c

Find a speed so that about 16% of his four-seam fastballs were slower than that speed.


