---
title: "DA"
author: "Your Full Name Here"
date: "8/28/2014"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# R as a Calculator

## Practice Part a

We can use R to find 2+7-3*5.  Run the code chunk below to see if it works:

```{r}
2+7-3*5
```

You should get -6.

## Pratice Part b

Next, insert write into the following the code needed to (25.5-20.2)/3:

```{r}

```

## Practice Part c

Next, insert your own code chunk below and write into it the code needed to produce the 12 squared plus 10 squared.




## Practice Part d

Finally, insert your own code chunk and put write into it one line of code that does all of the following:

* squares 3.1, then divides it by 25
* squares 5.2, then divides it by 36
* adds the previous two results together
* then takes the square root of the result



# Data Frames

## Practice Part 1

Run the following code chunk to learn about `m111survey`:

```{r eval=FALSE}
data(m111survey)
View(m111survey)
help(m111survey)
```

How many students took part in this survey?


## Practice Part b

Is **fastest** a numerical variable or a factor variable?


Is **seat** a numerical variable or a factor variable?


## Practice Part c

How many students did not answer the question about their ideal height?


# Making Lists with R

## Practice Part a

In the following code chunk, insert the code needed to make a list of 100 fives.

```{r}

```


## Practice Part b

In the following code chunk, insert the code needed to make a list of all the numbers from 100 to 220, going up in steps of 10.

```{r}

```

