---
title: "DA"
author: "Your Name Here"
date: "Wednesday, March 11 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```



# More Coin-Flipping

When you flip the fair coin 400 times.

## Part A

Use the 68-95 Rule to answer the following two questions:

About what is the chance that you will get fewer than 190 heads?


About what is the chance that you will get more than 220 heads?


## Part B

Next, use `pbinomGC()` to compute the "exact" probabilities:

Chance of fewer than 190 heads:

```{r}
pbinomGC(189,size=400,prob=0.5,
         region="below",graph=TRUE)
```

Next, the chance of more than 220 heads:

```{r}
pbinomGC(220,size=400,prob=0.5,
         region="above",graph=TRUE)
```

Compare these the "exact" values with the approximations from the 68-95 Rule.  Not bad, eh?


# Bats

A population of bats has wingspans that are normally distributed, with a mean μ of 25 centimeters and a standard deviation of 5 centimeters.

Let X= the wingspan of a randomly-captured bat. Find each of the following probabilities, using `pnormGC()`.  Insert the code chunks!


## Part A:  P(X<27)



    
## Part B:  P(X>29)




## Part C:  P(20<X<30)



# Bat Percentiles

Back to the population of bats whose wingspans average 25 centimeters, with a standard deviation of 5 cm.  Use `qnorm()` or `qnormGC()`to answer the following two questions.  Insert code chunks as needed.



## Part A

Harry the Bat has a wingspan that is right at the 35th percentile for all the bats. What is Harry's wingspan?




## Part B

Dolores the Bat has a wingspan that is longer than all but 25% of the bats. What is her wingspan?








