---
title: "DA"
author: "Your Name Here"
date: "Friday, April 24 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# M&M's

## Part a

Put your actual color-counts in the code below.  This will create a list of your counts.

```{r}
observedColors <- c(blue=10,orange=10,green=10,
                    yellow=10,red=10,brown=10)
```

## Part b

Next, create a list of the proportions of colors made at the factory back in 2008 (as claimed on the website).  You can do this just by running this code:

```{r}
nullProbs <- c(0.24,0.20,0.16,0.14,0.13,0.13)
```

## Part c

OK, now let's test to see if it's reasonable to believe that your bag is like a random sample from M&M's at the factory.  Modify the code below, and then run it.

```{r}
chisqtestGC(obs,p=rep(1/6,6))
```

## Part d

What was the observed value of the chi-square statistic?


## Part e

If the bag is like a random sample from the M&M's in the factory, about what would you expect the chi-square statistic to be?




##  Part f


If the bag is like a random sample from the M&M's in the factory, about what is the chance of getting a chi-square statistic at least as big as the one you actually got?


## Part g

Based on the data, is it reasonable to believe that your bag is like a random sample of candies from the factory?



