---
title: "DA"
author: "Your Name Here"
date: "Tuesday February 6, 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```


# Beetles and Stumps

## Part a

We decided that **larvae** and **stumps** are both numerical variables.

In our study, which one do you think is the explanatory variable?



Which one is the response?


## Part b

In the code-chunk below, write the code needed to make a scatter plot to investigate the relationship between **larvae** and **stumps**.  (**Tip**:  Find code from the slides that makes a scatter plot, then copy it into the eac of the two code vchunks below.  Using the top copy as a reference, modify the bottom copy so that it fits our situation.  Try it out.  When you are sure that it works, delete the top code chunk.)

Remember:  the explanatory variable should appear on the x-axis!


```{r}

```



```{r}

```


## Part c

Examining the scatter plot, describe the relationship:  is it positive, negative, curvilinear, or is there no relationship at all?


True or False:  the more stumps there are in a plot, the more larvae cluster tend to be found in it.


## Speed and Fuel Efficiency

## Part a

We decided that **speed** and **fuel** are both numerical variables.

In our study, which one do you think is the explanatory variable?



Which one is the response?


## Part b

Create a code chunk below and write into it the code to make a scatter plot to study the relationship between driving speed and fuel efficiency.  Remember:  the explanatory variable should be on the x-axis!






## Part c

Describe the relationship:  is it positive linear, negative linear, curvilinear, or is there no relationship?



## Part d

Units for **efficiency are** "liters per 100 km", the number of liters of fuel required to drive the car 100 kilometers at the given speed.

Does a high number for efficiency represent good or poor fuel efficiency?


At about what speed does the car have the best fuel efficiency?


# Justin Verlander's Pitches

Look at the scatter plots of **pz** vs. **px** for Justin Verlander's pitches.

For which types of pitch does there appear to be a relationship between the horizontal break over the plate (**px**) and the height of the pitch over the plate (**pz**)?

