---
title: "DA"
author: "Your Name Here"
date: "Tuesday February 11, 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```



# Beetles and Stumps

## Part a

In the code chunk below, we have the code need to predict **larvae** (the number of beetle larvae clusters) from **stumps**, the number of stumps in each plot:

```{r}
beetleMod <- lmGC(larvae~stumps,data=stumps)
beetleMod
```

Run the code chunk and examine the output.

What is the correlation coefficient r?  Based on r, would say the linear relationship is weak, moderate or strong?  (It's a judgement-call on your part?)



## Part b

Here is the code needed to predict the number of larvae cluster in a plot with 3 stumps:

```{r}
predict(beetleMod,x=3)
```


By how much is this prediction liable to be in error?



## Part d

What is the R-squared value?



# Latitude and Temperature

## Part a


In the code chunk below, insert the code needed to make a linear model to predict the temperature of a U.S. city from its latitude:

```{r}
tempLat <- # replace this remark (including the #) with the code needed
tempLat
```


## Part b

Use the model `tempLat` predict the August temperature of Georgetown, KY (latitude 39.21 deg N).  Insert the needed code in the chunk below:

```{r}

```


By about how much is the prediction liable to be in error?


## Part c

Use the `tempLat` model to predict the August temperature of Singapore (1.3 deg N).  Insert the needed code in the chunk below:

```{r}

```


Is your prediction reliable?  Why or why not?


# Gas Bill vs. Temperature

## Part a

Here is the code for a linear model to predict gasbill from temp:

```{r}
billMod <- lmGC(gasbill~temp,data=Utilities)
```

Run this code.  You've made the model.

Now use it to predict the gasbill in a billing period when the average temperature is 80 degrees.  Put the needed code into the chunk below:

```{r}

```


## Part b

Construct a second-degree polynomial model to predict gasbill from temp by running the code below:

```{r}
billMod2 <- polyfitGC(gasbill~temp,data=Utilities,degree=2)
```


Let's use the model to predict the gasbill in a billing period when the average temperature is 80 degrees.  Write the code you need for this in the chunk below:

```{r}

```


## Part c

Run this check for the linear fit:

```{r}
lmGC(gasbill~temp,data=Utilities,check=TRUE)
```


Then run a check on the quadratic fit:

```{r}
polyfitGC(gasbill~temp,data=Utilities,degree=2,check=TRUE)
```

Based on your checks, which of the predictions made in Part b above is more reliable?

