---
title: "DA"
author: "Your Name Here"
date: "Monday, March 23 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# One Proportion


Suppose that 45% of all registered voters in KY favor Alison Lundergren Grimes in the Senate race.

Suppose you take a random sample of 635 registered voters.

## Part a

About what percentage of the students in your sample will favor Alison? 

## Part b

Give or take about how much?  You will need to use R as a calculator.  Put your computations in the code chunk below:

```{r}

```

Turn your answer into a percentage:  othee people may find a percentage easier to understand!


# Sampling 64 College Students


In a population of college students, the mean fastest speed ever driven is 110 mph. The SD of the fastest speeds is 10 mph.

Herbert will sample 64 college students at random and compute the sample mean x-bar.

Let's find the approximate probability that x-bar will turn out to be less than 107 mph.

We do this in steps.

## Expected Value

What is the expected value of x-bar?


## Standard Deviation

What is the standard deviation of x-bar:  is it 10, or is it 10/sqrt(64)?


## Probability

Since the sample size of 64 is pretty large,  the Central Limit Theorem says that x-bar is approximately normal, as a random variable.

What's the correct way to compute the chance that x-bar will be less than 107 mph?

Is it this:

```{r}
pnormGC(107,region="below",mean=110,sd=10)
```

or this?

```{r}
pnormGC(107,region="below",mean=110,sd=10/sqrt(64))
```

Explain your choice briefly.



