---
title: "DA"
author: "Your Name Here"
date: "Monday, April 13 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```

# Miguel Cabrera



In the 2014 Major League season, Detroit Tiger Miguel Cabrera had a batting average of .313 on 611 at-bats. Can we conclude from this data that in 2014 his chance of getting a hit was more than 30 percent?

Define p to be the probability that Cabrera gets a hit when he is at bat.

$H_0:  p = 0.3$

$H_a:  p \neq 0.3$


## Part a

You will perform the two sided test with binomtestGC(), using the summary data given above.  To do this modify the code form our previous example:

```{r}
binomtestGC(x=1325,n=2500,p=0.50,graph=TRUE)
```


## Part b


Can you conclude that Cabrera's chance was better than 30%?  Why or why not?



# Political Party and Belief in Astrology


## Part a

Conduct the desired two-sided test by modifying the code in the example below:

```{r}
proptestGC(x=c(899,1183),
           n=c(1308,1824),
           p=0,graph=TRUE)
```


## Part b

What is the P-value?

## Part c

What's the value of the test statistic?


## Part d

If Democrats and Republicans out in the population at large are equally likely to believe in astrology, then what is the chance of getting a test statistic at least as far from 0 as the one we actually got in this study?


## Part e

What should we do about the Null Hypothesis:  reject it, or not?


## Part f

Did this study provide strong evidence that Democrats are more likely than Republicans to believe in astrology?

