---
title: "DA"
author: "Your Full Name Here"
date: "01/28/2015"
output: pdf_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```



# Kim Kardashian

Ten thousand people were asked to give their "temperature rating" of Kim Kardashian, on a scale from 0 to 100.  0 means you don't like her at all, and 100 means you like her very much.  Here is some code for a density plot of the ratings:

```{r}
densityplot(~kkardashtemp,data=imagpop,
            main="Kim Kardashian Ratings",
            xlab="rating",
            from=0,to=100,
            plot.points=FALSE)
```


Is the distribution symmetric or skewed?


Is the distribution unimodal, bimodal, or neither?


True or False:  Most of the ratings were between 40 and 60.



# Justin Verlander Again

## Practice Part a

About 68% of Justin Verlander's four-seam fast were between ? and ??  (Give the two speeds.)



## Practice Part b

About what percentage of the time were his four-seam fastballs faster than 100 miles per hour?


## Practice Part c

Find a speed so that about 16% of his four-seam fastballs were slower than that speed.


# z-Scores

## Practice Part a

In the chunk below, write the math code to find the z-score for a Justin Verlander four-seam fastball that has a speed of 91 mph.

```{r}

```


Would you say this fastball was unusually slow for Verlander?



## Practice Part b

In the chunk below, write the math code to find the z-score for a Justin Verlander four-seam fastball that has a speed of 95 mph.

```{r}

```


Would you say this fastball was unusually slow for Verlander?




