---
title: "DA"
author: "Your Name Here"
date: "Tuesday, April 28 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
library(readr) # for reading csv files
library(readxl) # for reading excel files
```


# Miguel Cabrera

In the 2014 Major League season, Detroit Tiger Miguel Cabrera had a batting average of .313 on 611 at-bats. Can we conclude from this data that in 2014 his chance of getting a hit was more than 30%?

Define $p$ to be the probability that Cabrera gets a hit when he is at bat.

* $H_0: p=0.3$
* $H_a: p \neq 0.3$

Perform the two sided test with binomtestGC(), using the summary data:

What should you enter for n, the number of trials?

What should you enter for x, the number of successes?

What should you enter for p?

Can you conclude that Cabrera's chance was better than 30%?



# Reading in CSV Files

In the code chunk below, paste the code needed to read in the `tornado` data:

```{r}

```

Then make a density plot of the tornado damages:

```{r}
densityplot(~ damage, data = tornado, from = 0,
            xlab = "mean damage (millions of $$)")
```


# Reading an Excel File

Below is a URL for an Excel workbook.  Copy the it onto your Clipboard:

https://dl.dropboxusercontent.com/u/30396928/datasets.xlsx

Now use the Import Dataset Utility to read in the data set `chugtime`.  Note that the workbook has two sheets.  You have to choose the correct sheet, and skip any useless lines.

Before you exit the utiltity, copy the code needed to read in the data into the chunk below:

```{r}

```

Now make yourself a nice scatterplot to examine the relationship between a frat boy's weight and the time it takes for him to chug 12 ounces of a certain beverage:

```{r}
xyplot(ChugTime ~ Weight,
	data = chugtime,
	type = c("p","r"),
	pch = 20,
	xlab = "weight (pounds)",
	ylab = "time to chug (seconds)")
```




