---
title: "DA"
author: "Your Name Here"
date: "Tuessday, April 12 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
```

# Exploring Safety Checks

## Part A

Try the Shiny app "Intervals for One Mean" with:

* the population set at "Way-Out Outlier Group"
* sample size at n = 10
* confidence level set at 95%

Take at least 10,000 samples.

What percent of the time did your confidence intervals contain the mean of the population?



Describe the distribtuion of the t-statistics? Does it resemble the t-curve very closely?

## Part B

Try the same things again, but with sample size set at n = 50. Even at sample size 50, can you fully "trust" confidence intervals based on samples from this population?



# Seating Preference at GC

Suppose you want to know:  Do more than one-third of all GC students prefer to the in the Middle?

## Part a

To answer this question, you will make a 95%-confidence interval for p, the proportion of all GC students who prefer the middle. Do this by modifying the code in the chunk below:

```{r}
binomtestGC(~sex,data=m111survey,success="female")
```

## Part b

What is the confidence interval for p?  (Give the lower and upper bounds.)


## Part c

Based on the the interval, can we conclude that more than one-third of all GC students prefer the Middle?  Why or why not?



# Barak Obama Approval

Out of 2500 randomly-selected voters, 1200 says that they approve of the job that Barak Obama is doing..

Make a 99%-confidence interval for p, the proportion of all voters who approve of the job Obama is doing. Do this by modifying the code in the chunk below:

(Hint: Add the argument conf.level=0.99 to your code.)

```{r}
binomtestGC(x=490,n=1000)
```



# UC-Davis


```{r eval=FALSE}
data(ucdavis1)
View(ucdavis1)
help(ucdavis1)
```

Let p1 be the proportion of all UC-Davis females who prefer to sit in the front.

Let p2 be the proportion of all UC-Davis males who prefer to sit in the front.

Your job is to make a 90%-confidence interval for p1−p2.  Do this by modifying the code below.  Here are some tips:

* use the `success` argument to count people who prefer the Front
* use the `first` argument to say which group should come first
* add the `conf.level` argument for confidence levels other than 95%


```{r}
proptestGC(~sex+extra_life,data=m111survey,
           success="yes",first="female")
```


# Difference of Two Proportions, Summary Data

* In a random sample of 2000 adults in the year 1940, 800 had completed high school.
* In a random sample of 2500 adults in the years 1991, 1900 had completed high school.


Let p1 be the proportion of all adults in 1991 who had completed high school.

Let p2 be the proportion of all adults in 1940 who had completed high school.

Find a 95%-confidence interval for p1−p2.  Do this by modifying the code below:

```{r}
proptestGC(x=c(23,33),n=c(100,110))
```



# A Tall Sample

You want to estimate the mean height of all GC students, so you record the heights of the students in your KHS Coaching basketball class. The mean height is 73 inches. Realizing that your sample may be biased toward taller students, you compute a 95% confidence interval for the mean. It turns out to be 71 to 75 inches.

True or False (and explain briefly):  "You can be about 95%-confident that the mean height of all GC students is between 71 and 75 inches.""


