---
title: "DA"
author: "Your Full Name Here"
date: "02/04/2016"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tigerstats)
```

# Sex and Seat:  Simulation

The code chunk below performs a chi-square test on the sex and seat question, using simulation.  A seed has been set, so the results will be the same whenver you run the code.  Run it now.


```{r}
set.seed(31415)
chisqtestGC(~sex+seat,data=m111survey,
            simulate.p.value="random",
            B=3000,graph=TRUE)
```

What's the P-value?

Here's the code we used earlier (no simulation).  Run it.

```{r}
chisqtestGC(~sex+seat,data=m111survey)
```


What's the P-value?


Is it about the same as the P-value you got using simulations?



# Beetles and Stumps

## Part a

We decided that **larvae** and **stumps** are both numerical variables.

In our study, which one do you think is the explanatory variable?



Which one is the response?


## Part b

In the code-chunk below, write the code needed to make a scatter plot to investigate the relationship between **larvae** and **stumps**.  (**Tip**:  Find code from the slides that makes a scatter plot, then copy it into the eac of the two code vchunks below.  Using the top copy as a reference, modify the bottom copy so that it fits our situation.  Try it out.  When you are sure that it works, delete the top code chunk.)

Remember:  the explanatory variable should appear on the x-axis!


```{r}

```



```{r}

```


## Part c

Examining the scatter plot, describe the relationship:  is it positive, negative, curvilinear, or is there no relationship at all?


True or False:  the more stumps there are in a plot, the more larvae cluster tend to be found in it.


## Speed and Fuel Efficiency

## Part a

We decided that **speed** and **fuel** are both numerical variables.

In our study, which one do you think is the explanatory variable?



Which one is the response?


## Part b

Create a code chunk below and write into it the code to make a scatter plot to study the relationship between driving speed and fuel efficiency.  Remember:  the explanatory variable should be on the x-axis!






## Part c

Describe the relationship:  is it positive linear, negative linear, curvilinear, or is there no relationship?



## Part d

Units for **efficiency are** "liters per 100 km", the number of liters of fuel required to drive the car 100 kilometers at the given speed.

Does a high number for efficiency represent good or poor fuel efficiency?


At about what speed does the car have the best fuel efficiency?


# Beetles and Stumps

## Part a

In the code chunk below, we have the code need to predict **larvae** (the number of beetle larvae clusters) from **stumps**, the number of stumps in each plot:

```{r}
beetleMod <- lmGC(larvae~stumps,data=stumps)
beetleMod
```

Run the code chunk and examine the output.

What is the correlation coefficient r?  Based on r, would say the linear relationship is weak, moderate or strong?  (It's a judgement-call on your part?)



## Part b

Here is the code needed to predict the number of larvae cluster in a plot with 3 stumps:

```{r}
predict(beetleMod,x=3)
```


By how much is this prediction liable to be in error?



## Part d

What is the R-squared value?



# Latitude and Temperature

## Part a


In the code chunk below, insert the code needed to make a linear model to predict the temperature of a U.S. city from its latitude:

```{r}
tempLat <- # replace this remark (including the #) with the code needed
tempLat
```


## Part b

Use the model `tempLat` predict the August temperature of Georgetown, KY (latitude 39.21 deg N).  Insert the needed code in the chunk below:

```{r}

```


By about how much is the prediction liable to be in error?


## Part c

Use the `tempLat` model to predict the August temperature of Singapore (1.3 deg N).  Insert the needed code in the chunk below:

```{r}

```


Is your prediction reliable?  Why or why not?