---
title: "DA"
author: "Your Name Here"
date: "Tuesday, April 19 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
```

# Drinking at UC-Davis

```{r eval=FALSE}
data(ucdavis1)
View(ucdavis1)
help(ucdavis1)
```


Let $\mu$ be the mean number of drinks consumed each week, for all UC-Davis students.

At Penn State, the mean is known to be 4. Is it the same at UC-Davis?

$H_0:  \mu = 4$

$H_a:  \mu \neq 4$

## Part a

Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~GPA,data=m111survey,
         mu=3.3,alternative="two.sided",
         graph=TRUE)
```


## Part b

Do you think the study provides strong enough evidence to conclude that the mean number of drinks consumed at UC-Davis differs from 4?  Why or why not?


# Spider Speeds (one last time!)

Does losing a pedipalp affect the sprinting speed of a Tidarren spider?

Let $\mu_d$ be the mean difference in speed (after amputation minus before) for all Tidarren spiders.

$H_0:  \mu_d = 0$

$H_a:  \mu_d \neq 0$


## Part a


Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~jiffrating-greatvaluerating,
      data=labels, mu=0)
```


## Part b

What's the value of the test statistic?


By how many standard errors does the sample mean of differences $\bar{d}$ differ from 0?


What's the P-value?


If amputation makes no difference in speed, on average, then what's the chance of getting a test statistic at least as far form 0 as the one we actually got?


What should we do with the Null Hypothesis this time?  Reject it or not?


Do you think that the data provided strong evidence that amputation makes a difference in speed, on average?


# Spider Speeds (one last time!)

Does losing a pedipalp affect the sprinting speed of a Tidarren spider?

Let $\mu_d$ be the mean difference in speed (after amputation minus before) for all Tidarren spiders.

$H_0:  \mu_d = 0$

$H_a:  \mu_d \neq 0$


## Part a


Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~jiffrating-greatvaluerating,
      data=labels, mu=0)
```


## Part b

What's the value of the test statistic?


By how many standard errors does the sample mean of differences $\bar{d}$ differ from 0?


What's the P-value?


If amputation makes no difference in speed, on average, then what's the chance of getting a test statistic at least as far form 0 as the one we actually got?


What should we do with the Null Hypothesis this time?  Reject it or not?


Do you think that the data provided strong evidence that amputation makes a difference in speed, on average?

# Spider Speeds (one last time!)

Does losing a pedipalp affect the sprinting speed of a Tidarren spider?

Let $\mu_d$ be the mean difference in speed (after amputation minus before) for all Tidarren spiders.

$H_0:  \mu_d = 0$

$H_a:  \mu_d \neq 0$


## Part a


Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~jiffrating-greatvaluerating,
      data=labels, mu=0)
```


## Part b

What's the value of the test statistic?


By how many standard errors does the sample mean of differences $\bar{d}$ differ from 0?


What's the P-value?


If amputation makes no difference in speed, on average, then what's the chance of getting a test statistic at least as far form 0 as the one we actually got?


What should we do with the Null Hypothesis this time?  Reject it or not?


Do you think that the data provided strong evidence that amputation makes a difference in speed, on average?


# Type-I and Type-II Errors

I plan to test:

* $H_0:  \mu = 50$
* $H_a:  \mu \neq 50$

Unknown to me, the mean of the population is 52.


Can I make a Type-I error?


Can I make a Type-II error?


# Sagebrush Crickets

What do you think?  Should we trust the results of the t-test?  Or should we ask a statistician for advice on how to analyze the data inferentially?
