---
title: "DA"
author: "Your Name Here"
date: "Thursday, March 24 2015"
output: html_document
---

```{r include=FALSE}
library(knitr)
library(mosaic)
library(tigerstats)
```


# Counting the Females in a Random Sample


* You sample 71 people at random from a large population.
* You record the sex of each person (male/female).
* 55% of the population is female

Let X = the total number of females in the sample.

Is X binomial? If so, what is n and what is p?


# Buzz at 70%

If Buzz can get it right only 70% of the time when Doris helps, then what is the chance that he gets it right 15 or more times out of 16 tries?

Use `pbinomGC()` to find out.  Write your code into the following chunk:

```{r}

```


# Coin Flipping

You flip a far coin 100 times.  Let X= the number of heads you get.

## Part A

Use `pbinomGC()` to find the chance that X is 45 or less.


```{r}

```

## Part B

What is the chance that X is less than 47 heads?

```{r}

```

## Part C

What is the chance that X is between 40 and 60 heads?

```{r}

```


## Part D

What is the chance that X is exactly 50 heads?  (Hint:  being exactly 50 is the same as being "between" 50 and 50.)

```{r}

```


# Free-Throw Shooting

You are a 70% free-throw shooter.  You stand at the line and take 20 shots.

The number of shots you make is a binomial random variable.  What is the n, the number of trials?


What is p, the chance of success on each trial?


How many shots do you expect to make?


Compute the SD, the give-or-take figure, using R as a calculator.  Use the code-chunk below.  Run the code to see what you get.

```{r}

```

# More Coin-Flipping

When you flip the fair coin 400 times.

## Part A

Use the 68-95 Rule to answer the following two questions:

About what is the chance that you will get fewer than 190 heads?


About what is the chance that you will get more than 220 heads?


## Part B

Next, use `pbinomGC()` to compute the "exact" probabilities:

Chance of fewer than 190 heads:

```{r}
pbinomGC(189,size=400,prob=0.5,
         region="below",graph=TRUE)
```

Next, the chance of more than 220 heads:

```{r}
pbinomGC(220,size=400,prob=0.5,
         region="above",graph=TRUE)
```

Compare these the "exact" values with the approximations from the 68-95 Rule.  Not bad, eh?


# Bats

A population of bats has wingspans that are normally distributed, with a mean μ of 25 centimeters and a standard deviation of 5 centimeters.

Let X= the wingspan of a randomly-captured bat. Find each of the following probabilities, using `pnormGC()`.  Insert the code chunks!


## Part A:  P(X<27)



    
## Part B:  P(X>29)




## Part C:  P(20<X<30)



# Bat Percentiles

Back to the population of bats whose wingspans average 25 centimeters, with a standard deviation of 5 cm.  Use `qnorm()` to answer the following two questions.  Insert code chunks as needed.

## Part A

Harry the Bat has a wingspan that is right at the 35th percentile for all the bats. What is Harry's wingspan?




## Part B

Dolores the Bat has a wingspan that is longer than all but 25% of the bats. What is her wingspan?








