---
title: "DA"
author: "Your Name Here"
date: "Tuesday, March 16 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
```

# Which Game do You Prefer?

In Game A, you you win either 2 dollars or 4 dollars, each time you play.  (%0% chance of either outcome).

In Game B you have a 75% chance to lose 8 dollars and a 25% chance to win 40 dollars, each time you play.

You can play Game A 10,000 times, or you can play Game B 10,000 times.

## Part A

Which would you prefer?  State a preference right now, before we do any math, and briefly explain your preference.


## Part B

Now let's look at the question from the point of view of expected value.  The following code chunk gives you the expected value of Game A:

```{r}
p.a <- c(1/2,1/2)
a <- c(2,4)
evA <- sum(a*p.a)
evA
```

The next code chunk gives you the expected value for game B:

```{r}
p.b <- c(3/4,1/4)
b <- c(-8,40)
evB <- sum(b*p.b)
evB
```

Do these computations affect your original answer?  Why or why not?


## Part C

Now let's examine the question using simulation.  The following code chunk illustrates what your total winnings could be if you played Game A 10,000 times:

```{r}
winningsA <- sample(a,size=10000,p=p.a,replace=TRUE)
sum(winningsA)
```

Run the chunk a few times.

The next code chunk illustrates what your total winnings might be like if you played Game B 10,000 times:

```{r}
winningsB <- sample(b,size=10000,p=p.b,replace=TRUE)
sum(winningsB)
```

Run this chunk a few times.

Does the results alter your original answer?  Why or why not?