---
title: "DA"
author: "Your Full Name Here"
date: "02/23/2016"
output: html_document
---

```{r, include=FALSE}
library(tigerstats)
library(tigerTree)
```

## Predicting the Speed of a Verlander Pitch

Let's make a regression tree to predict the speed of a Justin Verlander pitch using other variables in the `verlander` data frame.

First of all, we'll remove an outlier (the one very slow pitch) and we'll remove **gamedate** from the data frame:

```{r}
ver2 <- subset(verlander, speed > 60)
ver2$gamedate <- NULL
```

In the code chunk below, replace the `??` with the code needed to make a tree model to predict **speed** from the other variables in `ver2`:

```{r}
trSpeed <- tree(speed ~ ., data = ver2)
plot(trSpeed); text(trSpeed)
```

Let's figure out what the tree is telling us.  The code below gives you the values if the factor variable **pitch_type**:

```{r}
levels(ver2$pitch_type)
```

In the tree they are coded by the first five lower-case letters of the alphabet.  If you want to see them all together, run this code:

```{r}
data.frame(code = letters[1:5], type = levels(ver2$pitch_type))
```

If Verlander throws a four-seam fastball (FF), what does the tree predict its speed will be?


If Verlander throws a curve ball, what does the tree predict its speed will be?

The code below give a summary of the tree:

```{r}
summary(trSpeed)
```

What is the residual mean square error?


What is the root mean square error (RMSE)?  **Hint:** Just take `sqrt()` of the previous answer!





