---
title: "DA"
author: "Your Full Name Here"
date: "02/02/2016"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tigerstats)
```


# Kim Kardashian

Ten thousand people were asked to give their "temperature rating" of Kim Kardashian, on a scale from 0 to 100.  0 means you don't like her at all, and 100 means you like her very much.  Here is some code for a density plot of the ratings:

```{r}
densityplot(~kkardashtemp,data=imagpop,
            main="Kim Kardashian Ratings",
            xlab="rating",
            from=0,to=100,
            plot.points=FALSE)
```


Is the distribution symmetric or skewed?


Is the distribution unimodal, bimodal, or neither?


True or False:  Most of the ratings were between 40 and 60.



# Justin Verlander Again

## Practice Part a

About 68% of Justin Verlander's four-seam fast were between ? and ??  (Give the two speeds.)



## Practice Part b

About what percentage of the time were his four-seam fastballs faster than 100 miles per hour?


## Practice Part c

Find a speed so that about 16% of his four-seam fastballs were slower than that speed.

# Row Percents for Sex and Seat

What percentage of females preferred the back?  In the code chunk below, enter the math to find this percentage:

```{r}

```


What percentage of males preferred the back?  In the code chunk below, enter the math to find this percentage:

```{r}

```


# Sex and Belief in Love at First Sight

The code-chunk below gives code for a two way-table for the sex and seat question, followed by a table of row percents, and a barchart.  Run the code to see what it does.

```{r}
sexSeat <- xtabs(~sex+seat,data=m111survey)
sexSeat
rowPerc(sexSeat)
barchartGC(sexSeat,type="percent",flat=TRUE)
```

In the code chunk below, enter the code needed to make a two-way table, a table of row percents and a barchart for the question about the relationship between sex and belief in love at first sight.  **Tip**:  copy the above code into the chunk, then modify it.

```{r}

```

# Inference for Sex and Belief in Love at First Sight

Try the simulation app for the question about sex and belief in love at first sight:


```{r eval=FALSE}
chisqSh(~ sex + love_first, data = m111survey)
```

How many times did you re-sample?



What percentage of the time did the re sampled chi-square statistic exceed the chi-square statistic in the actual study?



Do you think there is overwhelming evidence that in the GC population the two sexes differ in whether they believe in love at first sight?



