---
title: "DA"
author: "Your Full Name Here"
date: "02/18/2016"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tigerstats)
library(tigerTree)
```

## Predicting the Species of an Iris

```{r eval = FALSE}
data(iris)
View(iris)
help(iris)
```

Use the code chunk below to make a tree model to predict **Species** from the other four variables in the `iris` data:

```{r}
tr.iris <- tree(Species ~ ., data = iris)
```

Now make a plot fo the model:

```{r}
plot(tr.iris); text(tr.iris)
```

Use the code chunk below to get a summary of your tree model:

```{r}
summary(tr.iris)
```

How many temrinal nodes does your tree have?


What's the mis-classification rate?


In the code chunk below, highlight the name `iris` and then head to the Cloud Plot Addin.  Make a cloud plot of the three varables that were actually used to constrcut the tree, and then group by **Species**.  Fix your plot up  nicely, and press Done to get the code for the plot in the chunk below.

```{r}
iris
```

Think about how/whether the tree model makes sense in term of the cloud plot you made.





