---
title: "DA"
author: "Your Name Here"
date: "Tuesday, April 26 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
```


# Spider Speeds (one last time!)

Does losing a pedipalp affect the sprinting speed of a Tidarren spider?

Let $\mu_d$ be the mean difference in speed (after amputation minus before) for all Tidarren spiders.

$H_0:  \mu_d = 0$

$H_a:  \mu_d \neq 0$


## Part a


Perform the test of significance.  Do this by modifying the code below:

```{r}
ttestGC(~jiffrating-greatvaluerating,
      data=labels, mu=0)
```


## Part b

What's the value of the test statistic?


By how many standard errors does the sample mean of differences $\bar{d}$ differ from 0?


What's the P-value?


If amputation makes no difference in speed, on average, then what's the chance of getting a test statistic at least as far form 0 as the one we actually got?


What should we do with the Null Hypothesis this time?  Reject it or not?


Do you think that the data provided strong evidence that amputation makes a difference in speed, on average?





# Type-I and Type-II Errors

I plan to test:

* $H_0:  \mu = 50$
* $H_a:  \mu \neq 50$

Unknown to me, the mean of the population is 52.


Can I make a Type-I error?


Can I make a Type-II error?


# Sagebrush Crickets

What do you think?  Should we trust the results of the t-test?  Or should we ask a statistician for advice on how to analyze the data inferentially?


# Miguel Cabrera

In the 2014 Major League season, Detroit Tiger Miguel Cabrera had a batting average of .313 on 611 at-bats. Can we conclude from this data that in 2014 his chance of getting a hit was more than 30%?

Define $p$ to be the probability that Cabrera gets a hit when he is at bat.

* $H_0: p=0.3$
* $H_a: p \neq 0.3$

Perform the two sided test with binomtestGC(), using the summary data:

What should you enter for n, the number of trials?

What should you enter for x, the number of successes?

What should you enter for p?

Can you conclude that Cabrera's chance was better than 30%?


