---
title: "DA"
author: "Your Name Here"
date: "Tuesday March 8, 2016"
output: html_document
---

```{r include=FALSE}
library(tigerstats)
```


# A Confounding Variable

Consider the following two variables, measured on adult males:

* **bifocals**:  whether or not you wear bifocals
* **prostate**:  whether or not you have prostate cancer

An observational study shows that **bifocals** and **prostate** are related:  the men in the study who reported wearing bifocals were about five times more likely to have prostate cancer than were the men who did not wear bifocals.

## Part a

What third variable might be a confounding variable in this study?


## Part b

Explain in a brief sentence why it could be confounder.


# Knife or Gun

Here is some code to randomly assign subject to treatment groups (i.e., a completely randomized design for this experiment).  After the assignment, `xtabs()` alows you to examine the two treatment groups:


```{r}
assign <- RandomExp(subjects,sizes=c(10,10),groups=c("knife","gun"))
assign
xtabs(~treat.grp+loud,data=assign)
```

Run the code a few times?  Do the groups always look the same with respect to hog-hollerin'?



Why is is a problem for the experiment if one group has more hog-hollerers than the other group does?



## Attitudes:  Race and Sentence

We are interested in the following Research Question:

Who recommends higher sentences, on average:

* subjects given a form where the defendant's suggested race is Black?
* or subjects given a form where the defendant's suggested race is White?


## Part a

Is the data for our Research Question based on an experiment or on an observational study?



## Part b

We could have asked each subject the same question twice:

* once for Tyrone Marcus Watson
* once for William Shane Winchester

Then we would have had a repeated measures design.

Why did we not do this?



